---
title: "02_supplement_dynamic_dfm"
author: "Yiming Zhao"
date: "2026-02-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)

library(tseries)
library(ggplot2)
library(reshape2)
library(patchwork)
library(dfms)
library(vars)
library(missMDA)
library(zoo)
library(knitr)
library(kableExtra)
library(KFAS)
library(tidyverse)

source("ifo_data_processing.R")
theme_set(theme_minimal())
```

# Data Preparation
```{r}
data_wide   <- read_csv("data_b_wide.csv")

dates <- as.Date(data_wide$date)
var_names <- colnames(data_wide)[-1]

# Extract industry and question code from variable name
extract_industry_code <- function(var_name) {
  strsplit(var_name, "_")[[1]][1]
}

extract_question_code <- function(var_name) {
  strsplit(var_name, "_")[[1]][2]
}

# Create metadata
var_metadata <- data.frame(
  var_name = var_names,
  industry_code = sapply(var_names, extract_industry_code),
  stringsAsFactors = FALSE
)

var_metadata$level <- sapply(var_metadata$industry_code, get_level)
```


\textbf{Extension 1: Level 2 only}

```{r, echo=FALSE}
# Keep only Level 2
keep_vars <- var_metadata$level == 2

X_raw <- as.matrix(data_wide[, c(FALSE, keep_vars)]) # drop date col
var_names_filtered <- var_names[keep_vars]
var_metadata_filtered <- var_metadata[keep_vars, ]

# Exclude KLD (aggregate index)
kld_vars <- grepl("_KLD$", var_names_filtered)

X_raw <- X_raw[, !kld_vars, drop = FALSE]
var_names_filtered <- var_names_filtered[!kld_vars]
var_metadata_filtered <- var_metadata_filtered[!kld_vars, ]

# Exclude Manufacture of Steel Tubes (C2420000), Animal Feed (C1090000), Basic Chemicals (C2010000), Fish Processing (C1020000) [results from "01_supplement_static_DFM"]

drop_industries <- c("C2420000", "C1090000", "C2010000", "C1020000")

pattern_drop <- paste0("^(", paste(drop_industries, collapse = "|"), ")_")
drop_vars <- grepl(pattern_drop, var_names_filtered)

cat("Dropping industries:",
    paste(drop_industries, collapse = ", "), "\n")
cat("Variables to drop:", sum(drop_vars), "\n")

if (sum(drop_vars) > 0) {
  drop_missing <- apply(X_raw[, drop_vars, drop = FALSE], 2,
                        function(x) mean(is.na(x)) * 100)
  cat("Avg missing rate (dropped vars):",
      round(mean(drop_missing), 2), "%\n")
}

X_raw <- X_raw[, !drop_vars, drop = FALSE]
var_names_filtered <- var_names_filtered[!drop_vars]
var_metadata_filtered <- var_metadata_filtered[!drop_vars, ]
```

```{r, include=FALSE}
# Linear Interpolation
X_filled <- apply(X_raw, 2, function(x) {
  if (all(is.na(x))) return(x) 
  zoo::na.approx(x, rule = 2) 
})

# Standardize using interpolated mean/sd (stable)
mu <- colMeans(X_filled, na.rm = TRUE)
sdv <- apply(X_filled, 2, sd, na.rm = TRUE)
sdv[sdv < 1e-10 | is.na(sdv)] <- 1

X_std <- sweep(X_filled, 2, mu, "-")
X_std <- sweep(X_std, 2, sdv, "/")

# Final finite check

valid <- apply(X_std, 2, function(x) all(is.finite(x)) && sd(x) > 1e-10)
X_std <- X_std[, valid, drop = FALSE]
X_raw <- X_raw[, valid, drop = FALSE]
X_filled <- X_filled[, valid, drop = FALSE]
var_names_filtered <- var_names_filtered[valid]
var_metadata_filtered <- var_metadata_filtered[valid, ]

N <- ncol(X_std)
T_obs <- nrow(X_std)

cat("=== Final Data ===\n")
cat("T =", T_obs, "months\n")
cat("N =", N, "variables\n")

# Standardized matrix with NA kept for Kalman
Y_kalman <- sweep(X_raw, 2, mu[valid], "-")
Y_kalman <- sweep(Y_kalman, 2, sdv[valid], "/")

```
The filtered dataset consists of 332 monthly observations (1997-07 to 2025-02) and 252 variables (18 Level-1 industries × 14 survey questions).

```{r, include=FALSE}
r <- 5  

pca_result <- prcomp(X_std, rank. = r)
F_hat <- pca_result$x
Lambda_hat <- pca_result$rotation

colnames(F_hat) <- paste0("F", 1:r)

cat("Factors:", nrow(F_hat), " x ", ncol(F_hat), "\n")
cat("Loadings:", nrow(Lambda_hat), " x ", ncol(Lambda_hat), "\n")

# sign normalization
expansion_period <- dates >= as.Date("2005-01-01") & dates <= as.Date("2008-12-31")

for (j in 1:r) {
mexp <- mean(F_hat[expansion_period, j], na.rm = TRUE)
if (mexp < 0) {
  F_hat[, j] <- -F_hat[, j]
  Lambda_hat[, j] <- -Lambda_hat[, j]
  cat("Factor", j, "flipped (aligned with expansion)\n")
  }
}
```


## VAR Lag Order Selection

```{r}
# fit VAR(p) on F_hat
max_p <- 6
ic_tbl <- purrr::map_dfr(1:max_p, function(p){
  fit <- vars::VAR(as.data.frame(F_hat), p = p, type = "const")
  tibble(p = p, AIC = AIC(fit), BIC = BIC(fit))
})

kbl(ic_tbl, digits = 3, caption = "VAR order selection (information criteria)") %>%
kable_styling(full_width = FALSE)

p_var <- ic_tbl$p[which.min(ic_tbl$BIC)]
cat("Selected VAR order: p =", p_var, "\n")
```

Based on the minimization of the Bayesian Information Criterion (BIC), a lag order of p=2 is selected for the VAR process. This ensures a stable balance between model fit and complexity.


## Parameter Estimation

```{r}
# 1. Fit the Vector Autoregression (VAR) on extracted factors
# F_t = c + A_1*F_{t-1} + A_2*F_{t-2} + ... + A_p*F_{t-p} + u_t
var_fit <- VAR(as.data.frame(F_hat), p = p_var, type = "const")

# 2. Robust extraction of VAR coefficient blocks using Bcoef()
# B = [c, A_1, A_2, ..., A_p] \in R^{r \times (1 + r*p)}
# Bcoef returns a concatenated matrix containing the intercept and all lag coefficient matrices
B <- Bcoef(var_fit)      # r x (1 + r*p)

# 3. Remove the intercept (constant) term
B_noC <- B[, colnames(B) != "const", drop = FALSE]

# 4. Reshape coefficients into a 3D array, separated by lag order
# Decomposition of B_{noC} into p distinct r x r matrices {A_1, A_2, ..., A_p}
A_array <- array(0, dim = c(r, r, p_var))
  for (lag in 1:p_var) {
  cols <- ((lag - 1) * r + 1):(lag * r)
  A_array[, , lag] <- B_noC[, cols, drop=FALSE]
}

# 5. Estimate the Factor Shock Covariance Matrix (Q) from residuals
# Q = E[u_t u_t'] \approx \frac{1}{T} \sum_{t=1}^T \hat{u}_t \hat{u}_t'
# Where \hat{u}_t represents the residuals from the VAR model
U <- resid(var_fit)
Q <- crossprod(as.matrix(U)) / nrow(U)

cat("VAR(p) fitted: r =", r, ", p =", p_var, "\n")
```

### Check VAR stability (all roots outside unit circle)
```{r}
# Check VAR stability (all roots outside unit circle)
roots <- roots(var_fit)
max(Mod(roots))
```
The VAR dynamics are stable, with the largest root less than one, indicating highly persistent but stationary factor processes.


```{r, include=FALSE}
library(ggplot2)
library(reshape2)

plot_A <- function(A, lag) {
  df <- melt(A)
  colnames(df) <- c("i", "j", "value")
  ggplot(df, aes(j, i, fill = value)) +
    geom_tile() +
    scale_fill_gradient2() +
    labs(
      title = paste("VAR Coefficients A", lag),
      x = expression(F[t - lag]),
      y = expression(F[t])
    ) +
    theme_minimal()
}

plot_A(A_array[,,1], 1)
plot_A(A_array[,,2], 2)


# Short-run factor dynamics are dominated by own-lag effects (A₁), while second-order lags (A₂) 
# mainly contribute to smoothing and minorcross-factor adjustments.
# This structure is well suited for Kalman smoothing and yields economically plausible, smoothly evolving dynamic factors.

```

### Calculate idiosyncratic residuals
```{r}
# Calculate the idiosyncratic residuals (error terms)
Xhat_pca <- F_hat %*% t(Lambda_hat)  # \hat{X}_t = \hat{\Lambda} \hat{F}_t; Dimensions: T x N
# Subtract the part explained by common factors from the standardized original data
E_hat <- X_std - Xhat_pca # \hat{\xi}_t = X_t - \hat{X}_t
psi_diag <- apply(E_hat, 2, var, na.rm = TRUE) #\hat{\Psi} = diag(Var(\hat{\xi}_{i,t}))

# Robustness Check: Handle missing values and near-zero variances
# If variance is zero or missing, set to a small floor (1e-8) to avoid division by zero in Kalman Filter
psi_diag[is.na(psi_diag) | psi_diag < 1e-8] <- 1e-8
summary(psi_diag) %>% print()

# Estimate the heteroskedastic idiosyncratic variance matrix Ψ by calculating the variance of the residuals 
# after removing the common components reconstructed via PCA. $\xi_t$ identifies the "noise" or variable-specific 
# movement that the common factors cannot explain.

```
The wide range of idiosyncratic variances demonstrates significant heterogeneity.


## Kalman smoother: build state-space model and smooth factors

```{r}
# Define augmented state dimension: m = r * p_var
m <- r * p_var

# Companion transition matrix Tmat (m x m)
Tmat <- matrix(0, m, m)

# Top block row: [A1 A2 ... Ap]
for (lag in 1:p_var) {
  Tmat[1:r, ((lag - 1) * r + 1):(lag * r)] <- A_array[, , lag]
}

# Sub-diagonal identity blocks
if (p_var > 1) {
  Tmat[(r + 1):m, 1:(m - r)] <- diag(m - r)
}

# Measurement matrix Z: N x m, only depends on current F_t (first r states)
Zmat <- matrix(0, N, m)
Zmat[, 1:r] <- Lambda_hat

# R maps factor shocks into the first r elements of the state
Rmat <- matrix(0, m, r)
Rmat[1:r, 1:r] <- diag(r)

# Observation noise H (Ψ) : N x N diagonal
Hmat <- diag(psi_diag, N, N)

a1 <- rep(0, m)
P1 <- diag(1e6, m)

# Kalman input: standardized data with NA kept
Y_kfas <- Y_kalman
colnames(Y_kfas) <- var_names_filtered
```



```{r}
kalman_smoother_missing <- function(Y, Z, Tmat, Rmat, Q, Hdiag, a1, P1) {
  # X: T x N (NA allowed)
  # Z: N x m
  # Tmat: m x m
  # Rmat: m x k
  # Q: k x k
  # Hdiag: length N (diagonal of H)
  # a1: length m
  # P1: m x m
  
  Y <- as.matrix(Y)
  storage.mode(Y) <- "double"
  
  T_obs <- nrow(Y)
  N <- ncol(Y)
  m <- ncol(Z)
  
  stopifnot(nrow(Z) == N, ncol(Tmat) == m, nrow(Tmat) == m)
  stopifnot(length(Hdiag) == N, length(a1) == m, all(dim(P1) == c(m, m)))
  
  # storage
  a_pred <- matrix(NA_real_, T_obs, m)
  a_filt <- matrix(NA_real_, T_obs, m)
  a_smooth <- matrix(NA_real_, T_obs, m)
  
  P_pred <- array(NA_real_, dim = c(m, m, T_obs))
  P_filt <- array(NA_real_, dim = c(m, m, T_obs))
  P_smooth <- array(NA_real_, dim = c(m, m, T_obs))
  
  # init
  a_prev <- as.numeric(a1)
  P_prev <- P1
  
  # filtering
  for (t in 1:T_obs) {
    # predict
    # a_t_pred = T * a_{t-1} 
    a_t_pred <- as.numeric(Tmat %*% a_prev) 
    # Var(a_t) P_t_pred = T * P_{t-1} * T' + R * Q * R' 
    P_t_pred <- Tmat %*% P_prev %*% t(Tmat) + Rmat %*% Q %*% t(Rmat) 
    
    a_pred[t, ] <- a_t_pred
    P_pred[, , t] <- P_t_pred
    
    obs <- which(!is.na(Y[t, ]))
    if (length(obs) == 0) {
      # no update
      a_t_filt <- a_t_pred
      P_t_filt <- P_t_pred
    } else {
      Zt <- Z[obs, , drop = FALSE]                  # n_t x m
      yt <- as.numeric(Y[t, obs])                   # n_t
      Ht <- diag(Hdiag[obs], nrow = length(obs))    # n_t x n_t
      
      v  <- yt - as.numeric(Zt %*% a_t_pred)        # innovation
      Fm <- Zt %*% P_t_pred %*% t(Zt) + Ht          # innovation cov
      
      # Numerical stabilization
      Fm <- (Fm + t(Fm)) / 2
      # Solve Fm * x = b
      K  <- P_t_pred %*% t(Zt) %*% solve(Fm)        # m x n_t
      
      a_t_filt <- as.numeric(a_t_pred + K %*% v)
      P_t_filt <- P_t_pred - K %*% Zt %*% P_t_pred
      P_t_filt <- (P_t_filt + t(P_t_filt)) / 2
    }
    
    a_filt[t, ] <- a_t_filt
    P_filt[, , t] <- P_t_filt
    
    a_prev <- a_t_filt
    P_prev <- P_t_filt
  }
  
  # Rauch–Tung–Striebel (RTS) smoother
  a_smooth[T_obs, ] <- a_filt[T_obs, ]
  P_smooth[, , T_obs] <- P_filt[, , T_obs]
  
  for (t in (T_obs - 1):1) {
    P_t_filt <- P_filt[, , t]
    P_tp1_pred <- P_pred[, , t + 1]
    
    # J = P_filt_t T' (P_pred_{t+1})^{-1}
    J <- P_t_filt %*% t(Tmat) %*% solve(P_tp1_pred)
    
    a_smooth[t, ] <- as.numeric(a_filt[t, ] + J %*% (a_smooth[t + 1, ] - a_pred[t + 1, ]))
    P_smooth[, , t] <- P_t_filt + J %*% (P_smooth[, , t + 1] - P_tp1_pred) %*% t(J)
    P_smooth[, , t] <- (P_smooth[, , t] + t(P_smooth[, , t])) / 2
  }
  
  list(
    a_pred = a_pred, a_filt = a_filt, a_smooth = a_smooth,
    P_pred = P_pred, P_filt = P_filt, P_smooth = P_smooth
  )
}

```

```{r}
# Ensure numeric matrix
Y_kfas <- as.matrix(Y_kalman)
storage.mode(Y_kfas) <- "double"
colnames(Y_kfas) <- var_names_filtered

Hdiag <- as.numeric(diag(Hmat))
Hdiag[!is.finite(Hdiag) | Hdiag < 1e-8] <- 1e-8

Q <- (Q + t(Q)) / 2
ev <- eigen(Q, symmetric = TRUE)
ev$values[ev$values < 1e-8] <- 1e-8
Q <- ev$vectors %*% diag(ev$values) %*% t(ev$vectors)

# Run smoother
ks <- kalman_smoother_missing(
  Y = Y_kfas,
  Z = Zmat,
  Tmat = Tmat,
  Rmat = Rmat,
  Q = Q,
  Hdiag = Hdiag,
  a1 = a1,
  P1 = P1
)

alpha_smooth <- ks$a_smooth   # T x m
F_smooth <- alpha_smooth[, 1:r, drop = FALSE]
colnames(F_smooth) <- paste0("F_smooth", 1:r)

```


```{r}
# Common component (standardized scale): X_common = Lambda * F_t
Lambda_hat <- Zmat[, 1:r, drop = FALSE]

X_common <- F_smooth %*% t(Lambda_hat)   # T x N
colnames(X_common) <- var_names_filtered

X_idio <- Y_kfas - X_common
colnames(X_idio) <- var_names_filtered
```

## Static PCA vs VAR Dynamics and Kalman Smoothing

The static PCA factors tend to exhibit noisy high-frequency fluctuations as shown in the figure below. These fluctuations often reflect measurement noise or temporary shocks rather than true economic trends.

In contrast, the dynamic factors are estimated under VAR(2) process within a state-space framework. The factors are no longer interpreted as isolated period-specific sentiment but as realizations of a continuously evolving latent economic state.


```{r, fig.width=10, fig.height=9.5, echo = FALSE}
# plot: Static PCA vs Dynamic (AR-smoothed via KF/RTS)

factor_titles <- c("Factor 1", "Factor 2", "Factor 3", "Factor 4", "Factor 5")
stopifnot(length(factor_titles) == r)


event_years  <- c(2008, 2011, 2020)
event_dates  <- as.Date(paste0(event_years, "-01-01"))
event_labels <- as.character(event_years)

# Keep only those inside sample
keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
event_dates  <- event_dates[keep_evt]
event_labels <- event_labels[keep_evt]

# High contrast
col_static  <- adjustcolor("gray20", alpha.f = 0.45)     # Static PCA
col_dynamic <- adjustcolor("dodgerblue3", alpha.f = 0.9) # Dynamic AR-smoothed
lwd_static  <- 1.8
lwd_dynamic <- 1.2
lty_static  <- 1            # solid
lty_dynamic <- 1            # dashed
evt_col     <- "red3"
evt_lty     <- 2            # dashed
evt_lwd     <- 1.2

# Layout
op <- par(no.readonly = TRUE)
on.exit(par(op), add = TRUE)

par(mfrow = c(r, 1),
    mar  = c(3.2, 4.2, 2.4, 1.2),
    oma  = c(3.0, 0.0, 3.0, 0.0))

for (j in 1:r) {
  y_rng <- range(F_hat[, j], F_smooth[, j], na.rm = TRUE)
  pad   <- 0.10 * diff(y_rng)
  if (!is.finite(pad) || pad == 0) pad <- 1
  y_lim <- c(y_rng[1] - pad, y_rng[2] + pad)

  plot(dates, F_hat[, j],
       type = "l",
       col  = col_static, lwd = lwd_static, lty = lty_static,
       ylim = y_lim,
       xlab = "", ylab = paste0("F", j),
       main = factor_titles[j])

  # Dynamic AR-smoothed
  lines(dates, F_smooth[, j],
        col = col_dynamic, lwd = lwd_dynamic, lty = lty_dynamic)

  abline(h = 0, col = "gray60", lty = 3)

  if (length(event_dates) > 0) {
    abline(v = event_dates, col = evt_col, lty = evt_lty, lwd = evt_lwd)
  }

  # label event years only on the first panel
  if (j == 1 && length(event_dates) > 0) {
    usr <- par("usr")   # x1 x2 y1 y2
    y_top <- usr[4]
    for (k in seq_along(event_dates)) {
      text(event_dates[k],
           y_top - 0.05 * diff(usr[3:4]),
           labels = event_labels[k],
           col = evt_col, cex = 0.85, pos = 4, xpd = NA)
    }
  }
}

# Outer title
mtext("Static PCA vs Dynamic DFM Factors", outer = TRUE, font = 2, cex = 1.2, line = 1.0)
mtext(
  "",
  outer = TRUE, cex = 0.95, line = -0.2
)


par(fig = c(0, 1, 0, 1), new = TRUE)
plot.new()
legend("bottom",
       inset = 0.02,
       legend = c("Static (PCA)", "Dynamic (AR-smoothed)"),
       col    = c(col_static, col_dynamic),
       lty    = c(lty_static, lty_dynamic),
       lwd    = c(lwd_static, lwd_dynamic),
       bty    = "n", horiz = TRUE, cex = 0.95)

```

The alignment between the static and dynamic estimates confirms that the cross-sectional factor structure extracted in the static step is well identified. The main medium-term movements are largely preserved when time-series dynamics are imposed, indicating that the static PCA already captures the dominant common variation in the data.

```{r}
# ---- Save outputs for Section 03 ----
dir.create("output", showWarnings = FALSE)

saveRDS(
  list(
    Y_kfas               = Y_kfas,
    F_smooth             = F_smooth,
    Hdiag                = Hdiag,
    r                    = r,
    dates                = dates,
    var_names_filtered   = var_names_filtered,
    var_metadata_filtered= var_metadata_filtered
  ),
  file = "output/dfm_results_02_level2.rds"
)
```



\textbf{Extension 2: Level 3 only}

```{r, echo=FALSE}
# Keep only Level 3
keep_vars <- var_metadata$level == 3

X_raw <- as.matrix(data_wide[, c(FALSE, keep_vars)]) # drop date col
var_names_filtered <- var_names[keep_vars]
var_metadata_filtered <- var_metadata[keep_vars, ]

# Exclude KLD (aggregate index)
kld_vars <- grepl("_KLD$", var_names_filtered)

X_raw <- X_raw[, !kld_vars, drop = FALSE]
var_names_filtered <- var_names_filtered[!kld_vars]
var_metadata_filtered <- var_metadata_filtered[!kld_vars, ]

# Exclude Manufacture of dyes and pigments (C2012000), other basic chemiclas (C2014000), plastics in primary forms (C2016000)
drop_industries <- c("C2012000", "C2014000", "C2016000")

pattern_drop <- paste0("^(", paste(drop_industries, collapse = "|"), ")_")
drop_vars <- grepl(pattern_drop, var_names_filtered)

cat("Dropping industries:",
    paste(drop_industries, collapse = ", "), "\n")
cat("Variables to drop:", sum(drop_vars), "\n")

if (sum(drop_vars) > 0) {
  drop_missing <- apply(X_raw[, drop_vars, drop = FALSE], 2,
                        function(x) mean(is.na(x)) * 100)
  cat("Avg missing rate (dropped vars):",
      round(mean(drop_missing), 2), "%\n")
}

X_raw <- X_raw[, !drop_vars, drop = FALSE]
var_names_filtered <- var_names_filtered[!drop_vars]
var_metadata_filtered <- var_metadata_filtered[!drop_vars, ]
```
```{r, include=FALSE}
# Linear Interpolation
X_filled <- apply(X_raw, 2, function(x) {
  if (all(is.na(x))) return(x) 
  zoo::na.approx(x, rule = 2) 
})

# Standardize using interpolated mean/sd (stable)
mu <- colMeans(X_filled, na.rm = TRUE)
sdv <- apply(X_filled, 2, sd, na.rm = TRUE)
sdv[sdv < 1e-10 | is.na(sdv)] <- 1

X_std <- sweep(X_filled, 2, mu, "-")
X_std <- sweep(X_std, 2, sdv, "/")

# Final finite check

valid <- apply(X_std, 2, function(x) all(is.finite(x)) && sd(x) > 1e-10)
X_std <- X_std[, valid, drop = FALSE]
X_raw <- X_raw[, valid, drop = FALSE]
X_filled <- X_filled[, valid, drop = FALSE]
var_names_filtered <- var_names_filtered[valid]
var_metadata_filtered <- var_metadata_filtered[valid, ]

N <- ncol(X_std)
T_obs <- nrow(X_std)

cat("=== Final Data ===\n")
cat("T =", T_obs, "months\n")
cat("N =", N, "variables\n")

# Standardized matrix with NA kept for Kalman
Y_kalman <- sweep(X_raw, 2, mu[valid], "-")
Y_kalman <- sweep(Y_kalman, 2, sdv[valid], "/")

```


The filtered dataset consists of 332 monthly observations (1997-07 to 2025-02) and 266 variables (19 Level-1 industries × 14 survey questions).

```{r, include=FALSE}
r <- 5  

pca_result <- prcomp(X_std, rank. = r)
F_hat <- pca_result$x
Lambda_hat <- pca_result$rotation

colnames(F_hat) <- paste0("F", 1:r)

cat("Factors:", nrow(F_hat), " x ", ncol(F_hat), "\n")
cat("Loadings:", nrow(Lambda_hat), " x ", ncol(Lambda_hat), "\n")

# sign normalization
expansion_period <- dates >= as.Date("2005-01-01") & dates <= as.Date("2008-12-31")

for (j in 1:r) {
mexp <- mean(F_hat[expansion_period, j], na.rm = TRUE)
if (mexp < 0) {
  F_hat[, j] <- -F_hat[, j]
  Lambda_hat[, j] <- -Lambda_hat[, j]
  cat("Factor", j, "flipped (aligned with expansion)\n")
  }
}
```


## VAR Lag Order Selection

```{r}
# fit VAR(p) on F_hat
max_p <- 6
ic_tbl <- purrr::map_dfr(1:max_p, function(p){
  fit <- vars::VAR(as.data.frame(F_hat), p = p, type = "const")
  tibble(p = p, AIC = AIC(fit), BIC = BIC(fit))
})

kbl(ic_tbl, digits = 3, caption = "VAR order selection (information criteria)") %>%
kable_styling(full_width = FALSE)

p_var <- ic_tbl$p[which.min(ic_tbl$BIC)]
cat("Selected VAR order: p =", p_var, "\n")
```

Based on the minimization of the Bayesian Information Criterion (BIC), a lag order of p=2 is selected for the VAR process. This ensures a stable balance between model fit and complexity.


## Parameter Estimation

```{r}
# 1. Fit the Vector Autoregression (VAR) on extracted factors
# F_t = c + A_1*F_{t-1} + A_2*F_{t-2} + ... + A_p*F_{t-p} + u_t
var_fit <- VAR(as.data.frame(F_hat), p = p_var, type = "const")

# 2. Robust extraction of VAR coefficient blocks using Bcoef()
# B = [c, A_1, A_2, ..., A_p] \in R^{r \times (1 + r*p)}
# Bcoef returns a concatenated matrix containing the intercept and all lag coefficient matrices
B <- Bcoef(var_fit)      # r x (1 + r*p)

# 3. Remove the intercept (constant) term
B_noC <- B[, colnames(B) != "const", drop = FALSE]

# 4. Reshape coefficients into a 3D array, separated by lag order
# Decomposition of B_{noC} into p distinct r x r matrices {A_1, A_2, ..., A_p}
A_array <- array(0, dim = c(r, r, p_var))
  for (lag in 1:p_var) {
  cols <- ((lag - 1) * r + 1):(lag * r)
  A_array[, , lag] <- B_noC[, cols, drop=FALSE]
}

# 5. Estimate the Factor Shock Covariance Matrix (Q) from residuals
# Q = E[u_t u_t'] \approx \frac{1}{T} \sum_{t=1}^T \hat{u}_t \hat{u}_t'
# Where \hat{u}_t represents the residuals from the VAR model
U <- resid(var_fit)
Q <- crossprod(as.matrix(U)) / nrow(U)

cat("VAR(p) fitted: r =", r, ", p =", p_var, "\n")
```

### Check VAR stability (all roots outside unit circle)
```{r}
# Check VAR stability (all roots outside unit circle)
roots <- roots(var_fit)
max(Mod(roots))
```
The VAR dynamics are stable, with the largest root less than one, indicating highly persistent but stationary factor processes.


```{r, include=FALSE}
library(ggplot2)
library(reshape2)

plot_A <- function(A, lag) {
  df <- melt(A)
  colnames(df) <- c("i", "j", "value")
  ggplot(df, aes(j, i, fill = value)) +
    geom_tile() +
    scale_fill_gradient2() +
    labs(
      title = paste("VAR Coefficients A", lag),
      x = expression(F[t - lag]),
      y = expression(F[t])
    ) +
    theme_minimal()
}

plot_A(A_array[,,1], 1)
plot_A(A_array[,,2], 2)


# Short-run factor dynamics are dominated by own-lag effects (A₁), while second-order lags (A₂) 
# mainly contribute to smoothing and minorcross-factor adjustments.
# This structure is well suited for Kalman smoothing and yields economically plausible, smoothly evolving dynamic factors.

```

### Calculate idiosyncratic residuals
```{r}
# Calculate the idiosyncratic residuals (error terms)
Xhat_pca <- F_hat %*% t(Lambda_hat)  # \hat{X}_t = \hat{\Lambda} \hat{F}_t; Dimensions: T x N
# Subtract the part explained by common factors from the standardized original data
E_hat <- X_std - Xhat_pca # \hat{\xi}_t = X_t - \hat{X}_t
psi_diag <- apply(E_hat, 2, var, na.rm = TRUE) #\hat{\Psi} = diag(Var(\hat{\xi}_{i,t}))

# Robustness Check: Handle missing values and near-zero variances
# If variance is zero or missing, set to a small floor (1e-8) to avoid division by zero in Kalman Filter
psi_diag[is.na(psi_diag) | psi_diag < 1e-8] <- 1e-8
summary(psi_diag) %>% print()

# Estimate the heteroskedastic idiosyncratic variance matrix Ψ by calculating the variance of the residuals 
# after removing the common components reconstructed via PCA. $\xi_t$ identifies the "noise" or variable-specific 
# movement that the common factors cannot explain.

```

The wide range of idiosyncratic variances demonstrates significant heterogeneity.


## Kalman smoother: build state-space model and smooth factors

```{r}
# Define augmented state dimension: m = r * p_var
m <- r * p_var

# Companion transition matrix Tmat (m x m)
Tmat <- matrix(0, m, m)

# Top block row: [A1 A2 ... Ap]
for (lag in 1:p_var) {
  Tmat[1:r, ((lag - 1) * r + 1):(lag * r)] <- A_array[, , lag]
}

# Sub-diagonal identity blocks
if (p_var > 1) {
  Tmat[(r + 1):m, 1:(m - r)] <- diag(m - r)
}

# Measurement matrix Z: N x m, only depends on current F_t (first r states)
Zmat <- matrix(0, N, m)
Zmat[, 1:r] <- Lambda_hat

# R maps factor shocks into the first r elements of the state
Rmat <- matrix(0, m, r)
Rmat[1:r, 1:r] <- diag(r)

# Observation noise H (Ψ) : N x N diagonal
Hmat <- diag(psi_diag, N, N)

a1 <- rep(0, m)
P1 <- diag(1e6, m)

# Kalman input: standardized data with NA kept
Y_kfas <- Y_kalman
colnames(Y_kfas) <- var_names_filtered
```



```{r}
kalman_smoother_missing <- function(Y, Z, Tmat, Rmat, Q, Hdiag, a1, P1) {
  # X: T x N (NA allowed)
  # Z: N x m
  # Tmat: m x m
  # Rmat: m x k
  # Q: k x k
  # Hdiag: length N (diagonal of H)
  # a1: length m
  # P1: m x m
  
  Y <- as.matrix(Y)
  storage.mode(Y) <- "double"
  
  T_obs <- nrow(Y)
  N <- ncol(Y)
  m <- ncol(Z)
  
  stopifnot(nrow(Z) == N, ncol(Tmat) == m, nrow(Tmat) == m)
  stopifnot(length(Hdiag) == N, length(a1) == m, all(dim(P1) == c(m, m)))
  
  # storage
  a_pred <- matrix(NA_real_, T_obs, m)
  a_filt <- matrix(NA_real_, T_obs, m)
  a_smooth <- matrix(NA_real_, T_obs, m)
  
  P_pred <- array(NA_real_, dim = c(m, m, T_obs))
  P_filt <- array(NA_real_, dim = c(m, m, T_obs))
  P_smooth <- array(NA_real_, dim = c(m, m, T_obs))
  
  # init
  a_prev <- as.numeric(a1)
  P_prev <- P1
  
  # filtering
  for (t in 1:T_obs) {
    # predict
    # a_t_pred = T * a_{t-1} 
    a_t_pred <- as.numeric(Tmat %*% a_prev) 
    # Var(a_t) P_t_pred = T * P_{t-1} * T' + R * Q * R' 
    P_t_pred <- Tmat %*% P_prev %*% t(Tmat) + Rmat %*% Q %*% t(Rmat) 
    
    a_pred[t, ] <- a_t_pred
    P_pred[, , t] <- P_t_pred
    
    obs <- which(!is.na(Y[t, ]))
    if (length(obs) == 0) {
      # no update
      a_t_filt <- a_t_pred
      P_t_filt <- P_t_pred
    } else {
      Zt <- Z[obs, , drop = FALSE]                  # n_t x m
      yt <- as.numeric(Y[t, obs])                   # n_t
      Ht <- diag(Hdiag[obs], nrow = length(obs))    # n_t x n_t
      
      v  <- yt - as.numeric(Zt %*% a_t_pred)        # innovation
      Fm <- Zt %*% P_t_pred %*% t(Zt) + Ht          # innovation cov
      
      # Numerical stabilization
      Fm <- (Fm + t(Fm)) / 2
      # Solve Fm * x = b
      K  <- P_t_pred %*% t(Zt) %*% solve(Fm)        # m x n_t
      
      a_t_filt <- as.numeric(a_t_pred + K %*% v)
      P_t_filt <- P_t_pred - K %*% Zt %*% P_t_pred
      P_t_filt <- (P_t_filt + t(P_t_filt)) / 2
    }
    
    a_filt[t, ] <- a_t_filt
    P_filt[, , t] <- P_t_filt
    
    a_prev <- a_t_filt
    P_prev <- P_t_filt
  }
  
  # Rauch–Tung–Striebel (RTS) smoother
  a_smooth[T_obs, ] <- a_filt[T_obs, ]
  P_smooth[, , T_obs] <- P_filt[, , T_obs]
  
  for (t in (T_obs - 1):1) {
    P_t_filt <- P_filt[, , t]
    P_tp1_pred <- P_pred[, , t + 1]
    
    # J = P_filt_t T' (P_pred_{t+1})^{-1}
    J <- P_t_filt %*% t(Tmat) %*% solve(P_tp1_pred)
    
    a_smooth[t, ] <- as.numeric(a_filt[t, ] + J %*% (a_smooth[t + 1, ] - a_pred[t + 1, ]))
    P_smooth[, , t] <- P_t_filt + J %*% (P_smooth[, , t + 1] - P_tp1_pred) %*% t(J)
    P_smooth[, , t] <- (P_smooth[, , t] + t(P_smooth[, , t])) / 2
  }
  
  list(
    a_pred = a_pred, a_filt = a_filt, a_smooth = a_smooth,
    P_pred = P_pred, P_filt = P_filt, P_smooth = P_smooth
  )
}

```

```{r}
# Ensure numeric matrix
Y_kfas <- as.matrix(Y_kalman)
storage.mode(Y_kfas) <- "double"
colnames(Y_kfas) <- var_names_filtered

Hdiag <- as.numeric(diag(Hmat))
Hdiag[!is.finite(Hdiag) | Hdiag < 1e-8] <- 1e-8

Q <- (Q + t(Q)) / 2
ev <- eigen(Q, symmetric = TRUE)
ev$values[ev$values < 1e-8] <- 1e-8
Q <- ev$vectors %*% diag(ev$values) %*% t(ev$vectors)

# Run smoother
ks <- kalman_smoother_missing(
  Y = Y_kfas,
  Z = Zmat,
  Tmat = Tmat,
  Rmat = Rmat,
  Q = Q,
  Hdiag = Hdiag,
  a1 = a1,
  P1 = P1
)

alpha_smooth <- ks$a_smooth   # T x m
F_smooth <- alpha_smooth[, 1:r, drop = FALSE]
colnames(F_smooth) <- paste0("F_smooth", 1:r)

```


```{r}
# Common component (standardized scale): X_common = Lambda * F_t
Lambda_hat <- Zmat[, 1:r, drop = FALSE]

X_common <- F_smooth %*% t(Lambda_hat)   # T x N
colnames(X_common) <- var_names_filtered

X_idio <- Y_kfas - X_common
colnames(X_idio) <- var_names_filtered
```

## Static PCA vs VAR Dynamics and Kalman Smoothing

```{r, fig.width=10, fig.height=9.5, echo = FALSE}
# plot: Static PCA vs Dynamic (AR-smoothed via KF/RTS)

factor_titles <- c("Factor 1", "Factor 2", "Factor 3", "Factor 4", "Factor 5")
stopifnot(length(factor_titles) == r)


event_years  <- c(2008, 2011, 2020)
event_dates  <- as.Date(paste0(event_years, "-01-01"))
event_labels <- as.character(event_years)

# Keep only those inside sample
keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
event_dates  <- event_dates[keep_evt]
event_labels <- event_labels[keep_evt]

# High contrast
col_static  <- adjustcolor("gray20", alpha.f = 0.45)     # Static PCA
col_dynamic <- adjustcolor("dodgerblue3", alpha.f = 0.9) # Dynamic AR-smoothed
lwd_static  <- 1.8
lwd_dynamic <- 1.2
lty_static  <- 1            # solid
lty_dynamic <- 1            # dashed
evt_col     <- "red3"
evt_lty     <- 2            # dashed
evt_lwd     <- 1.2

# Layout
op <- par(no.readonly = TRUE)
on.exit(par(op), add = TRUE)

par(mfrow = c(r, 1),
    mar  = c(3.2, 4.2, 2.4, 1.2),
    oma  = c(3.0, 0.0, 3.0, 0.0))

for (j in 1:r) {
  y_rng <- range(F_hat[, j], F_smooth[, j], na.rm = TRUE)
  pad   <- 0.10 * diff(y_rng)
  if (!is.finite(pad) || pad == 0) pad <- 1
  y_lim <- c(y_rng[1] - pad, y_rng[2] + pad)

  plot(dates, F_hat[, j],
       type = "l",
       col  = col_static, lwd = lwd_static, lty = lty_static,
       ylim = y_lim,
       xlab = "", ylab = paste0("F", j),
       main = factor_titles[j])

  # Dynamic AR-smoothed
  lines(dates, F_smooth[, j],
        col = col_dynamic, lwd = lwd_dynamic, lty = lty_dynamic)

  abline(h = 0, col = "gray60", lty = 3)

  if (length(event_dates) > 0) {
    abline(v = event_dates, col = evt_col, lty = evt_lty, lwd = evt_lwd)
  }

  # label event years only on the first panel
  if (j == 1 && length(event_dates) > 0) {
    usr <- par("usr")   # x1 x2 y1 y2
    y_top <- usr[4]
    for (k in seq_along(event_dates)) {
      text(event_dates[k],
           y_top - 0.05 * diff(usr[3:4]),
           labels = event_labels[k],
           col = evt_col, cex = 0.85, pos = 4, xpd = NA)
    }
  }
}

# Outer title
mtext("Static PCA vs Dynamic DFM Factors", outer = TRUE, font = 2, cex = 1.2, line = 1.0)
mtext(
  "",
  outer = TRUE, cex = 0.95, line = -0.2
)


par(fig = c(0, 1, 0, 1), new = TRUE)
plot.new()
legend("bottom",
       inset = 0.02,
       legend = c("Static (PCA)", "Dynamic (AR-smoothed)"),
       col    = c(col_static, col_dynamic),
       lty    = c(lty_static, lty_dynamic),
       lwd    = c(lwd_static, lwd_dynamic),
       bty    = "n", horiz = TRUE, cex = 0.95)

```

The alignment between the static and dynamic estimates confirms that the cross-sectional factor structure extracted in the static step is well identified. The main medium-term movements are largely preserved when time-series dynamics are imposed, indicating that the static PCA already captures the dominant common variation in the data.

In contrast to the dominant manufacturing cycle (Factor 1), which remains almost unchanged between the static and dynamic specifications due to its strong and persistent cross-sectional signal, Factor 5 shows a more noticeable deviation after Kalman smoothing. This factor captures comparatively heterogeneous and weaker sector-specific adjustments across industries such as food processing, construction materials, consumer durables, and selected metal segments. Because its loadings are more dispersed and its signal-to-noise ratio is lower, the static PCA tends to absorb short-lived survey fluctuations into this component. The dynamic model, by imposing a VAR structure on the latent factors, filters out these transitory movements unless they are supported by persistent dynamics. The resulting divergence therefore reflects the weaker structural coherence of Factor 5 rather than a change in its economic interpretation.

```{r}
# ---- Save outputs for Section 03 ----
dir.create("output", showWarnings = FALSE)

saveRDS(
  list(
    Y_kfas               = Y_kfas,
    F_smooth             = F_smooth,
    Hdiag                = Hdiag,
    r                    = r,
    dates                = dates,
    var_names_filtered   = var_names_filtered,
    var_metadata_filtered= var_metadata_filtered
  ),
  file = "output/dfm_results_02_level3.rds"
)
```