---
title: "03.1 Time-Varying Factor Loadings"
header-includes:
  - \usepackage{parskip}
  - \setlength{\parskip}{0.5em}
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5)
```

```{r}
dfm02 <- readRDS("output/dfm_results_02.rds")
source("ifo_data_processing.R")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered

```



While the dynamic factor model allows the latent factors to evolve over time, it typically maintains a time-invariant measurement relationship between the observed variables and the latent factors. However, the economic relevance of individual indicators may change gradually over time. Such shifts can arise from structural transformation of the economy, institutional reforms, or persistent changes in firms’ expectations.

To allow for **gradual structural change** in the measurement relationship, we extend the measurement equation by letting factor loadings vary over time. To maintain tractability, we first impose a parsimonious restriction that all variables within the same Level-1 industry group share the common (time-varying) factor loading vector.

Let \(g(i)\in\{1,\dots,G\}\) denote the Level-1 group membership of variable \(i\).
The measurement equation becomes

\[
X_{i,t} = \lambda_{g(i),t}^\top F_t + \xi_{i,t},
\qquad \xi_{i,t}\sim \mathcal N(0, H_i),
\]

where \(F_t\in\mathbb R^r\) is the dynamic factor (smoothed estimate from Section 02), and \(\lambda_{g,t}\in\mathbb R^r\) is the **group-level loading** shared by all variables in group \(g\).

We model gradual structural change via a random-walk evolution:

\[
\lambda_{g,t} = \lambda_{g,t-1} + u_{g,t},
\qquad u_{g,t}\sim \mathcal N(0, q_g I_r).
\]

Stacking the \(n_g\) observed series in group \(g\) yields a multivariate observation equation:

\[
X_{g,t} = \mathbf 1_{n_g}\lambda_{g,t}^\top F_t + \xi_{g,t},
\quad \xi_{g,t}\sim \mathcal N(0, H_g), \quad H_g=\mathrm{diag}(H_i: i\in g)
\]

The stacked observation vector $X_{g,t}$ represents all survey responses within industry $g$ at time $t$. Under the group-level loading structure, these responses share a common latent signal driven by the dynamic factors, while differences across questions are attributed to idiosyncratic measurement noise.


```{r}
library(KFAS)
library(tidyverse)

X_kfas <- Y_kfas
F_use <- as.matrix(F_smooth) # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(X_kfas)
stopifnot(ncol(F_use) == r, nrow(X_kfas) == T_obs, length(Hdiag) == N_obs)
```



```{r, include=FALSE}
# In 02 already kept only Level 1: keep_vars <- var_metadata$level == 1

stopifnot("industry_code" %in% colnames(var_metadata_filtered))
stopifnot("level" %in% colnames(var_metadata_filtered))

group_id <- as.character(var_metadata_filtered$industry_code)
groups <- sort(unique(group_id))
G <- length(groups)

cat("Number of Level-1 groups (industry_code):", G, "\n")

# Group sizes
group_sizes <- as.integer(table(group_id)[groups])
group_table <- data.frame(industry_code = groups, n_g = group_sizes) |>
arrange(desc(n_g))
head(group_table, 10)
```

```{r}
# Time-Varying Parameter (TVP) Factor Loading Estimation using KFAS
## Model Specification:
##   Measurement: X_{g,t} = Z_{g,t} * lambda_{g,t} + xi_{g,t}
##   State:       lambda_{g,t} = lambda_{g,t-1} + eta_{g,t}
##   Innovation:  eta_{g,t} ~ N(0, Q), where Q = q_g * I_r

fit_tvp_group <- function(X_group, Fmat, H_vec, q_init = 1e-4) {
  
  T_obs <- nrow(Fmat) 
  r     <- ncol(Fmat)     
  n_g   <- ncol(X_group)   
  
  
  #  Build time-varying measurement matrix Z_{g,t}
  ## For each time t, the measurement equation is:
  ##   X_{i,t} = F_{1,t} * lambda_{i,1,t} + ... + F_{r,t} * lambda_{i,r,t} + xi_{i,t}
  ##
  ## In matrix form for the entire group:
  ##   X_{g,t} = Z_{g,t} * lambda_{g,t} + xi_{g,t}
  ##
  ## where Z_{g,t} is an (n_g x r) matrix with structure:
  ##   Z_{g,t} = 1_{n_g} ⊗ F_t'  (each row is F_t')
  ##
  ##   Z_arr: (n_g, r, T) array
  
  Z_arr <- array(NA_real_, dim = c(n_g, r, T_obs))
  
  for (t in 1:T_obs) {
    Z_arr[, , t] <- matrix(
      rep(Fmat[t, ], each = n_g),  ## For each time t, create an (n_g x r) matrix
      nrow  = n_g,                 ## where each row is F_t' (the factor values at time t)
      byrow = FALSE
    )
  }
  
  # Specify State-Space Model using SSMcustom
  ## State-Space representation:
  ## MEASUREMENT EQUATION: y_t = Z_t * alpha_t + epsilon_t,  epsilon_t ~ N(0, H)
  ## STATE EQUATION: alpha_t = T * alpha_{t-1} + R * eta_t,  eta_t ~ N(0, Q)
  ##
  ## In our TVP loading model:
  ##   - y_t = X_{g,t}        (n_g x 1)
  ##   - alpha_t = lambda_{g,t}  (r x 1, the time-varying loadings)
  ##   - Z_t = Z_{g,t}        (n_g x r, time-varying)
  ##   - T = I_r              (r x r identity, random walk)
  ##   - R = I_r              (r x r identity, full-rank innovations)
  ##   - Q = q_g * I_r        (r x r, scalar variance * identity)
  ##   - H = diag(H_vec)      (n_g x n_g, fixed, diagonal)
  
  model0 <- SSModel(
    X_group ~ -1 +  ## Left-hand side: observed data (no intercept)
      
      SSMcustom(
        Z  = Z_arr,                    ## Time-varying measurement matrix (n_g, r, T)
        T  = diag(r),                  ## State transition: I_r (random walk)
        R  = diag(r),                  ## Selection matrix: I_r (full innovations)
        Q  = diag(NA_real_, r),        ## Innovation variance: q * I_r (to be estimated)
        a1 = rep(0, r),                ## Initial state mean: 0
        P1 = diag(10, r)               ## Initial state variance: diffuse prior
      ),
    
    H = diag(H_vec, n_g)               ## Observation error variance: fixed diagonal
  )
  
  
  # Define update function for Q matrix
  ## The KFAS package requires us to provide an "update function" that
  ## takes optimization parameters and updates the model accordingly.
  ##
  ## We estimate a SINGLE scalar parameter q_g > 0, and set:
  ##   Q = q_g * I_r
  ## To ensure q_g > 0, we optimize log(q_g) and then exponentiate.
  
  updatefn <- function(pars, model) {
    q <- exp(pars[1])            ## Exponentiate to ensure positivity
    model$Q[, , 1] <- diag(q, r) ## Update Q matrix: q * I_r
    return(model)
  }
  
  
  # Optimize the log-likelihood to estimate q_g
  fit <- fitSSM(
    model  = model0,
    inits  = log(q_init),      ## Starting value: log(q_init)
    updatefn = updatefn,       ## Function to update Q given parameters
    method = "BFGS"            ## Optimization algorithm
  )
  
  model <- fit$model
  q_hat <- exp(fit$optim.out$par[1])  ## Estimated innovation variance
  
  
  # Kalman Smoothing - Smoothed states: alpha_{t|T} (uses full-sample information)
  kfs <- KFS(
    model = model,
    smoothing = c("state")  ## Request smoothed state estimates
  )
  
  ## Extract smoothed state estimates: lambda_{g,t|T}
  lambda_t <- kfs$alphahat  ## T x r matrix
  colnames(lambda_t) <- paste0("lambda_", 1:r)
  
  return(list(
    lambda_t = lambda_t,  ## Smoothed time-varying loadings (T x r)
    q_hat    = q_hat      ## Estimated innovation variance (scalar)
  ))
  
}
```

```{r}
## Estimation of Time-Varying Loadings for All Industry Groups

## Storage for time-varying loadings
## Each element will be a T x r matrix for one industry group
Lambda_g_tv <- vector("list", G)
names(Lambda_g_tv) <- groups

## Storage for estimated innovation variances
## One scalar q_g per group
q_hat_g <- rep(NA_real_, G)
names(q_hat_g) <- groups


for (gg in seq_along(groups)) {
  
  g <- groups[gg]                ## Group name (e.g., "C2800000")
  idx <- which(group_id == g)    ## Column indices for this group in X_kfas
  n_g <- length(idx)             ## Number of series in this group
  
  # Extract observed data for this group: T x n_g matrix
  X_group <- as.matrix(X_kfas[, idx, drop = FALSE])
  
  # Extract fixed observation error variances for this group
  H_vec <- Hdiag[idx]

  # Estimate time-varying loadings using KFAS
  res <- fit_tvp_group(
    X_group = X_group,     ## T x n_g observed data
    Fmat    = F_use,       ## T x r static factors
    H_vec   = H_vec,       ## n_g observation error variances
    q_init  = 1e-4         ## Initial value for innovation variance
  )

  Lambda_g_tv[[g]] <- res$lambda_t
  q_hat_g[g] <- res$q_hat
}  

```


```{r}
## Expand Group-Level TV Loadings to Variable Level - maps group-level loading 
## paths to individual variables. Each variable inherits the loading path of its parent group.
##
##   - Input:  Lambda_g_tv (list of G elements, each T x r)
##   - Output: Lambda_tv_list (list of N elements, each T x r)
##
## Mapping:  lambda_{i,t} = lambda_{g(i),t}
##           where g(i) is the group that variable i belongs to

N_obs <- length(var_names_filtered)

Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

# Map group-level loadings to individual variables
for (i in 1:N_obs) {
  g <- group_id[i]
  Lambda_tv_list[[i]] <- Lambda_g_tv[[g]]  # T x r 
}
```


```{r, include=FALSE}
# Instability Measures for Each Group
instability_by_group <- lapply(groups, function(g) {
  
  L <- Lambda_g_tv[[g]]  # T x r
  
  dL <- apply(L, 2, diff)
  
  data.frame(
    industry_code       = g,
    n_g                 = sum(group_id == g),
    q_hat               = q_hat_g[g],
    mean_abs_dlambda    = mean(abs(dL), na.rm = TRUE),
    sd_dlambda          = sd(as.numeric(dL), na.rm = TRUE)
  )
  
}) |>
  bind_rows() |>
  arrange(desc(mean_abs_dlambda))

head(instability_by_group, 15)
```


```{r}
plot_factor_all_industries <- function(
  Lambda_g_tv,
  dates,
  factor_j = 1,
  industries = NULL,
  event_years = c(2008, 2011, 2020),
  title_prefix = "Industry-only time-varying loadings",
  show_legend = TRUE,
  fade_highvol = TRUE,  
  top_k = 5
) {
  gs <- names(Lambda_g_tv)
  if (!is.null(industries)) gs <- intersect(gs, industries)
  stopifnot(length(gs) > 0)

  # compute volatility per industry (for this factor)
  vol_df <- data.frame(
    industry_code = gs,
    vol = sapply(gs, function(g) {
      L <- Lambda_g_tv[[g]][, factor_j]
      stats::sd(diff(L), na.rm = TRUE)
    }),
    stringsAsFactors = FALSE
  )

  highvol_industries <- vol_df |>
    dplyr::arrange(dplyr::desc(vol)) |>
    dplyr::slice_head(n = top_k) |>
    dplyr::pull(industry_code)

  # build long data
  df_long <- do.call(rbind, lapply(gs, function(g) {
    L <- Lambda_g_tv[[g]]  # T x r
    stopifnot(nrow(L) == length(dates))
    data.frame(
      date = dates,
      industry_code = g,
      industry_name = get_industry_name_en(g),
      loading = as.numeric(L[, factor_j]),
      is_highvol = g %in% highvol_industries,  
      stringsAsFactors = FALSE
    )
  }))

  # events
  event_dates  <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates  <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  # base plot: keep global color mapping so ALL are colored
  p <- ggplot2::ggplot(df_long, ggplot2::aes(x = date, y = loading, color = industry_name)) +
    ggplot2::geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    ggplot2::scale_color_viridis_d(option = "turbo") +
    ggplot2::labs(
      title = paste0(title_prefix, " — Factor ", factor_j),
      subtitle = if (fade_highvol)
        paste0("High-volatility industries (top ", top_k, ") are faded")
      else
        "Each line corresponds to one Level-1 industry",
      x = NULL,
      y = paste0("Loading (lambda_", factor_j, ")"),
      color = "Industry"
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(
      plot.title.position = "plot",
      legend.position = if (show_legend) "bottom" else "none",
      legend.key.height = grid::unit(0.45, "lines"),
      legend.text = ggplot2::element_text(size = 9),
      plot.margin = grid::unit(c(6, 6, 6, 6), "pt")
    )

  # draw lines in two layers to control emphasis
  if (fade_highvol) {
    # 1) draw HIGH-vol first (faded, but still colored)
    p <- p + ggplot2::geom_line(
      data = df_long[df_long$is_highvol, ],
      linewidth = 0.7,
      alpha = 0.25
    )
    # 2) draw LOW-vol on top (more visible)
    p <- p + ggplot2::geom_line(
      data = df_long[!df_long$is_highvol, ],
      linewidth = 1.0,
      alpha = 0.95
    )
  } else {
    # no fading: all same style
    p <- p + ggplot2::geom_line(linewidth = 0.9, alpha = 0.85)
  }

  if (show_legend) {
    p <- p + ggplot2::guides(color = ggplot2::guide_legend(ncol = 3))
  }

  if (length(event_dates) > 0) {
    evt_df <- data.frame(date = event_dates, label = event_labels)
    p <- p +
      ggplot2::geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.8) +
      ggplot2::geom_text(
        data = evt_df,
        ggplot2::aes(x = date, y = Inf, label = label),
        inherit.aes = FALSE,
        vjust = 1.2, hjust = 0,
        color = "red3",
        size = 3
      ) +
      ggplot2::coord_cartesian(clip = "off")
  }

  p
}
```


```{r fig.width=14, fig.height=6, echo=FALSE, include=FALSE}
#for (j in 1:r) {
#  print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = j, show_legend = (j == 5)))
#}

for (j in 1:r) {
  print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = j, fade_highvol = TRUE, top_k = 5))
}
```




```{r}
print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = 1, fade_highvol = TRUE, top_k = 5))
```

**Factor 1: Capital goods Manufacturing cycle**

Factor 1 captures the largest share of common variation in the data, reflecting the investment-driven core of German manufacturing activity. It summarizes the synchronized expansion and contraction of key capital goods industries and serves as a broad indicator of underlying macroeconomic conditions in the manufactuaring sector.

  Core capital-goods industries: Machinery, Electrical Equipment, Fabricated Metal Products, these three industries exhibit persistently positive and relatively high loadings throughout the sample. This tight clustering of loadings indicates strong cross-sectional co-movement, confirming that Factor 1 remains firmly anchored in investment-related manufacturing dynamics.
  
  Rubber and Plastics show a marked increase after 2011, likely reflecting their integration into machinery and automotive supply chains. 
  The relatively high volatility of Factor-1 loadings reflects the broad and heterogeneous nature of the investment cycle. As this factor aggregates multiple stages of the production chain, shifts in investment composition and crisis-driven disruptions tend to reallocate exposure across industries, resulting in visible cross-sectional and time-series variation in loadings.

\vspace{0.5cm}

```{r, fig.height = 5}
print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = 2, fade_highvol = TRUE, top_k = 5))
```

**Factor 2: Supply-side & Productivity Cycle**

Leather Products show a clear increase in loadings after 2011, indicating an increasing exposure of Leather Products to supply-side conditions in the post-crisis period. Printing loads relatively strongly before 2005 and weakens thereafter. Paper Products exhibit persistently positive and relatively high loadings throughout the period, indicating a stable and sustained exposure to supply-side conditions.

  Petroleum refining loads strongly on the supply-side factor around 2005, weakens for a long period after the financial crisis, and re-emerges again after 2023. Taken together, these patterns point to a time-varying role of energy-related and input-intensive industries in shaping the cross-sectional structure of supply-side conditions, with their relevance declining in the post-crisis period and re-intensifying amid recent energy-market disruptions.



```{r}
print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = 3, fade_highvol = TRUE, top_k = 5))
```
  
**Factor 3: Consumer Demand & Price Dynamics**

Factor 3 has a diffuse and low-amplitude cross-sectional structure with limited concentration in any single sector. This reflects the smoother and more persistent nature of consumer demand dynamics, which tend to adjust gradually and are less prone to abrupt reallocation across industries than investment-driven cycles. Beverages increase gradually over time, signaling growing integration with household demand cycles. 

It indicates that no industry persistently dominates this dimension, resulting in weaker cross-sectional identification and higher uncertainty in the estimated factor path, consistent with its relatively high estimation uncertainty and moderate sample stability.


\vspace{0.3cm}


```{r}
print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = 4, fade_highvol = TRUE, top_k = 5))
```

**Factor 4: Global Industrial Demand**

Basic Metals maintain high loadings over the entire sample, underscoring their sustained exposure to external demand. Wood Products load weakly throughout, suggesting limited sensitivity to global trade cycles. Wearing Apparel shows a gradual increase during 2005–2008, consistent with the pre-crisis expansion in global trade. Coke and Petroleum Refining collapse after 2021, coinciding with energy market shocks on global industrial demand.

\vspace{0.5cm}

```{r}
print(plot_factor_all_industries(Lambda_g_tv, dates, factor_j = 5, fade_highvol = TRUE, top_k = 5))
```

**Factor 5: Specialized Manufacturing**

Factor 5 exhibits the greatest dispersion in loadings across industries, demonstrating its highly specialized structure. Motor Vehicles and Beverages load most strongly, while Other Manufacturing and Other Transport Equipment also show relatively pronounced and persistent exposure. 

  The strong cross-sectional heterogeneity and moderate time variation of Factor-5 loadings indicate that this factor captures a narrow, project-based or sector-specific dimension rather than a broad cyclical force. Consequently, differences across industries dominate over common temporal movements. As such, Factor 5 should be interpreted as reflecting localized manufacturing dynamics rather than a persistent source of aggregate co-movement.




\vspace{0.3cm}

```{r}
plot_industry_loadings <- function(
  g,
  Lambda_g_tv,
  dates,
  event_years = c(2008, 2011, 2020),
  show_event_labels = TRUE
) {
  stopifnot(g %in% names(Lambda_g_tv))
  lambda_t <- Lambda_g_tv[[g]]  # T x r
  stopifnot(nrow(lambda_t) == length(dates))

  r <- ncol(lambda_t)

  df <- data.frame(date = dates, lambda_t, check.names = FALSE)
  df_long <- tidyr::pivot_longer(
    df,
    cols = starts_with("lambda_"),
    names_to = "factor",
    values_to = "loading"
  )

  # events
  event_dates  <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates  <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  ind_name <- get_industry_name_en(g)

  # aesthetics (match your earlier "clean academic" style)
  line_col <- adjustcolor("dodgerblue3", alpha.f = 0.95)
  line_lwd <- 0.95
  evt_col  <- "red3"
  evt_lty  <- 2
  evt_lwd  <- 0.8

  p <- ggplot2::ggplot(df_long, ggplot2::aes(x = date, y = loading)) +
    ggplot2::geom_line(color = line_col, linewidth = line_lwd) +
    ggplot2::geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    ggplot2::facet_wrap(~ factor, ncol = 1, scales = "free_y") +
    ggplot2::labs(
      title = paste0("Time-varying loadings (industry-only): ", ind_name),
      subtitle = paste0("Industry code: ", g),
      x = NULL,
      y = "Loading"
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(
      plot.title.position = "plot",
      strip.text = ggplot2::element_text(face = "bold"),
      panel.grid.minor = ggplot2::element_blank(),
      plot.margin = grid::unit(c(8, 8, 8, 8), "pt")
    )

  # event lines
  if (length(event_dates) > 0) {
    p <- p +
      ggplot2::geom_vline(
        xintercept = event_dates,
        linetype = evt_lty,
        color = evt_col,
        linewidth = evt_lwd
      )
  }

  if (show_event_labels && length(event_dates) > 0) {
    first_factor <- sort(unique(df_long$factor))[1]

    evt_df <- data.frame(
      date = event_dates,
      label = event_labels,
      factor = first_factor
    )

    p <- p +
      ggplot2::geom_text(
        data = evt_df,
        ggplot2::aes(x = date, y = Inf, label = label),
        inherit.aes = FALSE,
        vjust = 1.2,
        hjust = 0,
        color = evt_col,
        size = 3
      ) +
      ggplot2::coord_cartesian(clip = "off")
  }

  p
}

```


```{r fig.width=12, fig.height=8, include=FALSE}
g_example <- names(Lambda_g_tv)[5]
plot_industry_loadings(g_example, Lambda_g_tv, dates)
```


```{r}
plot_many_industries_5facets <- function(
  Lambda_g_tv,
  dates,
  industries = NULL,       
  event_years = c(2008, 2011, 2020),
  n_per_page = 6            
) {
  gs <- names(Lambda_g_tv)
  if (!is.null(industries)) gs <- intersect(gs, industries)
  stopifnot(length(gs) > 0)

  event_dates  <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates  <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  pages <- split(gs, ceiling(seq_along(gs) / n_per_page))

  plot_one_page <- function(gs_page) {
    df_long <- do.call(rbind, lapply(gs_page, function(g) {
      L <- Lambda_g_tv[[g]]  # T x r
      stopifnot(nrow(L) == length(dates))
      df <- data.frame(date = dates, L, check.names = FALSE)
      df <- tidyr::pivot_longer(df, cols = starts_with("lambda_"),
                                names_to = "factor", values_to = "loading")
      df$industry_code <- g
      df$industry_name <- get_industry_name_en(g)
      df
    }))

    evt_df <- data.frame(
      date = event_dates,
      label = event_labels,
      industry_name = unique(df_long$industry_name)[1],
      factor = sort(unique(df_long$factor))[1]
    )

    p <- ggplot2::ggplot(df_long, ggplot2::aes(x = date, y = loading)) +
      ggplot2::geom_line(color = adjustcolor("dodgerblue3", 0.95), linewidth = 0.6) +
      ggplot2::geom_hline(yintercept = 0, linetype = 3, color = "gray60", linewidth = 0.3) +
      ggplot2::facet_grid(industry_name ~ factor, scales = "free_y") +
      ggplot2::labs(
        title = "Industry-only time-varying loadings (many industries on one page)",
        subtitle = "Rows = industries, columns = loadings (lambda_1 ... lambda_r)",
        x = NULL, y = "Loading"
      ) +
      ggplot2::theme_minimal(base_size = 11) +
      ggplot2::theme(
        panel.grid.minor = ggplot2::element_blank(),
        strip.text.x = ggplot2::element_text(face = "bold"),
        strip.text.y = ggplot2::element_text(face = "bold"),
        plot.margin = grid::unit(c(8, 8, 8, 8), "pt")
      )

    if (length(event_dates) > 0) {
      p <- p +
        ggplot2::geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.4) +
        ggplot2::geom_text(
          data = evt_df,
          ggplot2::aes(x = date, y = Inf, label = label),
          inherit.aes = FALSE,
          vjust = 1.2, hjust = 0,
          color = "red3",
          size = 2.8
        ) +
        ggplot2::coord_cartesian(clip = "off")
    }

    p
  }

  lapply(pages, plot_one_page)
}

```


```{r fig.width=14, fig.height=10, include=FALSE}
plots <- plot_many_industries_5facets(Lambda_g_tv, dates, n_per_page = 9)
print(plots[[1]])  
```






### Instability in Factor Exposure

```{r, include=FALSE}
get_top5_industries_by_factor <- function(Lambda_g_tv, top_k = 5) {
  gs <- names(Lambda_g_tv)
  r  <- ncol(Lambda_g_tv[[1]])

  res <- do.call(rbind, lapply(1:r, function(j) {

    vol_df <- data.frame(
      industry_code = gs,
      industry_name = sapply(gs, get_industry_name_en),
      volatility = sapply(gs, function(g) {
        L <- Lambda_g_tv[[g]][, j]
        sd(diff(L), na.rm = TRUE)
      }),
      stringsAsFactors = FALSE
    )

    vol_df |>
      dplyr::arrange(dplyr::desc(volatility)) |>
      dplyr::slice_head(n = top_k) |>
      dplyr::mutate(
        factor = paste0("Factor ", j),
        rank = dplyr::row_number()
      ) |>
      dplyr::select(factor, rank, industry_code, industry_name, volatility)

  }))
  
  rownames(res) <- NULL
  res
}
```

\[
\text{volatility}_{g,j}
=
\operatorname{sd}\!\left(\Delta \lambda_{g,j,t}\right),
\]
where $\Delta \lambda_{g,j,t}$ denotes the first difference of the time-varying loading for industry $g$ and factor $j$.

This measure captures temporal instability in factor exposure and captures industries whose relationship with a given factor is subject to pronounced structural adjustment.


```{r}
top5_by_factor <- get_top5_industries_by_factor(Lambda_g_tv, top_k = 5)

print(top5_by_factor)
```


Leather Products exhibits the highest instability across multiple factors, it appear subject to erratic external shocks, likely driven by sudden supply chain disruptions and abrupt shifts in consumer preferences. Coke and Petroleum Refining also has consistently high volatility across all factors, aligning with the twin shocks of aggressive climate policy pivots and the Russia-Ukraine conflict-induced energy crisis.

Printing, and Furniture experience slow but persistent shifts rather than abrupt breaks, consistent with a declining role of traditional manufacturing segments.


**Summary**

The results show that allowing loadings to vary reveals non-negligible heterogeneity across industries. Importantly, the magnitude and timing of these variations are economically meaningful and align with major business-cycle disruptions. Several industries exhibit relatively stable loading paths throughout the sample, suggesting a persistent and robust association with the underlying common factors. In contrast, some industries display pronounced time variation.

However, the industry-level grouping remains too coarse to attribute the detected instability to specific types of survey information. In particular, indicators capturing current assessments, expectations, prices, and short-term changes may differ substantially in how their informational content evolves over time. 

For next step, we therefore extend the analysis by allowing factor loadings to vary jointly across industries and question blocks, enabling a sharper assessment of which types of survey indicators drive the common factors in different economic regimes.







