---
title: "03.2 Industry × Question Block Time-Varying Factor Loadings"
author: "Weiyun Hong"
date: "2025-12-28"
header-includes:
  - \usepackage{parskip}
  - \setlength{\parskip}{0.5em}
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
```

```{r}
dfm02 <- readRDS("output/dfm_results_02.rds")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered
source("ifo_data_processing.R")
```

### Industry $\times$ Question Block Time-Varying Factor Loadings

While the dynamic factor model introduced in Section~02 allows the latent factors 
$F_t$ to evolve over time, it assumes a stable measurement relationship between 
observed survey variables and the factors. In practice, however, the relevance of 
different survey questions for the underlying business cycle may change gradually 
due to structural transformation, institutional reforms, or shifts in economic 
expectations.

To allow for \emph{gradual structural change} in the measurement equation, we extend 
the model by allowing factor loadings to vary over time. To retain parsimony, we impose 
a grouping structure based on both industry classification and question block.

Let variable $i$ belong to Level-1 industry group $g(i)\in\{1,\dots,G\}$ and question 
block $b(i)\in\{A,B,C,D\}$. All variables within the same $(g,b)$ cell share a common, 
possibly time-varying, loading vector.

**Measurement Equation**

The measurement equation becomes
\begin{equation}
X_{i,t}
=
\lambda_{g(i),b(i),t}^\top F_t + \xi_{i,t},
\qquad
\xi_{i,t} \sim \mathcal{N}(0,H_i),
\end{equation}
where $F_t\in\mathbb{R}^r$ denotes the dynamic factor vector obtained from the 
Kalman smoother in Section~02, and 
$\lambda_{g,b,t}\in\mathbb{R}^r$ is the loading vector shared by all variables 
belonging to industry $g$ and question block $b$.

**Time Evolution of Loadings**

We model gradual structural change in the measurement relationship by assuming 
that the group-level loadings follow a random walk:
\begin{equation}
\lambda_{g,b,t}
=
\lambda_{g,b,t-1} + \eta_{g,b,t},
\qquad
\eta_{g,b,t} \sim \mathcal{N}(0, q_{g,b} I_r),
\end{equation}
where $q_{g,b}$ controls the degree of time variation in the loadings for 
industry--block cell $(g,b)$.

**Stacked Observation Equation**

Stacking the $n_{g,b}$ observed variables belonging to cell $(g,b)$ yields the 
multivariate observation equation
\begin{equation}
X_{g,b,t}
=
Z_{g,b,t}\,\lambda_{g,b,t} + \xi_{g,b,t},
\qquad
\xi_{g,b,t} \sim \mathcal{N}(0,H_{g,b}),
\end{equation}
where the time-varying design matrix is given by
\begin{equation}
Z_{g,b,t}
=
\mathbf{1}_{n_{g,b}} F_t^\top
\;\in\;
\mathbb{R}^{n_{g,b}\times r},
\end{equation}
and
\begin{equation}
H_{g,b}
=
\mathrm{diag}(H_i : i\in(g,b)).
\end{equation}

```{r}
library(KFAS)
library(tidyverse)
library(dplyr)

X_kfas <- Y_kfas
F_use <- as.matrix(F_smooth) # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(X_kfas)
stopifnot(ncol(F_use) == r, nrow(X_kfas) == T_obs, length(Hdiag) == N_obs)
```




```{r}
extract_qcode <- function(vn) {
  m <- str_match(vn, "_([A-Za-z]{2,4})$")[,2]
  toupper(m)
}
q_code <- extract_qcode(var_names_filtered)

block_map <- function(q) {
  if (is.na(q)) return(NA_character_)
  if (q %in% c("GUS","BUS","XUS","LUS")) return("A") # Current
  if (q %in% c("GES","QES","XES","BES","PRS")) return("B") # Expectations
  if (q %in% c("PWS","PVS")) return("C") # Prices
  if (q %in% c("AVS", "BVS","QVS")) return("D") # MoM changes
  return(NA_character_)
}
q_block <- vapply(q_code, block_map, character(1))

table(q_block, useNA = "ifany")
```

```{r}
industry_id <- as.character(var_metadata_filtered$industry_code)

cell_id <- ifelse(is.na(q_block),
                  paste0(industry_id, "_", "UNK"),
                  paste0(industry_id, "_", q_block))

cells <- sort(unique(cell_id))
K <- length(cells)
cat("Number of (industry × block) cells:", K, "\n")

cell_sizes <- as.integer(table(cell_id)[cells])
cell_table <- data.frame(cell = cells, n_cell = cell_sizes) |>
  arrange(desc(n_cell))
```

```{r}
fit_tvp_cell <- function(X_cell, Fmat, H_vec, q_init = 1e-4) {
  
  T_obs <- nrow(Fmat)
  r <- ncol(Fmat)
  n_cell <- ncol(X_cell)

  # Z_{t} = 1_{n_cell} x F_t'
  Z_arr <- array(NA_real_, dim = c(n_cell, r, T_obs))
  for (t in 1:T_obs) {
    Z_arr[,,t] <- matrix(rep(Fmat[t, ], each = n_cell), nrow = n_cell, byrow = FALSE)
  }
  
  model0 <- SSModel(
    X_cell ~ -1 + 
      SSMcustom(
        Z = Z_arr,
        T = diag(r),
        R = diag(r),
        Q = diag(NA_real_, r), # updated to q*I_r in updatefn
        a1 = rep(0, r),
        P1 = diag(10, r)
      ),
    H = diag(H_vec, n_cell)    # fixed diagonal obs noise per series
  )
  
  updatefn <- function(pars, model) {
    q <- exp(pars[1])
    model$Q[,,1] <- diag(q, r)
    model
  }
  
  # Solving for maximum likelihood estimation q
  fit <- fitSSM(
    model0,
    inits = log(q_init),
    updatefn = updatefn,
    method = "BFGS"
  )
  model <- fit$model
  q_hat <- exp(fit$optim.out$par[1])
  
  kfs <- KFS(model, smoothing = c("state"))
  lambda_t <- kfs$alphahat # T x r
  
  colnames(lambda_t) <- paste0("lambda_", 1:r)
  list(lambda_t = lambda_t, q_hat = q_hat)
}
```

```{r, include=T}
# Estimate time-varying loadings for all (industry × block) cells
Lambda_cell_tv <- vector("list", K)
q_hat_cell <- rep(NA_real_, K)
names(Lambda_cell_tv) <- cells
names(q_hat_cell) <- cells

for (kk in seq_along(cells)) {
  c_id <- cells[kk]
  idx  <- which(cell_id == c_id)
  
  X_cell <- as.matrix(X_kfas[, idx, drop = FALSE]) # T x n_cell
  H_vec  <- Hdiag[idx]
  
  res <- fit_tvp_cell(X_cell = X_cell, Fmat = F_use, H_vec = H_vec, q_init = 1e-4)
  
  Lambda_cell_tv[[c_id]] <- res$lambda_t
  q_hat_cell[c_id]       <- res$q_hat
  
  if (kk %% 20 == 0) cat("Done cell", kk, "/", K, "\n")
}

summary(q_hat_cell)

# Expand cell-level paths to variable-level time-varying loadings:
# each variable i inherits its cell's loading path lambda_{industry(i),block(i),t}
Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

for (i in 1:N_obs) {
  c_id <- cell_id[i]
  Lambda_tv_list[[i]] <- Lambda_cell_tv[[c_id]] # T x r
}
```

```{r, include=T}
parse_cell <- function(cell) {
  parts <- str_split(cell, "_", simplify = TRUE)
  data.frame(
    industry_code = parts[1], 
    block = parts[2], 
    cell = cell, 
    stringsAsFactors = FALSE
  )
}

# Create a metadata table for all unique cells
cell_meta <- bind_rows(lapply(cells, parse_cell)) |>
  mutate(n_cell = as.integer(table(cell_id)[cell]))

head(cell_meta, 10)
```

```{r, include=T}
# Plot loading paths over time for a specific cell
plot_cell_loadings <- function(cell) {
  # Retrieve estimated time-varying loadings (T x r)
  lambda_t <- Lambda_cell_tv[[cell]]
  meta <- parse_cell(cell)
  
  # Prepare data for ggplot
  df <- data.frame(date = dates, lambda_t)
  df_long <- pivot_longer(
    df, 
    cols = starts_with("lambda_"),
    names_to = "factor", 
    values_to = "loading"
  )
  
  # Create the plot
  ggplot(df_long, aes(x = date, y = loading, group = factor)) +
    geom_line(color = "steelblue") +
    facet_wrap(~ factor, ncol = 1, scales = "free_y") +
    labs(
      title = paste0("TV Loadings: Industry = ", meta$industry_code, ", Block = ", meta$block),
      subtitle = paste("Based on", meta$n_cell, "series"),
      x = NULL, 
      y = "Loading Coefficient"
    ) +
    theme_minimal()
}

# Example: Visualize the largest cell (the one with the most series)
cell_example <- cell_table$cell[1]
print(plot_cell_loadings(cell_example))
```


```{r, include=T}
# Calculate instability metrics for every cell
instability_by_cell <- lapply(cells, function(cell) {
  L  <- Lambda_cell_tv[[cell]] # T x r matrix
  dL <- apply(L, 2, diff)      # Calculate first differences (T-1 x r)
  
  meta <- parse_cell(cell)
  
  data.frame(
    industry_code    = meta$industry_code,
    block            = meta$block,
    cell             = cell,
    n_cell           = sum(cell_id == cell),
    q_hat            = q_hat_cell[cell],
    mean_abs_dlambda = mean(abs(dL), na.rm = TRUE),
    sd_dlambda       = sd(as.numeric(dL), na.rm = TRUE)
  )
}) |> 
  bind_rows() |> 
  arrange(desc(mean_abs_dlambda))

# Show the top 20 most unstable cells
head(instability_by_cell, 20)
```

```{r}
qtype_labels <- c(
  A = "Current assessments (GUS/LUS/BUS/XUS)",
  B = "Expectations (GES/QES/XES/BES/PRS)",
  C = "Prices (PVS/PWS)",
  D = "MoM changes (AVS/BVS/QVS)"
)
```


```{r}
instability_by_block <- instability_by_cell |>
  dplyr::group_by(block) |>
  dplyr::summarise(
    n_cells = dplyr::n(),
    total_series = sum(n_cell),
    median_q = median(q_hat, na.rm = TRUE),
    median_mean_abs_dlambda = median(mean_abs_dlambda, na.rm = TRUE),
    .groups = "drop"
  )


# Group instability metrics by business block (A, B, C, D)
instability_by_qtype <- instability_by_block |>
  mutate(
    question_type = qtype_labels[block]
  ) |>
  dplyr::select(
    question_type,
    n_cells,
    total_series,
    median_q,
    median_mean_abs_dlambda
  ) |>
  rename(
    `Question type`              = question_type,
    `# Cells (industry × type)`  = n_cells,
    `# Series`                   = total_series,
    `Median q (RW variance)`     = median_q,
    `Median |Δλ|`                = median_mean_abs_dlambda
  )

print(instability_by_qtype)

```

The stability of factor loadings differs across question types: current assessments are the most stable, expectations and price-related questions exhibit moderate time variation, while month-on-month indicators display the highest degree of instability.



```{r}
plot_loading_trends_by_qtype <- function(
  Lambda_cell_tv,
  dates,
  factor_j = 1,
  qtype = "B",                      # A/B/C/D
  industries = NULL,                # NULL = All industries
  event_years = c(2008, 2011, 2020),
  show_legend = TRUE
) {
  stopifnot(qtype %in% c("A","B","C","D"))

  parse_cell <- function(cell) {
    parts <- strsplit(cell, "_", fixed = TRUE)[[1]]
    list(industry_code = parts[1], qtype = parts[2])
  }

  cells <- names(Lambda_cell_tv)
  meta <- lapply(cells, parse_cell)

  meta_df <- data.frame(
    cell = cells,
    industry = vapply(meta, `[[`, character(1), "industry_code"),
    qtype = vapply(meta, `[[`, character(1), "qtype"),
    stringsAsFactors = FALSE
  )

  meta_df <- meta_df[meta_df$qtype == qtype, , drop = FALSE]

  # optional subset industries
  if (!is.null(industries)) {
    meta_df <- meta_df[meta_df$industry %in% industries, , drop = FALSE]
  }
  stopifnot(nrow(meta_df) > 0)

  # build long df: one line per industry
  df_long <- do.call(rbind, lapply(seq_len(nrow(meta_df)), function(k) {
    cell_k <- meta_df$cell[k]
    ind_k  <- meta_df$industry[k]
    L <- Lambda_cell_tv[[cell_k]]  # T x r
    if (is.null(L) || ncol(L) < factor_j) return(NULL)
    data.frame(
      date = dates,
      industry = ind_k,
      industry_name = get_industry_name_en(ind_k),
      loading = as.numeric(L[, factor_j]),
      stringsAsFactors = FALSE
    )
  }))

  # event lines
  event_dates  <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates  <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  # title label
  qtype_title <- unname(qtype_labels[qtype])

  p <- ggplot2::ggplot(df_long, aes(x = date, y = loading, color = industry_name)) +
    ggplot2::geom_line(linewidth = 0.9, alpha = 0.85) +
    ggplot2::geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    ggplot2::scale_color_viridis_d(option = "turbo") +
    ggplot2::labs(
      title = paste0("Time-varying loadings by industry — ", qtype_title),
      subtitle = paste0("Factor ", factor_j),
      x = NULL,
      y = paste0("Loading (lambda_", factor_j, ")"),
      color = "Industry (Level 1)"
    ) +
    ggplot2::theme_minimal(base_size = 12)
  

  if (!show_legend) {
    p <- p + ggplot2::theme(legend.position = "none")
  } else {
    p <- p + ggplot2::theme(
      legend.position = "bottom",
      legend.key.height = grid::unit(0.45, "lines")
    ) +
      ggplot2::guides(color = ggplot2::guide_legend(ncol = 3))
  }

  if (length(event_dates) > 0) {
    p <- p +
      ggplot2::geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.8) +
      ggplot2::annotate(
        "text",
        x = event_dates,
        y = Inf,
        label = event_labels,
        vjust = 1.2,
        hjust = 0,
        color = "red3",
        size = 3
      )
  }

  p
}

```

**Factor 1 — Capital goods Manufacturing cycle**

Current assessment indicators provide a stable and persistent anchor for the factor, while month-on-month changes contribute higher-frequency variation that reacts more strongly to short-term shocks. The higher volatility of expectations and monthly changes are particularly informative during periods of economic stress, when they respond rapidly to shocks and temporary decouplings in business sentiment. While their loading paths appear irregular, this behavior reflects their role as high-frequency adjustment signals that react rapidly to sudden shocks, such as production disruptions or abrupt changes in demand, rather than persistent cyclical movements.

```{r fig.width=14, fig.height=5}
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 1, qtype = "A")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 1, qtype = "B")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 1, qtype = "C")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 1, qtype = "D")
```
**Factor 2 — Supply-side & Productivity Cycle**
Factor 2 represents a "benign supply-side expansion," characterized by a distinct inverse relationship where high optimism in upstream industries coexists with contained price pressures downstream.


```{r fig.width=14, fig.height=5}
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 2, qtype = "A")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 2, qtype = "B")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 2, qtype = "C")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 2, qtype = "D")
```




```{r fig.width=14, fig.height=5}
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 3, qtype = "A")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 3, qtype = "B")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 3, qtype = "C")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 3, qtype = "D")

```
**Factor 3: Consumer Demand & Price Dynamics**





```{r fig.width=14, fig.height=5}
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 4, qtype = "A")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 4, qtype = "B")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 4, qtype = "C")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 4, qtype = "D")
```

```{r fig.width=14, fig.height=5}
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 5, qtype = "A")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 5, qtype = "B")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 5, qtype = "C")
plot_loading_trends_by_qtype(Lambda_cell_tv, dates, factor_j = 5, qtype = "D")
```
Across all factors, indicators in Group A - the contemporaneous assessments of business conditions and order books (GUS/LUS/BUS/XUS) — tend to deliver the most stable and persistent measurement relationship with the latent factors, making them a reliable backbone for tracking the underlying manufacturing cycle. By contrast, Group D — month-on-month change indicators (AVS/BVS/QVS) — exhibits markedly higher volatility and stronger time dependence in loadings. While these trajectories can look visually “noisy”, the pattern is consistent with D capturing short-term adjustment dynamics that become especially informative during abrupt disruptions, when firms react quickly to shocks and constraints. 

Group B expectations (GES/QES/XES/BES/PRS) typically lie between these two extremes: their loadings vary more than A but less erratically than D, with more pronounced fluctuations around major stress episodes, suggesting that forward-looking sentiment can intermittently decouple from realized conditions under heightened uncertainty. Group C price indicators (PVS/PWS) also show intermediate behavior and tend to become more salient when cost pressures and inflation-related forces dominate the manufacturing environment. Taken together, the results imply that the informational content of the ifo manufacturing survey is systematically state-dependent: stable “level-type” assessments anchor the common component, whereas high-frequency changes and expectations contribute disproportionately around turning points and short-lived shocks.

These suggest that, separating structural signal from shock sensitivity is crucial when studying propagation and regime changes. For causal and lead–lag analysis, it is generally more defensible to treat Group A as the anchor (a stable measurement layer) and use Groups B and D for early-moving or shock-responsive channels, rather than pooling all questions indiscriminately. 


```{r, include=T}
plot_factor_block_all_industries <- function(
  Lambda_cell_tv, dates,
  factor_j = 1, block = "B",
  event_years = c(2008, 2011, 2020),
  highlight_top_k = 6,
  highlight_by = c("instability", "mean_abs"),
  alpha_other = 0.25
) {
  highlight_by <- match.arg(highlight_by)

  # cell metadata
  parse_cell <- function(cell) {
    parts <- strsplit(cell, "__", fixed = TRUE)[[1]]
    if (length(parts) == 1) parts <- strsplit(cell, "_", fixed = TRUE)[[1]]
    list(industry = parts[1], block = parts[2])
  }
  cells <- names(Lambda_cell_tv)
  meta <- tibble(cell = cells) |>
    mutate(tmp = lapply(cell, parse_cell),
           industry = map_chr(tmp, "industry"),
           blk      = map_chr(tmp, "block")) |>
    dplyr:: select(-tmp) |>
    filter(blk == block)

  # build long df: 21 lines
  df <- map_dfr(meta$cell, function(c_id) {
    L <- Lambda_cell_tv[[c_id]]
    tibble(date = dates,
           industry = meta$industry[meta$cell == c_id],
           loading = as.numeric(L[, factor_j]))
  })

  # choose which lines to highlight
  score <- df |>
    group_by(industry) |>
    summarise(
      mean_abs = mean(abs(loading), na.rm = TRUE),
      instability = mean(abs(diff(loading)), na.rm = TRUE),
      .groups = "drop"
    )

  top_inds <- score |>
    arrange(desc(.data[[highlight_by]])) |>
    slice_head(n = highlight_top_k) |>
    pull(industry)

  df <- df |>
    mutate(is_highlight = industry %in% top_inds)

  # events
  event_dates <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  ggplot(df, aes(x = date, y = loading, group = industry)) +
    geom_line(aes(color = industry, alpha = is_highlight), linewidth = 0.9) +
    scale_alpha_manual(values = c(`TRUE` = 0.95, `FALSE` = alpha_other), guide = "none") +
    scale_color_viridis_d(option = "turbo") +
    geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.8) +
    annotate("text", x = event_dates, y = Inf, label = event_labels,
             vjust = 1.2, hjust = 0, color = "red3", size = 3) +
    labs(
      title = paste0("TV Loadings across industries: Factor ", factor_j, ", Block = ", block),
      subtitle = paste0("Highlight top-", highlight_top_k, " by ", highlight_by, " (others faded)"),
      x = NULL,
      y = paste0("Loading (lambda_", factor_j, ")"),
      color = "Industry"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom") +
    guides(color = guide_legend(ncol = 3, override.aes = list(alpha = 1)))
}

```



```{r fig.width=13, fig.height=6, include=T}
plot_factor_block_all_industries(Lambda_cell_tv, dates, factor_j = 2, block = "A",
                                 highlight_top_k = 6, highlight_by = "instability")
```






```{r, include=T}
source("0302_followup_analysis.R")
   
all_results <- run_all_0302_followup_analyses(
     Lambda_cell_tv, dates, instability_by_qtype
   )
```


```{r 0302-save-for-04, message=FALSE, warning=FALSE}
# Save objects for Section 04
stopifnot(exists("X_kfas"), exists("F_smooth"), exists("dates"))
stopifnot(nrow(X_kfas) == nrow(as.matrix(F_smooth)))
stopifnot(nrow(X_kfas) == length(dates))

var_map_04 <- data.frame(
  series_id = colnames(X_kfas),
  var_name  = var_names_filtered,
  industry  = as.character(var_metadata_filtered$industry_code),
  q_block   = q_block,
  cell_id   = cell_id,
  stringsAsFactors = FALSE
)

dfm_state_04 <- list(
  created_at = Sys.time(),
  dates      = as.Date(dates),
  r          = r,
  F_smooth   = as.matrix(F_smooth),   # T x r
  X_kfas     = as.matrix(X_kfas),     # T x N
  var_map    = var_map_04,            # N rows
  Lambda_cell_tv = if (exists("Lambda_cell_tv")) Lambda_cell_tv else NULL
)

# save 
out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
saveRDS(dfm_state_04, file = file.path(out_dir, "dfm_state_04.rds"))

str(dfm_state_04, max.level = 1)

```




你在 0302 已经把每个变量分到了 `cell_id = industry_code + "_" + block`，而且一共 `K = 84` 个 cells :contentReference[oaicite:1]{index=1}。  
所以 04 最自然的候选预测变量就是：

\[
x_{c,t} \equiv x_{g,b,t} = \frac{1}{n_{g,b}} \sum_{i \in (g,b)} X_{i,t}
\]

也就是：**同一个 industry×block cell 内的变量在每期取均值**，得到一个 \(T \times 84\) 的候选矩阵。













