---
title: "0302_supplement_YVFL"
author: "Yiming Zhao"
date: "2026-02-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
```
Industry $\times$ Question Block Time-Varying Factor Loadings

# Extension 1: Level 2
```{r}
dfm02 <- readRDS("output/dfm_results_02_level2.rds")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered
source("ifo_data_processing.R")

library(KFAS)
library(tidyverse)
library(dplyr)

X_kfas <- Y_kfas
F_use <- as.matrix(F_smooth) # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(X_kfas)
stopifnot(ncol(F_use) == r, nrow(X_kfas) == T_obs, length(Hdiag) == N_obs)
```


```{r }
extract_qcode <- function(vn) {
  m <- str_match(vn, "_([A-Za-z]{2,4})$")[,2]
  toupper(m)
}
q_code <- extract_qcode(var_names_filtered)

block_map <- function(q) {
  if (is.na(q)) return(NA_character_)
  if (q %in% c("GUS","BUS","XUS","LUS")) return("A") # Current
  if (q %in% c("GES","QES","XES","BES","PRS")) return("B") # Expectations
  if (q %in% c("PWS","PVS")) return("C") # Prices
  if (q %in% c("AVS", "BVS","QVS")) return("D") # MoM changes
  return(NA_character_)
}
q_block <- vapply(q_code, block_map, character(1))

table(q_block, useNA = "ifany")
```
```{r}
industry_id <- as.character(var_metadata_filtered$industry_code)

cell_id <- ifelse(is.na(q_block),
                  paste0(industry_id, "_", "UNK"),
                  paste0(industry_id, "_", q_block))

cells <- sort(unique(cell_id))
K <- length(cells)
cat("Number of (industry × block) cells:", K, "\n")

cell_sizes <- as.integer(table(cell_id)[cells])
cell_table <- data.frame(cell = cells, n_cell = cell_sizes) |>
  arrange(desc(n_cell))
```

```{r}
fit_tvp_cell <- function(X_cell, Fmat, H_vec, q_init = 1e-4) {
  
  T_obs <- nrow(Fmat)
  r <- ncol(Fmat)
  n_cell <- ncol(X_cell)

  # Z_{t} = 1_{n_cell} x F_t'
  Z_arr <- array(NA_real_, dim = c(n_cell, r, T_obs))
  for (t in 1:T_obs) {
    Z_arr[,,t] <- matrix(rep(Fmat[t, ], each = n_cell), nrow = n_cell, byrow = FALSE)
  }
  
  model0 <- SSModel(
    X_cell ~ -1 + 
      SSMcustom(
        Z = Z_arr,
        T = diag(r),
        R = diag(r),
        Q = diag(NA_real_, r), # updated to q*I_r in updatefn
        a1 = rep(0, r),
        P1 = diag(10, r)
      ),
    H = diag(H_vec, n_cell)    # fixed diagonal obs noise per series
  )
  
  updatefn <- function(pars, model) {
    q <- exp(pars[1])
    model$Q[,,1] <- diag(q, r)
    model
  }
  
  # Solving for maximum likelihood estimation q
  fit <- fitSSM(
    model0,
    inits = log(q_init),
    updatefn = updatefn,
    method = "BFGS"
  )
  model <- fit$model
  q_hat <- exp(fit$optim.out$par[1])
  
  kfs <- KFS(model, smoothing = c("state"))
  lambda_t <- kfs$alphahat # T x r
  
  colnames(lambda_t) <- paste0("lambda_", 1:r)
  list(lambda_t = lambda_t, q_hat = q_hat)
}
```

```{r, include=FALSE}
# Estimate time-varying loadings for all (industry × block) cells
Lambda_cell_tv <- vector("list", K)
q_hat_cell <- rep(NA_real_, K)
names(Lambda_cell_tv) <- cells
names(q_hat_cell) <- cells

for (kk in seq_along(cells)) {
  c_id <- cells[kk]
  idx  <- which(cell_id == c_id)
  
  X_cell <- as.matrix(X_kfas[, idx, drop = FALSE]) # T x n_cell
  H_vec  <- Hdiag[idx]
  
  res <- fit_tvp_cell(X_cell = X_cell, Fmat = F_use, H_vec = H_vec, q_init = 1e-4)
  
  Lambda_cell_tv[[c_id]] <- res$lambda_t
  q_hat_cell[c_id]       <- res$q_hat
  
  if (kk %% 20 == 0) cat("Done cell", kk, "/", K, "\n")
}

summary(q_hat_cell)

# Expand cell-level paths to variable-level time-varying loadings:
# each variable i inherits its cell's loading path lambda_{industry(i),block(i),t}
Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

for (i in 1:N_obs) {
  c_id <- cell_id[i]
  Lambda_tv_list[[i]] <- Lambda_cell_tv[[c_id]] # T x r
}

parse_cell <- function(cell) {
  parts <- str_split(cell, "_", simplify = TRUE)
  data.frame(
    industry_code = parts[1], 
    block = parts[2], 
    cell = cell, 
    stringsAsFactors = FALSE
  )
}

# Create a metadata table for all unique cells
cell_meta <- bind_rows(lapply(cells, parse_cell)) |>
  mutate(n_cell = as.integer(table(cell_id)[cell]))

head(cell_meta, 10)
```

```{r, include=FALSE}
# Calculate instability metrics for every cell
instability_by_cell <- lapply(cells, function(cell) {
  L  <- Lambda_cell_tv[[cell]] # T x r matrix
  dL <- apply(L, 2, diff)      # Calculate first differences (T-1 x r)
  
  meta <- parse_cell(cell)
  
  data.frame(
    industry_code    = meta$industry_code,
    block            = meta$block,
    cell             = cell,
    n_cell           = sum(cell_id == cell),
    q_hat            = q_hat_cell[cell],
    mean_abs_dlambda = mean(abs(dL), na.rm = TRUE),
    sd_dlambda       = sd(as.numeric(dL), na.rm = TRUE)
  )
}) |> 
  bind_rows() |> 
  arrange(desc(mean_abs_dlambda))

# Show the top 20 most unstable cells
head(instability_by_cell, 20)
```

The instability ranking reveals a striking concentration of high-volatility cells in block D (Month-on-Month changes). This indicates that temporal variation in factor loadings is predominantly driven by short-term change indicators rather than by current assessments, expectations, or price components.


```{r}
qtype_labels <- c(
  A = "Current assessments (GUS/LUS/BUS/XUS)",
  B = "Expectations (GES/QES/XES/BES/PRS)",
  C = "Prices (PVS/PWS)",
  D = "MoM changes (AVS/BVS/QVS)"
)

instability_by_block <- instability_by_cell |>
  dplyr::group_by(block) |>
  dplyr::summarise(
    n_cells = dplyr::n(),
    total_series = sum(n_cell),
    median_q = median(q_hat, na.rm = TRUE),
    median_mean_abs_dlambda = median(mean_abs_dlambda, na.rm = TRUE),
    .groups = "drop"
  )


# Group instability metrics by business block (A, B, C, D)
instability_by_qtype <- instability_by_block |>
  mutate(
    question_type = qtype_labels[block]
  ) |>
  dplyr::select(
    question_type,
    n_cells,
    total_series,
    median_q,
    median_mean_abs_dlambda
  ) |>
  rename(
    `Question type`              = question_type,
    `# Cells (industry × type)`  = n_cells,
    `# Series`                   = total_series,
    `Median q (RW variance)`     = median_q,
    `Median |Δλ|`                = median_mean_abs_dlambda
  )

print(instability_by_qtype)
```
The stability of factor loadings differs across question types: current assessments are the most stable, expectations and price-related questions exhibit moderate time variation, while month-on-month indicators display the highest degree of instability.

```{r }
plot_factor_block_all_industries <- function(
  Lambda_cell_tv, dates,
  factor_j = 1, block = "B",
  event_years = c(2008, 2011, 2020),
  top_k = 20,
  rank_by = c("instability", "mean_abs"),
  show_legend = FALSE
) {
  rank_by <- match.arg(rank_by)

  # parse cell id robustly: "C1010000_D" or "C1010000__D"
  parse_cell <- function(cell) {
    parts <- strsplit(cell, "__", fixed = TRUE)[[1]]
    if (length(parts) == 1) parts <- strsplit(cell, "_", fixed = TRUE)[[1]]
    list(industry = parts[1], blk = parts[2])
  }

  cells <- names(Lambda_cell_tv)

  meta <- tibble(cell = cells) |>
    mutate(tmp = lapply(cell, parse_cell),
           industry = purrr::map_chr(tmp, "industry"),
           blk      = purrr::map_chr(tmp, "blk")) |>
    dplyr::select(-tmp) |>
    dplyr::filter(blk == block)

  stopifnot(nrow(meta) > 0)

  # long df: one line per industry in this block
  df <- purrr::map_dfr(meta$cell, function(c_id) {
    L <- Lambda_cell_tv[[c_id]]
    tibble(
      date = dates,
      industry = meta$industry[meta$cell == c_id][1],
      loading = as.numeric(L[, factor_j])
    )
  }) |>
    dplyr::arrange(industry, date)

  # rank industries within this (factor, block)
  score <- df |>
    dplyr::group_by(industry) |>
    dplyr::summarise(
      mean_abs    = mean(abs(loading), na.rm = TRUE),
      instability = mean(abs(diff(loading)), na.rm = TRUE),
      .groups = "drop"
    )

  top_inds <- score |>
    dplyr::arrange(dplyr::desc(.data[[rank_by]])) |>
    dplyr::slice_head(n = top_k) |>
    dplyr::pull(industry)

  # keep only top-k industries
  df <- df |>
    dplyr::filter(industry %in% top_inds)

  # events
  event_dates <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  p <- ggplot(df, aes(x = date, y = loading, group = industry, color = industry)) +
    geom_line(linewidth = 0.9, alpha = 0.9) +
    scale_color_viridis_d(option = "turbo") +
    geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    labs(
      title = paste0("TV Loadings across industries: Factor ", factor_j, ", Block = ", block),
      subtitle = paste0("Showing top-", top_k, " industries by ", rank_by),
      x = NULL,
      y = paste0("Loading (lambda_", factor_j, ")"),
      color = "Industry"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = if (show_legend) "bottom" else "none") +
    guides(color = guide_legend(ncol = 3))

  if (length(event_dates) > 0) {
    p <- p +
      geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.8) +
      annotate("text", x = event_dates, y = Inf, label = event_labels,
               vjust = 1.2, hjust = 0, color = "red3", size = 3) +
      coord_cartesian(clip = "off")
  }

  p
}
```



```{r fig.width=13, fig.height=6}
for (j in 1:5) {

  print(plot_factor_block_all_industries(
    Lambda_cell_tv, dates,
    factor_j = j, block = "D",
    top_k = 20, rank_by = "instability",
    show_legend = TRUE
  ))

  print(plot_factor_block_all_industries(
    Lambda_cell_tv, dates,
    factor_j = j, block = "B",
    top_k = 20, rank_by = "instability",
    show_legend = TRUE
  ))
}
```

At the Level-2 aggregation, time variation in factor loadings is predominantly driven by short-term change indicators (Block D), while expectations (Block B) exhibit smoother and more structural dynamics. Crisis periods (2008, 2020) coincide with pronounced dispersion and structural shifts in industry-specific factor exposures, indicating that macro shocks alter not only factor realizations but also the cross-industry loading structure.


# Extension 2: Level 3

```{r}
dfm02 <- readRDS("output/dfm_results_02_level3.rds")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered
source("ifo_data_processing.R")

library(KFAS)
library(tidyverse)
library(dplyr)

X_kfas <- Y_kfas
F_use <- as.matrix(F_smooth) # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(X_kfas)
stopifnot(ncol(F_use) == r, nrow(X_kfas) == T_obs, length(Hdiag) == N_obs)
```


```{r }
extract_qcode <- function(vn) {
  m <- str_match(vn, "_([A-Za-z]{2,4})$")[,2]
  toupper(m)
}
q_code <- extract_qcode(var_names_filtered)

block_map <- function(q) {
  if (is.na(q)) return(NA_character_)
  if (q %in% c("GUS","BUS","XUS","LUS")) return("A") # Current
  if (q %in% c("GES","QES","XES","BES","PRS")) return("B") # Expectations
  if (q %in% c("PWS","PVS")) return("C") # Prices
  if (q %in% c("AVS", "BVS","QVS")) return("D") # MoM changes
  return(NA_character_)
}
q_block <- vapply(q_code, block_map, character(1))

table(q_block, useNA = "ifany")
```
```{r}
industry_id <- as.character(var_metadata_filtered$industry_code)

cell_id <- ifelse(is.na(q_block),
                  paste0(industry_id, "_", "UNK"),
                  paste0(industry_id, "_", q_block))

cells <- sort(unique(cell_id))
K <- length(cells)
cat("Number of (industry × block) cells:", K, "\n")

cell_sizes <- as.integer(table(cell_id)[cells])
cell_table <- data.frame(cell = cells, n_cell = cell_sizes) |>
  arrange(desc(n_cell))
```

```{r}
fit_tvp_cell <- function(X_cell, Fmat, H_vec, q_init = 1e-4) {
  
  T_obs <- nrow(Fmat)
  r <- ncol(Fmat)
  n_cell <- ncol(X_cell)

  # Z_{t} = 1_{n_cell} x F_t'
  Z_arr <- array(NA_real_, dim = c(n_cell, r, T_obs))
  for (t in 1:T_obs) {
    Z_arr[,,t] <- matrix(rep(Fmat[t, ], each = n_cell), nrow = n_cell, byrow = FALSE)
  }
  
  model0 <- SSModel(
    X_cell ~ -1 + 
      SSMcustom(
        Z = Z_arr,
        T = diag(r),
        R = diag(r),
        Q = diag(NA_real_, r), # updated to q*I_r in updatefn
        a1 = rep(0, r),
        P1 = diag(10, r)
      ),
    H = diag(H_vec, n_cell)    # fixed diagonal obs noise per series
  )
  
  updatefn <- function(pars, model) {
    q <- exp(pars[1])
    model$Q[,,1] <- diag(q, r)
    model
  }
  
  # Solving for maximum likelihood estimation q
  fit <- fitSSM(
    model0,
    inits = log(q_init),
    updatefn = updatefn,
    method = "BFGS"
  )
  model <- fit$model
  q_hat <- exp(fit$optim.out$par[1])
  
  kfs <- KFS(model, smoothing = c("state"))
  lambda_t <- kfs$alphahat # T x r
  
  colnames(lambda_t) <- paste0("lambda_", 1:r)
  list(lambda_t = lambda_t, q_hat = q_hat)
}
```

```{r, include=FALSE}
# Estimate time-varying loadings for all (industry × block) cells
Lambda_cell_tv <- vector("list", K)
q_hat_cell <- rep(NA_real_, K)
names(Lambda_cell_tv) <- cells
names(q_hat_cell) <- cells

for (kk in seq_along(cells)) {
  c_id <- cells[kk]
  idx  <- which(cell_id == c_id)
  
  X_cell <- as.matrix(X_kfas[, idx, drop = FALSE]) # T x n_cell
  H_vec  <- Hdiag[idx]
  
  res <- fit_tvp_cell(X_cell = X_cell, Fmat = F_use, H_vec = H_vec, q_init = 1e-4)
  
  Lambda_cell_tv[[c_id]] <- res$lambda_t
  q_hat_cell[c_id]       <- res$q_hat
  
  if (kk %% 20 == 0) cat("Done cell", kk, "/", K, "\n")
}

summary(q_hat_cell)

# Expand cell-level paths to variable-level time-varying loadings:
# each variable i inherits its cell's loading path lambda_{industry(i),block(i),t}
Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

for (i in 1:N_obs) {
  c_id <- cell_id[i]
  Lambda_tv_list[[i]] <- Lambda_cell_tv[[c_id]] # T x r
}

parse_cell <- function(cell) {
  parts <- str_split(cell, "_", simplify = TRUE)
  data.frame(
    industry_code = parts[1], 
    block = parts[2], 
    cell = cell, 
    stringsAsFactors = FALSE
  )
}

# Create a metadata table for all unique cells
cell_meta <- bind_rows(lapply(cells, parse_cell)) |>
  mutate(n_cell = as.integer(table(cell_id)[cell]))

head(cell_meta, 10)
```


```{r, include=FALSE}
# Calculate instability metrics for every cell
instability_by_cell <- lapply(cells, function(cell) {
  L  <- Lambda_cell_tv[[cell]] # T x r matrix
  dL <- apply(L, 2, diff)      # Calculate first differences (T-1 x r)
  
  meta <- parse_cell(cell)
  
  data.frame(
    industry_code    = meta$industry_code,
    block            = meta$block,
    cell             = cell,
    n_cell           = sum(cell_id == cell),
    q_hat            = q_hat_cell[cell],
    mean_abs_dlambda = mean(abs(dL), na.rm = TRUE),
    sd_dlambda       = sd(as.numeric(dL), na.rm = TRUE)
  )
}) |> 
  bind_rows() |> 
  arrange(desc(mean_abs_dlambda))

# Show the top 20 most unstable cells
head(instability_by_cell, 20)
```

The instability ranking reveals a striking concentration of high-volatility cells in block D (Month-on-Month changes). This indicates that temporal variation in factor loadings is predominantly driven by short-term change indicators rather than by current assessments, expectations, or price components.


```{r}
qtype_labels <- c(
  A = "Current assessments (GUS/LUS/BUS/XUS)",
  B = "Expectations (GES/QES/XES/BES/PRS)",
  C = "Prices (PVS/PWS)",
  D = "MoM changes (AVS/BVS/QVS)"
)

instability_by_block <- instability_by_cell |>
  dplyr::group_by(block) |>
  dplyr::summarise(
    n_cells = dplyr::n(),
    total_series = sum(n_cell),
    median_q = median(q_hat, na.rm = TRUE),
    median_mean_abs_dlambda = median(mean_abs_dlambda, na.rm = TRUE),
    .groups = "drop"
  )


# Group instability metrics by business block (A, B, C, D)
instability_by_qtype <- instability_by_block |>
  mutate(
    question_type = qtype_labels[block]
  ) |>
  dplyr::select(
    question_type,
    n_cells,
    total_series,
    median_q,
    median_mean_abs_dlambda
  ) |>
  rename(
    `Question type`              = question_type,
    `# Cells (industry × type)`  = n_cells,
    `# Series`                   = total_series,
    `Median q (RW variance)`     = median_q,
    `Median |Δλ|`                = median_mean_abs_dlambda
  )

print(instability_by_qtype)
```

The stability of factor loadings differs across question types: current assessments are the most stable, expectations and price-related questions exhibit moderate time variation, while month-on-month indicators display the highest degree of instability.

```{r }
plot_factor_block_all_industries <- function(
  Lambda_cell_tv, dates,
  factor_j = 1, block = "B",
  event_years = c(2008, 2011, 2020),
  top_k = 20,
  rank_by = c("instability", "mean_abs"),
  show_legend = FALSE
) {
  rank_by <- match.arg(rank_by)

  # parse cell id robustly: "C1010000_D" or "C1010000__D"
  parse_cell <- function(cell) {
    parts <- strsplit(cell, "__", fixed = TRUE)[[1]]
    if (length(parts) == 1) parts <- strsplit(cell, "_", fixed = TRUE)[[1]]
    list(industry = parts[1], blk = parts[2])
  }

  cells <- names(Lambda_cell_tv)

  meta <- tibble(cell = cells) |>
    mutate(tmp = lapply(cell, parse_cell),
           industry = purrr::map_chr(tmp, "industry"),
           blk      = purrr::map_chr(tmp, "blk")) |>
    dplyr::select(-tmp) |>
    dplyr::filter(blk == block)

  stopifnot(nrow(meta) > 0)

  # long df: one line per industry in this block
  df <- purrr::map_dfr(meta$cell, function(c_id) {
    L <- Lambda_cell_tv[[c_id]]
    tibble(
      date = dates,
      industry = meta$industry[meta$cell == c_id][1],
      loading = as.numeric(L[, factor_j])
    )
  }) |>
    dplyr::arrange(industry, date)

  # rank industries within this (factor, block)
  score <- df |>
    dplyr::group_by(industry) |>
    dplyr::summarise(
      mean_abs    = mean(abs(loading), na.rm = TRUE),
      instability = mean(abs(diff(loading)), na.rm = TRUE),
      .groups = "drop"
    )

  top_inds <- score |>
    dplyr::arrange(dplyr::desc(.data[[rank_by]])) |>
    dplyr::slice_head(n = top_k) |>
    dplyr::pull(industry)

  # keep only top-k industries
  df <- df |>
    dplyr::filter(industry %in% top_inds)

  # events
  event_dates <- as.Date(paste0(event_years, "-01-01"))
  keep_evt <- event_dates >= min(dates) & event_dates <= max(dates)
  event_dates <- event_dates[keep_evt]
  event_labels <- as.character(event_years[keep_evt])

  p <- ggplot(df, aes(x = date, y = loading, group = industry, color = industry)) +
    geom_line(linewidth = 0.9, alpha = 0.9) +
    scale_color_viridis_d(option = "turbo") +
    geom_hline(yintercept = 0, linetype = 3, color = "gray60") +
    labs(
      title = paste0("TV Loadings across industries: Factor ", factor_j, ", Block = ", block),
      subtitle = paste0("Showing top-", top_k, " industries by ", rank_by),
      x = NULL,
      y = paste0("Loading (lambda_", factor_j, ")"),
      color = "Industry"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = if (show_legend) "bottom" else "none") +
    guides(color = guide_legend(ncol = 3))

  if (length(event_dates) > 0) {
    p <- p +
      geom_vline(xintercept = event_dates, linetype = 2, color = "red3", linewidth = 0.8) +
      annotate("text", x = event_dates, y = Inf, label = event_labels,
               vjust = 1.2, hjust = 0, color = "red3", size = 3) +
      coord_cartesian(clip = "off")
  }

  p
}
```


```{r fig.width=13, fig.height=6}
for (j in 1:5) {

  print(plot_factor_block_all_industries(
    Lambda_cell_tv, dates,
    factor_j = j, block = "D",
    top_k = 20, rank_by = "instability",
    show_legend = TRUE
  ))

  print(plot_factor_block_all_industries(
    Lambda_cell_tv, dates,
    factor_j = j, block = "B",
    top_k = 20, rank_by = "instability",
    show_legend = TRUE
  ))
}
```

At the Level-3 disaggregation, cross-sectional heterogeneity in time-varying loadings becomes substantially more pronounced. While the overall block- and factor-level patterns remain consistent with Level 2, the finer industry classification reveals strong and persistent sectoral divergence, particularly in high-frequency (Block D) indicators. Crisis episodes, especially 2020, coincide with structural reallocation rather than temporary dispersion, suggesting that macro shocks alter the industry-specific exposure to latent drivers in a lasting manner.