---
title: "03_supplement_TVFL"
author: "Yiming Zhao"
date: "2026-02-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)

# --- Preconditions: objects created in 02_supplement_Dynamic_DFM.Rmd ---
dfm02 <- readRDS("output/dfm_results_02_level2.rds")
source("ifo_data_processing.R")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered

library(KFAS)
library(tidyverse)

# Factors used as observed regressors (from Section 02)
F_use <- as.matrix(F_smooth)  # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(Y_kfas)

stopifnot(ncol(F_use) == r, nrow(Y_kfas) == T_obs, length(Hdiag) == N_obs)
```

## Time-Varying Factor Loadings (Two-step TVP Regression)

We extend the measurement equation by letting factor loadings vary over time. We impose a parsimonious restriction that all survey variables within the same industry group under each level share the same (possibly time-varying) factor loading vector.


## EXtension 1: Level 2 only

```{r }
# Keep only Level 2
keep_vars <- var_metadata$level == 2

stopifnot("industry_code" %in% colnames(var_metadata_filtered))
stopifnot("level" %in% colnames(var_metadata_filtered))

group_id <- as.character(var_metadata_filtered$industry_code)
groups <- sort(unique(group_id))
G <- length(groups)

cat("Number of Level-2 groups (industry_code):", G, "\n")
```


```{r }
# For group g with indices idx (n_g variables):

# y_t (n_g x 1) = Z_t (n_g x r) * beta_t (r x 1) + eps_t

# Z_t = 1_{n_g} %*% t(F_t)  -> each row is F_t'

# beta_t = beta_{t-1} + u_t,  u_t ~ N(0, q * I_r)

#

# We estimate scalar q_g via MLE (fitSSM), keep H_g fixed as diag(Hdiag[idx]).

fit_tvp_group <- function(Y_group, Fmat, H_vec, q_init = 1e-4) {

# Y_group: T x n_g (NA allowed)

# Fmat:    T x r

# H_vec:   n_g vector (fixed obs variances)

T_obs <- nrow(Fmat)
r <- ncol(Fmat)
n_g <- ncol(Y_group)

# Build time-varying Z: array dim (n_g, r, T)

Z_arr <- array(NA_real_, dim = c(n_g, r, T_obs))
for (t in 1:T_obs) {
Z_arr[,,t] <- matrix(rep(Fmat[t, ], each = n_g), nrow = n_g, byrow = FALSE)
# each row = F_t'
}

# Custom state-space:

# alpha_t = beta_t (dimension r)

# y_t = Z_t alpha_t + eps_t

# alpha_{t+1} = alpha_t + eta_t  (RW)

model0 <- SSModel(
Y_group ~ -1 +
SSMcustom(
Z = Z_arr,
T = diag(r),
R = diag(r),
Q = diag(NA_real_, r),   # q * I_r to be set in updatefn
a1 = rep(0, r),
P1 = diag(10, r)
),
H = diag(H_vec, n_g)        # fixed diagonal obs noise per series
)

updatefn <- function(pars, model) {
q <- exp(pars[1])           # enforce positivity
model$Q[,,1] <- diag(q, r)  # q * I_r
model
}

fit <- fitSSM(model0, inits = log(q_init), updatefn = updatefn, method = "BFGS")
model <- fit$model
q_hat <- exp(fit$optim.out$par[1])

kfs <- KFS(model, smoothing = c("state"))
beta_t <- kfs$alphahat        # T x r

colnames(beta_t) <- paste0("lambda_", 1:r)
list(beta_t = beta_t, q_hat = q_hat)
}

# --- Estimate group-level TV loadings for all groups ---

Lambda_g_tv <- vector("list", G)
q_hat_g <- rep(NA_real_, G)
names(Lambda_g_tv) <- groups
names(q_hat_g) <- groups

for (gg in seq_along(groups)) {
g <- groups[gg]
idx <- which(group_id == g)
n_g <- length(idx)

Y_group <- as.matrix(Y_kfas[, idx, drop = FALSE])   # T x n_g
H_vec <- Hdiag[idx]

res <- fit_tvp_group(Y_group = Y_group, Fmat = F_use, H_vec = H_vec, q_init = 1e-4)

Lambda_g_tv[[g]] <- res$beta_t   # T x r
q_hat_g[g] <- res$q_hat

if (gg %% 10 == 0) cat("Done group", gg, "/", G, "\n")
}

summary(q_hat_g)

# --- Expand to variable-level loadings (each variable inherits its group's path) ---

Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

for (i in 1:N_obs) {
g <- group_id[i]
Lambda_tv_list[[i]] <- Lambda_g_tv[[g]]  # T x r, shared within group
}
```
The estimated loading innovation variance Q is quantitatively very small, indicating that time variation in the factor loadings is limited.

```{r}
# --- Plot: one group's loading paths (pick a group with many vars, or specify) ---

plot_group_loadings <- function(g) {
beta_t <- Lambda_g_tv[[g]]
df <- data.frame(date = dates, beta_t)
df_long <- tidyr::pivot_longer(df, cols = starts_with("lambda_"),
names_to = "factor", values_to = "loading")

ggplot(df_long, aes(x = date, y = loading, group = factor)) +
geom_line() +
facet_wrap(~ factor, ncol = 1, scales = "free_y") +
labs(title = paste("Group-level time-varying loadings (industry_code):", g),
x = NULL, y = "Loading") +
theme_minimal()
}

# choose a group example (largest group)

group_sizes <- table(group_id)
g_example <- names(which.max(group_sizes))
print(plot_group_loadings(g_example))

```
Repeating the analysis at finer industry aggregation levels yields the same qualitative pattern: although disaggregation slightly increases cross-sectional heterogeneity, the overall factor structure remains remarkably stable. This suggests that the underlying economic relationships captured by the common factors are structurally persistent rather than regime-dependent.



## EXtension 2: Level 3 only

```{r }
# --- Preconditions: objects created in 02_supplement_Dynamic_DFM.Rmd ---
dfm02 <- readRDS("output/dfm_results_02_level3.rds")
source("ifo_data_processing.R")

Y_kfas <- dfm02$Y_kfas
F_smooth <- dfm02$F_smooth
Hdiag <- dfm02$Hdiag
r <- dfm02$r
dates <- dfm02$dates
var_names_filtered <- dfm02$var_names_filtered
var_metadata_filtered <- dfm02$var_metadata_filtered

library(KFAS)
library(tidyverse)

# Factors used as observed regressors (from Section 02)
F_use <- as.matrix(F_smooth)  # T x r
T_obs <- nrow(F_use)
N_obs <- ncol(Y_kfas)

stopifnot(ncol(F_use) == r, nrow(Y_kfas) == T_obs, length(Hdiag) == N_obs)
```

```{r}
# Keep only Level 3
keep_vars <- var_metadata$level == 3

stopifnot("industry_code" %in% colnames(var_metadata_filtered))
stopifnot("level" %in% colnames(var_metadata_filtered))

group_id <- as.character(var_metadata_filtered$industry_code)
groups <- sort(unique(group_id))
G <- length(groups)

cat("Number of Level-3 groups (industry_code):", G, "\n")
```

```{r }
# For group g with indices idx (n_g variables):

# y_t (n_g x 1) = Z_t (n_g x r) * beta_t (r x 1) + eps_t

# Z_t = 1_{n_g} %*% t(F_t)  -> each row is F_t'

# beta_t = beta_{t-1} + u_t,  u_t ~ N(0, q * I_r)

#

# We estimate scalar q_g via MLE (fitSSM), keep H_g fixed as diag(Hdiag[idx]).

fit_tvp_group <- function(Y_group, Fmat, H_vec, q_init = 1e-4) {

# Y_group: T x n_g (NA allowed)

# Fmat:    T x r

# H_vec:   n_g vector (fixed obs variances)

T_obs <- nrow(Fmat)
r <- ncol(Fmat)
n_g <- ncol(Y_group)

# Build time-varying Z: array dim (n_g, r, T)

Z_arr <- array(NA_real_, dim = c(n_g, r, T_obs))
for (t in 1:T_obs) {
Z_arr[,,t] <- matrix(rep(Fmat[t, ], each = n_g), nrow = n_g, byrow = FALSE)
# each row = F_t'
}

# Custom state-space:

# alpha_t = beta_t (dimension r)

# y_t = Z_t alpha_t + eps_t

# alpha_{t+1} = alpha_t + eta_t  (RW)

model0 <- SSModel(
Y_group ~ -1 +
SSMcustom(
Z = Z_arr,
T = diag(r),
R = diag(r),
Q = diag(NA_real_, r),   # q * I_r to be set in updatefn
a1 = rep(0, r),
P1 = diag(10, r)
),
H = diag(H_vec, n_g)        # fixed diagonal obs noise per series
)

updatefn <- function(pars, model) {
q <- exp(pars[1])           # enforce positivity
model$Q[,,1] <- diag(q, r)  # q * I_r
model
}

fit <- fitSSM(model0, inits = log(q_init), updatefn = updatefn, method = "BFGS")
model <- fit$model
q_hat <- exp(fit$optim.out$par[1])

kfs <- KFS(model, smoothing = c("state"))
beta_t <- kfs$alphahat        # T x r

colnames(beta_t) <- paste0("lambda_", 1:r)
list(beta_t = beta_t, q_hat = q_hat)
}

# --- Estimate group-level TV loadings for all groups ---

Lambda_g_tv <- vector("list", G)
q_hat_g <- rep(NA_real_, G)
names(Lambda_g_tv) <- groups
names(q_hat_g) <- groups

for (gg in seq_along(groups)) {
g <- groups[gg]
idx <- which(group_id == g)
n_g <- length(idx)

Y_group <- as.matrix(Y_kfas[, idx, drop = FALSE])   # T x n_g
H_vec <- Hdiag[idx]

res <- fit_tvp_group(Y_group = Y_group, Fmat = F_use, H_vec = H_vec, q_init = 1e-4)

Lambda_g_tv[[g]] <- res$beta_t   # T x r
q_hat_g[g] <- res$q_hat

if (gg %% 10 == 0) cat("Done group", gg, "/", G, "\n")
}

summary(q_hat_g)

# --- Expand to variable-level loadings (each variable inherits its group's path) ---

Lambda_tv_list <- vector("list", N_obs)
names(Lambda_tv_list) <- var_names_filtered

for (i in 1:N_obs) {
g <- group_id[i]
Lambda_tv_list[[i]] <- Lambda_g_tv[[g]]  # T x r, shared within group
}
```
The estimated loading innovation variance Q is quantitatively very small, indicating that time variation in the factor loadings is limited.

```{r}
# --- Plot: one group's loading paths (pick a group with many vars, or specify) ---

plot_group_loadings <- function(g) {
beta_t <- Lambda_g_tv[[g]]
df <- data.frame(date = dates, beta_t)
df_long <- tidyr::pivot_longer(df, cols = starts_with("lambda_"),
names_to = "factor", values_to = "loading")

ggplot(df_long, aes(x = date, y = loading, group = factor)) +
geom_line() +
facet_wrap(~ factor, ncol = 1, scales = "free_y") +
labs(title = paste("Group-level time-varying loadings (industry_code):", g),
x = NULL, y = "Loading") +
theme_minimal()
}

# choose a group example (largest group)

group_sizes <- table(group_id)
g_example <- names(which.max(group_sizes))
print(plot_group_loadings(g_example))

```

Repeating the analysis at finer industry aggregation levels yields the same qualitative pattern: although disaggregation slightly increases cross-sectional heterogeneity, the overall factor structure remains remarkably stable. This suggests that the underlying economic relationships captured by the common factors are structurally persistent rather than regime-dependent.