---
title: "04.1 Incremental Predictive Power Analysis"
author: "Weiyun Hong"
date: "2026-01-03"
header-includes:
  - \usepackage{parskip}
  - \setlength{\parskip}{0.5em}
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE,
  fig.width = 12, 
  fig.align = "center"
)
```


In Section 03, we established that the measurement relationship between observed survey variables and latent factors exhibits significant time variation. The loadings $\lambda_{g,b,t}$ describe how each industry *reflects* the common factor (measurement direction: $F_t \to X_{i,t}$).

This section addresses the reverse question: which industry-block combinations can *predict* future factor movements? This identifies **leading indicators** that provide advance signals of aggregate cycle turning points.

| Aspect | Time-Varying Loadings (03) | Predictive Power (04) |
|--------|---------------------------|----------------------|
| Direction | $F_t \to X_{i,t}$ | $\bar{x}_{c,t-\ell} \to F_{j,t}$ |
| Question | How does industry reflect the cycle? | Which industries act as leading indicators for the factor? |
| Coefficient | $\lambda$ (loading) | $\beta$ (regression) |

```{r load-packages}
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)

source("ifo_data_processing.R")
theme_set(theme_minimal())
```

```{r load-data}
# Load saved state from Section 03
dfm_state <- readRDS("output/dfm_results_02.rds")
dfm_state <- readRDS("output/dfm_state_04.rds")

dates      <- dfm_state$dates
r          <- dfm_state$r
F_smooth   <- dfm_state$F_smooth    # T x r
X_kfas     <- dfm_state$X_kfas      # T x N
var_map    <- dfm_state$var_map     # N rows

T_obs <- nrow(F_smooth)
N_obs <- ncol(X_kfas)

# Cell structure
cells <- sort(unique(var_map$cell_id))
K <- length(cells)

cell_meta <- data.frame(cell_id = cells, stringsAsFactors = FALSE) |>
  mutate(
    industry = sub("_[ABCD]$", "", cell_id),
    block = sub("^.*_", "", cell_id),
    n_series = sapply(cell_id, function(c) sum(var_map$cell_id == c))
  )

cat("Sample: ", as.character(min(dates)), " to ", as.character(max(dates)), 
    " (T=", T_obs, ", N=", N_obs, ", K=", K, " cells)\n", sep = "")
```


### Candidate Predictor Construction

For cell $c = (g, b)$ containing $n_c$ individual series, the candidate predictor is:
\begin{equation*}
\bar{x}_{c,t} = \frac{1}{n_c} \sum_{i \in c} X_{i,t}
\end{equation*}

This aggregation reduces idiosyncratic noise while preserving cell-specific signals.

```{r construct-cell-averages}
X_cell <- matrix(NA_real_, nrow = T_obs, ncol = K)
colnames(X_cell) <- cells

for (k in seq_along(cells)) {
  idx <- which(var_map$cell_id == cells[k])
  X_cell[, k] <- if (length(idx) == 1) X_kfas[, idx] else rowMeans(X_kfas[, idx], na.rm = TRUE)
}
```

### Granger Causality Test

We test whether cell $c$ provides **incremental predictive power** for factor $F_j$ beyond the factor's own history.

**Restricted Model** (baseline):
\begin{equation*}
F_{j,t} = \alpha_0 + \sum_{k=1}^{p} \phi_k F_{j,t-k} + \varepsilon_t^{(R)}
\end{equation*}

**Unrestricted Model** (adding candidate):
\begin{equation*}
F_{j,t} = \alpha_0 + \sum_{k=1}^{p} \phi_k F_{j,t-k} + \sum_{\ell=1}^{q} \beta_\ell \bar{x}_{c,t-\ell} + \varepsilon_t^{(U)}
\end{equation*}

The null hypothesis $H_0: \beta_1 = \cdots = \beta_q = 0$ is tested via F-statistic:
\begin{equation*}
F = \frac{(SSE_R - SSE_U)/q}{SSE_U/(T-p-q-1)} \sim F(q, T-p-q-1)
\end{equation*}

**Model Fit Measures**: The coefficient of determination $R^2$ quantifies explained variance:
\begin{align}
R^2_R &= 1 - \frac{SSE_R}{SST}, \quad \text{(restricted model)} \\
R^2_U &= 1 - \frac{SSE_U}{SST}, \quad \text{(unrestricted model)}
\end{align}
where $SST = \sum_t (F_{j,t} - \bar{F}_j)^2$ is the total sum of squares.

**Incremental $R^2$** measures additional variance explained by the cell's lagged values:
\begin{equation*}
\Delta R^2 = R^2_U - R^2_R = \frac{SSE_R - SSE_U}{SST}
\end{equation*}

Information Gain: A high $\Delta R^2$ indicates that the industry cell provides significant unique information that is not already captured by the factor's own historical price action.

**Sign Interpretation**: The sum $\sum_\ell \beta_\ell$ indicates the direction of the predictive relationship:

- $\sum \beta > 0$: Increases in cell average predict factor increases (procyclical, e.g., business expectations)
- $\sum \beta < 0$: Cell increases predict factor decreases (countercyclical, e.g., cost pressures, inventorybuildup)

Both types are valid leading indicators. Countercyclical indicators (e.g., inventory buildup, cost pressures) can be equally informative for forecasting.

The autoregressive order is set to $p=2$. The maximum lag length of the candidate predictors is set to $q=5$, allowing for distributed lead effects over a short- to medium-horizon window. 

```{r granger-test-function}
granger_test_factor <- function(F_j, x_c, p = 2, q = 5) {
  T_full <- length(F_j)
  max_lag <- max(p, q)
  T_eff <- T_full - max_lag
  
  y <- F_j[(max_lag + 1):T_full]
  
  # Factor lags
  F_lags <- matrix(NA_real_, nrow = T_eff, ncol = p)
  for (k in 1:p) F_lags[, k] <- F_j[(max_lag + 1 - k):(T_full - k)]
  colnames(F_lags) <- paste0("F_lag", 1:p)
  
  # Predictor lags
  x_lags <- matrix(NA_real_, nrow = T_eff, ncol = q)
  for (ell in 1:q) x_lags[, ell] <- x_c[(max_lag + 1 - ell):(T_full - ell)]
  colnames(x_lags) <- paste0("x_lag", 1:q)
  
  # Models
  fit_R <- lm(y ~ ., data = data.frame(y = y, F_lags))
  fit_U <- lm(y ~ ., data = data.frame(y = y, F_lags, x_lags))
  
  SSE_R <- sum(residuals(fit_R)^2)
  SSE_U <- sum(residuals(fit_U)^2)
  SST <- sum((y - mean(y))^2)
  
  # F-test
  df1 <- q
  df2 <- T_eff - p - q - 1
  F_stat <- ((SSE_R - SSE_U) / df1) / (SSE_U / df2)
  p_value <- pf(F_stat, df1, df2, lower.tail = FALSE)
  
  # Coefficients
  beta_coefs <- coef(fit_U)[grep("^x_lag", names(coef(fit_U)))]
  
  # R-squared values
  R2_R <- summary(fit_R)$r.squared
  R2_U <- summary(fit_U)$r.squared
  
  list(
    F_stat = F_stat,
    p_value = p_value,
    R2_R = R2_R,
    R2_U = R2_U,
    delta_R2 = R2_U - R2_R,
    beta_coefs = beta_coefs,
    beta_sum = sum(beta_coefs),
    T_eff = T_eff
  )
}
```

## Full-Sample Analysis

```{r run-granger-tests, cache=TRUE}
p_lags <- 2
q_lags <- 5

granger_results <- vector("list", r)
names(granger_results) <- paste0("Factor_", 1:r)

for (j in 1:r) {
  results_j <- data.frame(
    cell_id = cells,
    industry = cell_meta$industry,
    block = cell_meta$block,
    F_stat = NA_real_,
    p_value = NA_real_,
    R2_R = NA_real_,
    R2_U = NA_real_,
    delta_R2 = NA_real_,
    beta_sum = NA_real_,
    stringsAsFactors = FALSE
  )
  
  for (k in seq_along(cells)) {
    test <- tryCatch(
      granger_test_factor(F_smooth[, j], X_cell[, k], p = p_lags, q = q_lags),
      error = function(e) NULL
    )
    if (!is.null(test)) {
      results_j$F_stat[k] <- test$F_stat
      results_j$p_value[k] <- test$p_value
      results_j$R2_R[k] <- test$R2_R
      results_j$R2_U[k] <- test$R2_U
      results_j$delta_R2[k] <- test$delta_R2
      results_j$beta_sum[k] <- test$beta_sum
    }
  }
  
  granger_results[[j]] <- results_j
}
```



### Leading Indicator Selection
### Benjamini-Hochberg Correction

With `r K` cells tested per factor across `r r` factors, multiple testing correction is essential. After Benjamini-Hochberg (BH) correction, we identify cells with **statistically significant incremental predictive power** (BH-adjusted $p < 0.05$). The primary metric is $\Delta R^2$, which measures how much additional variance in the factor is explained by the cell's lagged values.

```{r bh-correction, include=FALSE}
# We apply the Benjamini-Hochberg (BH) procedure to control the false discovery rate (FDR) at 5%.
# Apply BH correction within each factor
for (j in 1:r) {
  granger_results[[j]]$p_adj_bh <- p.adjust(granger_results[[j]]$p_value, method = "BH")
}

# Summary of correction impact
bh_summary <- lapply(1:r, function(j) {
  res <- granger_results[[j]]
  data.frame(
    Factor = j,
    Raw_Significant = sum(res$p_value < 0.05, na.rm = TRUE),
    BH_Significant = sum(res$p_adj_bh < 0.05, na.rm = TRUE)
  )
}) |> bind_rows()

bh_summary$Retention_Rate <- round(bh_summary$BH_Significant / bh_summary$Raw_Significant * 100, 1)

kable(bh_summary,
      col.names = c("Factor", "Raw p < 0.05", "BH-adjusted p < 0.05", "Retention (%)"),
      caption = "Impact of Benjamini-Hochberg Correction on Significant Indicators")
```

```{r select-leading-indicators}
# Summary across factors (using BH-corrected p-values)
leading_summary <- lapply(1:r, function(j) {
  res <- granger_results[[j]]
  sig_mask <- res$p_adj_bh < 0.05
  n_sig <- sum(sig_mask, na.rm = TRUE)
  n_pos <- sum(sig_mask & res$beta_sum > 0, na.rm = TRUE)
  n_neg <- sum(sig_mask & res$beta_sum < 0, na.rm = TRUE)
  max_dR2 <- max(res$delta_R2[sig_mask], na.rm = TRUE)
  
  data.frame(
    Factor = j,
    Significant = n_sig,
    Procyclical = n_pos,
    Countercyclical = n_neg,
    Max_Delta_R2 = ifelse(is.finite(max_dR2), max_dR2 * 100, NA)
  )
}) |> bind_rows()

kable(leading_summary, 
      col.names = c("Factor", "Significant (BH)", "Σβ > 0", "Σβ < 0", "Max ΔR² (%)"),
      digits = c(0, 0, 0, 0, 3),
      caption = "Leading Indicator Summary by Factor (BH-corrected p < 0.05)")
```


Factor 2 shows mostly countercyclical relationships (79 significant indicators have $\sum\beta < 0$). This suggests Factor 2 captures a dimension that moves inversely to industry-level improvements, potentially reflecting cost pressures or capacity constraints.


\newpage

### Top Leading Indicators by Factor

```{r top-indicators-table}
# Top 5 per factor by delta_R2
top_indicators <- lapply(1:r, function(j) {
  res <- granger_results[[j]] |>
    filter(p_value < 0.05) |>
    arrange(desc(delta_R2)) |>
    slice_head(n = 5)
  
  if (nrow(res) > 0) {
    res$factor <- j
    res$industry_name <- as.character(sapply(res$industry, function(x) 
      tryCatch(get_industry_name_en(x), error = function(e) as.character(x))))
    res
  } else {
    NULL
  }
}) |> bind_rows()

# Create the formatted table
top_indicators |>
  dplyr::select(factor, industry_name, block, delta_R2, p_value, beta_sum) |>
  mutate(
    delta_R2 = paste0(round(delta_R2 * 100, 3), "%"),
    p_value = format.pval(p_value, digits = 2),
    beta_sum = round(beta_sum, 3)
  ) |>
  kable(
    col.names = c("Factor", "Industry", "Block", "ΔR²", "p-value", "ΣB"),
    caption = "Top 5 Leading Indicators per Factor (by Incremental R²)"
  )
```


### Block-Level Predictive Power

```{r block-analysis, out.width = "100%"}
# Aggregate by block (no sign filter)
all_results <- bind_rows(lapply(1:r, function(j) {
  granger_results[[j]] |> mutate(factor = paste0("Factor ", j))
}))

block_summary <- all_results |>
  filter(p_value < 0.05) |>
  group_by(factor, block) |>
  summarise(
    n_valid = n(),
    mean_delta_R2 = mean(delta_R2, na.rm = TRUE) * 100,
    .groups = "drop"
  )

ggplot(block_summary, aes(x = block, y = mean_delta_R2, fill = block)) +
  geom_col(alpha = 0.85) +
  geom_text(aes(label = paste0("n=", n_valid)), vjust = -0.3, size = 3) +
  facet_wrap(~ factor, nrow = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Mean Predictive Power by Question Block",
    subtitle = "Significant indicators only (p<0.05) | Labels show count",
    x = "Question Block", y = "Mean ΔR² (%)"
  ) +
  theme(legend.position = "none")
```



Block D (month-on-month changes) exhibit the highest average incremental predictive power across all factors, showing their central role in capturing short-horizon dynamics. Expectations (Block B) also provide meaningful forward-looking information. Current assessments (Block A) and price indicators (Block C) display weaker incremental contributions, suggesting that their information content is largely contemporaneous once factor persistence is accounted for. 




```{r top-indicators-plot, out.width = "100%", fig.height=9}
if (nrow(top_indicators) > 0) {
  top_indicators |>
    mutate(
      label = paste0(industry_name, " (", block, ")"),
      sign_label = ifelse(beta_sum > 0, "+", "−")
    ) |>
    ggplot(aes(x = reorder(label, delta_R2), y = delta_R2 * 100, fill = block)) +
    geom_col(alpha = 0.85) +
    geom_text(aes(label = paste0(round(delta_R2 * 100, 2), "% ", sign_label)), 
              hjust = -0.1, size = 3) +
    coord_flip(clip = "off") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
    facet_wrap(~ paste("Factor", factor), scales = "free", ncol = 2) +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = "Top Leading Indicators by Factor",
      subtitle = "Ranked by ΔR² | Sign (+/−) indicates direction of predictive relationship",
      x = NULL, y = "Incremental R² (%)", fill = "Block"
    ) +
    theme(legend.position = "bottom", strip.text = element_text(face = "bold"))
}
```



## Rolling Window Analysis: Time-Varying Predictive Power

The full-sample analysis assumes stable predictive relationships. We now examine whether predictive power varies over the business cycle using rolling 60-month windows.

```{r rolling-function}
rolling_granger_test <- function(F_j, x_c, dates, window_width = 60, p = 2, q = 5) {
  T_full <- length(F_j)
  start_idx <- window_width + max(p, q)
  
  if (start_idx >= T_full) return(NULL)
  
  results <- data.frame(
    date = dates[start_idx:T_full],
    delta_R2 = NA_real_,
    beta_sum = NA_real_
  )
  
  for (t in start_idx:T_full) {
    window_idx <- (t - window_width + 1):t
    test <- tryCatch(
      granger_test_factor(F_j[window_idx], x_c[window_idx], p = p, q = q),
      error = function(e) NULL
    )
    row_idx <- t - start_idx + 1
    if (!is.null(test)) {
      results$delta_R2[row_idx] <- test$delta_R2
      results$beta_sum[row_idx] <- test$beta_sum
    }
  }
  results
}
```

```{r rolling-analysis-all-factors, fig.height=17, out.width = "100%"}
event_dates  <- as.Date(c("2008-09-01", "2011-08-01", "2020-03-01"))
event_labels <- c("2008", "2011", "2020")

# Create rolling analysis plots for ALL 5 factors
rolling_plots <- list()

for (fac in 1:r) {
  # Get top 5 indicators for this factor
  top5_fac <- granger_results[[paste0("Factor_", fac)]] |>
    filter(p_value < 0.05) |>
    arrange(desc(delta_R2)) |>
    slice_head(n = 5) |>
    pull(cell_id)
  
  if (length(top5_fac) > 0) {
    rolling_results_fac <- lapply(top5_fac, function(c_id) {
      k <- which(cells == c_id)
      res <- rolling_granger_test(F_smooth[, fac], X_cell[, k], dates, window_width = 60)
      if (!is.null(res)) {
        res$cell_id <- c_id
        res$industry <- cell_meta$industry[cell_meta$cell_id == c_id]
        res$block <- cell_meta$block[cell_meta$cell_id == c_id]
      }
      res
    }) |> bind_rows()
    
    if (nrow(rolling_results_fac) > 0) {
      rolling_df_fac <- rolling_results_fac |>
        mutate(
          industry_name = as.character(sapply(industry, function(x) 
            tryCatch(get_industry_name_en(x), error = function(e) as.character(x)))),
          label = paste0(industry_name, " (", block, ")")
        )
      
      p_fac <- ggplot(rolling_df_fac, aes(x = date, y = delta_R2 * 100, color = label)) +
        geom_line(linewidth = 0.9, alpha = 0.8) +
        geom_hline(yintercept = 0, linetype = 3, color = "gray50") +
        # event dashed lines + labels
        geom_vline(xintercept = event_dates, linetype = "dashed", color = "red", linewidth = 0.6) + 
            annotate(
            "text", x = event_dates, y = max(rolling_df_fac$delta_R2 * 100, na.rm = TRUE),
             label = event_labels, angle = 90, vjust = -0.3, hjust = 0, color = "red", size = 2) +
        coord_cartesian(clip = "off") +
        scale_color_viridis_d(option = "turbo") +
        labs(
          title = paste0("Factor ", fac),
          x = NULL, y = "ΔR² (%)", color = "Predictor"
        ) +
        theme(legend.position = "bottom", legend.text = element_text(size = 7)) +
        guides(color = guide_legend(ncol = 2))
      
      rolling_plots[[fac]] <- p_fac
    }
  }
}

# Combine all plots
if (length(rolling_plots) > 0) {
  library(patchwork)
  combined_plot <- wrap_plots(rolling_plots, ncol = 1) +
    plot_annotation(
      title = "Time-Varying Predictive Power by Factor",
      subtitle = "Rolling 60-month window | Top 5 leading indicators"
    )
  print(combined_plot)
}
```

Predictive power is not constant over time, it spikes around turning points and recessions.

### Composite Leading Indicator

We construct a composite leading indicator using the top 3 predictors weighted by $\Delta R^2$. The sign of each component's contribution is determined by its $\sum \beta$:

```{r composite-indicator}
# Select top 3 predictors for Factor 1 (no sign filter)
top3_F1 <- granger_results[["Factor_1"]] |>
  filter(p_value < 0.05) |>
  arrange(desc(delta_R2)) |>
  slice_head(n = 3)

if (nrow(top3_F1) >= 1) {
  # Weight by delta_R2
  weights <- top3_F1$delta_R2 / sum(top3_F1$delta_R2)
  # Sign adjustment: if beta_sum < 0, flip the contribution
  signs <- sign(top3_F1$beta_sum)
  
  leading_indicator_F1 <- rep(0, T_obs)
  for (i in 1:nrow(top3_F1)) {
    k <- which(cells == top3_F1$cell_id[i])
    # Multiply by sign so that increases in composite always predict factor increases
    leading_indicator_F1 <- leading_indicator_F1 + weights[i] * signs[i] * X_cell[, k]
  }
  leading_indicator_F1 <- scale(leading_indicator_F1)[, 1]
  
  cat("Composite Leading Indicator Components:\n")
  for (i in 1:nrow(top3_F1)) {
    ind_name <- tryCatch(get_industry_name_en(top3_F1$industry[i]), 
                         error = function(e) top3_F1$industry[i])
    sign_str <- ifelse(signs[i] > 0, "+", "−")
    cat("  ", sign_str, " ", ind_name, " (", top3_F1$block[i], "): ", 
        round(weights[i] * 100, 1), "%\n", sep = "")
  }
}
```

```{r plot-composite, fig.height=5, out.width = "100%"}
if (exists("leading_indicator_F1")) {
  plot_df <- data.frame(
    date = dates,
    Factor_1 = scale(F_smooth[, 1])[, 1],
    Leading_Indicator = leading_indicator_F1
  ) |>
    pivot_longer(cols = -date, names_to = "series", values_to = "value")
  
  ggplot(plot_df, aes(x = date, y = value, color = series)) +
    geom_line(linewidth = 1, alpha = 0.85) +
    geom_hline(yintercept = 0, linetype = 3, color = "gray50") +
    # event dashed lines + labels
    geom_vline(xintercept = event_dates, linetype = "dashed", color = "red", linewidth = 0.6) +
    annotate(
       "text",
        x = event_dates,
        y = max(plot_df$value, na.rm = TRUE),
        label = event_labels,
        angle = 90,
        vjust = -0.3,
        hjust = 0,
        color = "red",
        size = 3
      ) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(5.5, 20, 5.5, 5.5))+
    scale_color_manual(
      values = c("Factor_1" = "steelblue", "Leading_Indicator" = "darkorange"),
      labels = c("Factor 1 (Capital Goods)", "Composite Leading Indicator")
    ) +
    labs(
      title = "Factor 1 vs Composite Leading Indicator",
      x = NULL, y = "Standardized value", color = NULL
    ) +
    theme(legend.position = "bottom")
}
```


The composite indicator systematically leads major downturns in the factor, with pronounced declines preceding the 2008–2009 financial crisis, the European debt crisis, and the COVID-19 shock. It captures early stress signals that are not yet reflected in the latent factor itself. This makes the indicator particularly suitable for driving time-varying regime transition probabilities in the subsequent regime-switching framework.





## Implications for Regime-Switching Models

We consider following transition probability specification in which regime changes depend on lagged leading signals. The signals are summarized by a composite leading indicator $z_{t-1}$, constructed as a weighted combination of the top 10 industry–block predictors ranked by incremental $R^2$. 

\begin{equation*}
P(s_t = 1 | s_{t-1}, z_{t-1}) = \Phi(\gamma_0 + \gamma_1 z_{t-1})
\end{equation*}

where the latent state variable $s_t \in \{0,1\}$ indicates the current business cycle regime, and $\Phi(\cdot)$ is the standard normal CDF. When leading indicators deteriorate, the probability of transitioning to recession regime increases.



```{r save-results}
# Save for Section 05
predictive_power_results <- list(
  created_at = Sys.time(),
  granger_results = granger_results,
  X_cell = X_cell,
  cells = cells,
  cell_meta = cell_meta,
  leading_indicator_F1 = if (exists("leading_indicator_F1")) leading_indicator_F1 else NULL,
  top_indicators = top_indicators
)

saveRDS(predictive_power_results, file = "output/predictive_power_04.rds")
#$P(s_t=1\mid s_{t-1}, \mathbf{x}_{t-1})=\Phi\big(\gamma_0+\boldsymbol{\gamma}'\mathbf{x}_{t-1}\big)$
```






\newpage



### Appendix

```{r}
top_k <- 10     
alpha_sig <- 0.05

leading_indicator <- vector("list", length = 5)
names(leading_indicator) <- paste0("Factor_", 1:5)

top_predictors <- vector("list", length = 5)
names(top_predictors) <- paste0("Factor_", 1:5)

for (j in 1:5) {

  key <- paste0("Factor_", j)

  dfj <- granger_results[[key]] |>
    filter(p_value < alpha_sig) |>
    arrange(desc(delta_R2)) |>
    slice_head(n = top_k)

  top_predictors[[key]] <- dfj

  if (nrow(dfj) >= 1) {

    # weights by delta_R2
    weights <- dfj$delta_R2 / sum(dfj$delta_R2)

    # sign adjustment: ensure increases in composite predict factor increases
    signs <- sign(dfj$beta_sum)
    signs[signs == 0] <- 1  

    z <- rep(0, T_obs)

    for (i in seq_len(nrow(dfj))) {
      k <- which(cells == dfj$cell_id[i])
      if (length(k) != 1) next  
      z <- z + weights[i] * signs[i] * X_cell[, k]
    }

    leading_indicator[[key]] <- scale(z)[, 1]

    cat("\n## Composite Leading Indicator Components for ", key, ":\n", sep = "")
    for (i in seq_len(nrow(dfj))) {
      ind_name <- tryCatch(
        get_industry_name_en(dfj$industry[i]),
        error = function(e) dfj$industry[i]
      )
      sign_str <- ifelse(signs[i] > 0, "+", "−")
      cat("  ", sign_str, " ", ind_name, " (", dfj$block[i], "): ",
          round(weights[i] * 100, 1), "%\n", sep = "")
    }

  } else {
    leading_indicator[[key]] <- rep(NA_real_, T_obs)
    cat("\n## ", key, ": no significant predictors at alpha = ", alpha_sig, "\n", sep = "")
  }
}

leading_indicator_F1 <- leading_indicator[["Factor_1"]]
leading_indicator_F2 <- leading_indicator[["Factor_2"]]
leading_indicator_F3 <- leading_indicator[["Factor_3"]]
leading_indicator_F4 <- leading_indicator[["Factor_4"]]
leading_indicator_F5 <- leading_indicator[["Factor_5"]]

```

```{r plot-composite-all, fig.height=9, out.width="100%"}
library(patchwork)

plot_one_factor <- function(factor_j,
                            F_smooth, dates,
                            leading_indicator,
                            event_dates, event_labels) {

  key <- paste0("Factor_", factor_j)

  # leading_indicator: list (recommended) or fall back to leading_indicator_Fj
  z <- NULL
  if (is.list(leading_indicator) && !is.null(leading_indicator[[key]])) {
    z <- leading_indicator[[key]]
  } else {
    varname <- paste0("leading_indicator_F", factor_j)
    if (exists(varname, inherits = TRUE)) {
      z <- get(varname, inherits = TRUE)
    }
  }

  if (is.null(z) || all(is.na(z))) {
    message("Skip ", key, ": leading indicator not found or all NA.")
    return(NULL)
  }

  plot_df <- data.frame(
    date = dates,
    Factor = scale(F_smooth[, factor_j])[, 1],
    Leading_Indicator = as.numeric(z)
  ) |>
    pivot_longer(cols = -date, names_to = "series", values_to = "value")

  ggplot(plot_df, aes(x = date, y = value, color = series)) +
    geom_line(linewidth = 1, alpha = 0.85) +
    geom_hline(yintercept = 0, linetype = 3, color = "gray50") +
    geom_vline(xintercept = event_dates, linetype = "dashed",
               color = "red", linewidth = 0.6) +
    annotate(
      "text",
      x = event_dates,
      y = max(plot_df$value, na.rm = TRUE),
      label = event_labels,
      angle = 90,
      vjust = -0.3,
      hjust = 0,
      color = "red",
      size = 3
    ) +
    coord_cartesian(clip = "off") +
    theme(plot.margin = margin(5.5, 20, 5.5, 5.5)) +
    scale_color_manual(
      values = c("Factor" = "steelblue", "Leading_Indicator" = "darkorange"),
      labels = c(paste0("Factor ", factor_j), "Composite Leading Indicator")
    ) +
    labs(
      title = paste0("Factor ", factor_j, " vs Composite Leading Indicator"),
      x = NULL, y = "Standardized value", color = NULL
    ) +
    theme(legend.position = "bottom")
}

# generate 5 plots
plots <- lapply(1:5, function(j) {
  plot_one_factor(
    factor_j = j,
    F_smooth = F_smooth,
    dates = dates,
    leading_indicator = leading_indicator,  
    event_dates = event_dates,
    event_labels = event_labels
  )
})

# drop NULLs (in case some are missing)
plots <- Filter(Negate(is.null), plots)

wrap_plots(plots, ncol = 2)
```







