---
title: "06 Factor-Augmented VAR: Out-of-Sample Forecasting Evaluation"
header-includes:
- \usepackage{parskip}
- \setlength{\parskip}{0.5em}
- \usepackage{amsmath}
- \usepackage{booktabs}
- \usepackage{float}
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
  word_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6, fig.pos = "H")
options(knitr.table.format = "latex")
```


We construct a Factor-Augmented Vector Autoregression (FAVAR) to predict the aggregate manufacturing Business Climate Index ($\text{KLD}_{t}$, industry code \texttt{C0000000}). 

We evaluate the FAVAR against a hierarchy of benchmarks to determine whether

(i) any predictability exists beyond a random walk,
(ii) the extracted factors provide incremental forecasting gains over a pure AR benchmark, given that $\text{KLD}_{t}$ is constructed from only two survey dimensions (current situation and expectations),
(iii) such gains (if present) are stable across horizons and regimes or concentrated in specific episodes.

To assess robustness, we consider multiple training/evaluation splits.

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(zoo)
library(knitr)
library(kableExtra)
library(patchwork)

source("ifo_data_processing.R")
theme_set(theme_minimal())
```


## Model Specification

### FAVAR Forecasting Equation

The direct $h$-step-ahead forecasting regression takes the form:
\begin{equation}
y_{t+h} = \alpha^{(h)} + \sum_{k=1}^{p} \beta_k^{(h)} \, y_{t+1-k}
          + \sum_{k=1}^{q} \gamma_k^{(h)\prime} \hat{F}_{t+1-k}
          + \varepsilon_{t+h}^{(h)},
\label{eq:favar}
\end{equation}

where $y_t \equiv \text{KLD}_t$ denotes the aggregate manufacturing Business Climate Index (\texttt{C0000000\_KLD}), $\hat{F}_t = (\hat{F}_{1,t},\ldots,\hat{F}_{r,t})'$ is the $r \times 1$ vector of estimated common factors, $p=2$ is the AR lag order, and $q$ is the factor lag order. We consider both $q=p=2$ (full FAVAR) and $q=1$ (parsimonious FAVAR).

Rather than iterating a one-step-ahead VAR forward, we estimate separate projection regressions for each horizon $h \in \{1, 3, 6, 12\}$, avoiding the compounding of one-step-ahead estimation errors.

The factors entering Equation\eqref{eq:favar} are principal components $\hat{F}_t$ estimated by static PCA on the standardised panel $X_{1:t}$, conditioning only on information available up to time $t$. Static PCA is retained here for computational tractability in recursive estimation, as re-estimating and smoothing a state-space model would substantially increase computational cost and introduce additional tuning choices. Consequently, the results should be interpreted as evidence on whether cross-sectional information in the survey panel adds incremental forecasting value beyond AR.


### Benchmark Models

We compare the FAVAR against the following benchmarks:

\begin{center}
\begin{tabular}{llc}
\toprule
\textbf{Model} & \textbf{Specification} & \textbf{Params} \\
\midrule
Random Walk (RW)
  & $\hat{y}_{t+h|t} = y_t$ & 0 \\[3pt]
AR($p$)
  & $y_{t+h} = \alpha + \sum_{k=1}^{p} \beta_k \, y_{t+1-k} + \varepsilon_{t+h}$
  & 3 \\[3pt]
AR + F1
  & AR($p$) augmented with $q=1$ lag of Factor~1
  & 4 \\[3pt]
FAVAR-P
  & AR($p$) + $q=1$ lag of all $r=5$ factors (parsimonious)
  & 8 \\[3pt]
FAVAR
  & AR($p$) + $q=2$ lags of all $r=5$ factors (full)
  & 13 \\
\bottomrule
\end{tabular}
\end{center}

With an initial training window of approximately 126 months (1997:07--2007:12), the usable sample at $h = 12$ declines to roughly 114--120 observations, depending on lag structure. 
The full FAVAR specification with 13 parameters implies that the model may be susceptible to overfitting in a limited sample.

By contrast, the parsimonious FAVAR-P specification retains only 8 parameters. This reduction in dimensionality lowers estimation variance while preserving most of the relevant forecasting signal.

## Evaluation Metrics

For each model $m$ and horizon $h$, let
$e_{t+h|t}^{(m)} = y_{t+h} - \hat{y}_{t+h|t}^{(m)}$
denote the forecast error. We report:

\vspace{0.5em}

\textbf{Root Mean Squared Forecast Error (RMSFE):}
$$\text{RMSFE}^{(m)}_h = \sqrt{\frac{1}{P} \sum_{t=t_0}^{t_0+P-1}
  \left(e_{t+h|t}^{(m)}\right)^2}$$
where $P$ is the number of evaluation periods.

\vspace{1em}

\textbf{Relative RMSFE:}
$\text{RMSFE}^{(m)}_h / \text{RMSFE}^{(\text{RW})}_h$; values below 1 indicate
improvement over the random walk.

\vspace{1em}

\textbf{Out-of-sample $R^2$:}
$$R^2_{\text{OOS},h} = 1 - \frac{\sum_t (e_{t+h|t}^{(m)})^2}
                              {\sum_t (e_{t+h|t}^{(\text{RW})})^2}$$
Positive values indicate that model $m$ outperforms the random walk.

\vspace{1em}

\textbf{Diebold--Mariano test:}
$$\text{DM} = \frac{\bar{d}}{\hat{\sigma}_d / \sqrt{P}},
\quad d_t = (e_{t}^{(A)})^2 - (e_{t}^{(B)})^2$$
with HAC-robust standard errors (Newey--West, bandwidth $= \lfloor h^{1/3} \rfloor$).
A significantly negative DM statistic indicates that model A produces smaller squared
errors than model B.

\vspace{1em}

\textbf{Cumulative Sum of Squared Forecast Error Differences (CSSED):}
$$\text{CSSED}_t = \sum_{\tau=t_0}^{t}
  \left[(e_{\tau}^{(\text{AR})})^2 - (e_{\tau}^{(m)})^2\right]$$
An upward-sloping CSSED indicates periods where model $m$ outperforms the AR benchmark.


\vspace{1em}

## Data Preparation

```{r load-data, results='asis'}
data_wide <- read_csv("data_b_wide.csv", show_col_types = FALSE)
dates     <- as.Date(data_wide$date)
var_names <- colnames(data_wide)[-1]

y_all <- data_wide$C0000000_KLD
```


```{r prepare-panel, results='asis'}
var_metadata <- data.frame(
  var_name      = var_names,
  industry_code = sapply(var_names, extract_industry_code),
  stringsAsFactors = FALSE
)
var_metadata$level <- sapply(var_metadata$industry_code, get_level)

keep_vars <- var_metadata$level == 1
X_raw_full <- as.matrix(data_wide[, c(FALSE, keep_vars)])
vnames     <- var_names[keep_vars]

kld_idx    <- grepl("_KLD$", vnames)
X_raw_full <- X_raw_full[, !kld_idx]
vnames     <- vnames[!kld_idx]

pharma_idx <- grepl("^C2100000_", vnames)
X_raw_full <- X_raw_full[, !pharma_idx]
vnames     <- vnames[!pharma_idx]
```

The dataset comprises `r nrow(data_wide)` months (`r format(min(dates))` to `r format(max(dates))`) with `r ncol(X_raw_full)` variables in the factor extraction panel (21 industries $\times$ 14 questions = 294 cells, excluding KLD and Pharmaceuticals).
The target variable is the aggregate manufacturing Business Climate Index (C0000000\_KLD), with mean `r round(mean(y_all, na.rm=TRUE), 1)` and SD `r round(sd(y_all, na.rm=TRUE), 1)`.


## Recursive Out-of-Sample Estimation

The evaluation window starts at 2008:01, providing an initial training sample of approximately 126 months (1997:07--2007:12). At each subsequent month $t$, the entire estimation pipeline is re-run using only data available up to $t$.

```{r Helper function}
# Helper function: build direct h-step regression matrix
# s ranges from s_min to s_max; Y = y_{s+h}, X = [1, y_{s}, y_{s-1}, ..., F_{s}, F_{s-1}, ...]
build_direct_reg <- function(y_t, F_mat, h, p_ar, p_fac, T_t) {
  # p_fac: number of lags for factors
  s_min <- max(p_ar, p_fac)
  s_max <- T_t - h
  if (s_max <= s_min + 2) return(NULL)

  n_obs <- s_max - s_min + 1
  Y_h   <- y_t[(s_min + h):(s_max + h)]

  r_fac <- if (!is.null(F_mat)) ncol(F_mat) else 0
  n_reg <- 1 + p_ar + p_fac * r_fac
  X_h   <- matrix(NA_real_, n_obs, n_reg)
  X_h[, 1] <- 1
  for (k in 1:p_ar) {
    X_h[, 1 + k] <- y_t[(s_min + 1 - k):(s_max + 1 - k)]
  }
  if (r_fac > 0 && p_fac > 0) {
    for (k in 1:p_fac) {
      for (j in 1:r_fac) {
        col_idx <- 1 + p_ar + (k - 1) * r_fac + j
        X_h[, col_idx] <- F_mat[(s_min + 1 - k):(s_max + 1 - k), j]
      }
    }
  }
  list(Y = Y_h, X = X_h, n_reg = n_reg)
}

# Build prediction vector for time T_t
build_xnew <- function(y_t, F_mat, p_ar, p_fac, T_t) {
  r_fac <- if (!is.null(F_mat)) ncol(F_mat) else 0
  n_reg <- 1 + p_ar + p_fac * r_fac
  x_new <- numeric(n_reg)
  x_new[1] <- 1
  for (k in 1:p_ar) x_new[1 + k] <- y_t[T_t + 1 - k]
  if (r_fac > 0 && p_fac > 0) {
    for (k in 1:p_fac) {
      for (j in 1:r_fac) {
        x_new[1 + p_ar + (k - 1) * r_fac + j] <- F_mat[T_t + 1 - k, j]
      }
    }
  }
  x_new
}
```



```{r run-oos-fn}
# Main recursive OOS evaluation function
run_oos_evaluation <- function(eval_start_date, X_raw_full, y_all, dates,
                                r = 5, p_ar = 2,
                                horizons = c(1, 3, 6, 12)) {

  T_total  <- length(dates)
  max_h    <- max(horizons)
  eval_idx <- which(dates >= eval_start_date)
  eval_idx <- eval_idx[eval_idx + max_h <= T_total]
  n_eval   <- length(eval_idx)
  if (n_eval < 10) return(NULL)

  # Models: RW, AR, AR_F1(q=1), FAVAR_P(q=1), FAVAR(q=2)
  model_names <- c("RW", "AR", "AR_F1", "FAVAR_P", "FAVAR")
  n_models    <- length(model_names)
  n_horizons  <- length(horizons)

  fcst   <- array(NA_real_, dim = c(n_eval, n_horizons, n_models),
                  dimnames = list(NULL, paste0("h", horizons), model_names))
  actual <- matrix(NA_real_, nrow = n_eval, ncol = n_horizons,
                   dimnames = list(NULL, paste0("h", horizons)))

  exp_start <- as.Date("2005-01-01")
  exp_end   <- as.Date("2007-12-01")

  for (i in seq_len(n_eval)) {
    t_origin <- eval_idx[i]
    X_t      <- X_raw_full[1:t_origin, ]
    y_t      <- y_all[1:t_origin]
    dates_t  <- dates[1:t_origin]
    T_t      <- length(y_t)

    for (hi in seq_along(horizons)) {
      actual[i, hi] <- y_all[t_origin + horizons[hi]]
    }

    # (1) Random Walk
    for (hi in seq_along(horizons)) fcst[i, hi, "RW"] <- y_t[T_t]

    # (2) AR(p) - direct h-step
    if (T_t > p_ar + max_h) {
      for (hi in seq_along(horizons)) {
        h   <- horizons[hi]
        reg <- build_direct_reg(y_t, NULL, h, p_ar, 0, T_t)
        if (is.null(reg)) next
        valid <- complete.cases(reg$Y, reg$X)
        if (sum(valid) > reg$n_reg + 2) {
          fit <- lm.fit(reg$X[valid, , drop=FALSE], reg$Y[valid])
          x_new <- build_xnew(y_t, NULL, p_ar, 0, T_t)
          fcst[i, hi, "AR"] <- sum(fit$coefficients * x_new)
        }
      }
    }

    # Factor extraction: recursive PCA
    X_filled_t <- apply(X_t, 2, function(x) {
      if (all(is.na(x))) return(x); zoo::na.approx(x, rule = 2)
    })
    mu_t <- colMeans(X_filled_t, na.rm = TRUE)
    sd_t <- apply(X_filled_t, 2, sd, na.rm = TRUE)
    sd_t[sd_t < 1e-10 | is.na(sd_t)] <- 1
    X_std_t <- sweep(sweep(X_filled_t, 2, mu_t, "-"), 2, sd_t, "/")

    valid_cols <- apply(X_std_t, 2, function(x) all(is.finite(x)) && sd(x) > 1e-10)
    X_std_t    <- X_std_t[, valid_cols, drop = FALSE]

    pca_t   <- prcomp(X_std_t, rank. = r, center = FALSE, scale. = FALSE)
    F_pca_t <- pca_t$x

    exp_idx_t <- dates_t >= exp_start & dates_t <= exp_end
    if (sum(exp_idx_t) > 0) {
      for (j in 1:r) {
        if (mean(F_pca_t[exp_idx_t, j], na.rm = TRUE) < 0) {
          F_pca_t[, j] <- -F_pca_t[, j]
        }
      }
    }
    colnames(F_pca_t) <- paste0("F", 1:r)

    if (T_t > p_ar + max_h) {

      # (3) AR + F1: p_ar lags of y + 1 lag of F1 only
      F1_mat <- F_pca_t[, 1, drop = FALSE]
      for (hi in seq_along(horizons)) {
        h   <- horizons[hi]
        reg <- build_direct_reg(y_t, F1_mat, h, p_ar, 1, T_t)
        if (is.null(reg)) next
        valid <- complete.cases(reg$Y, reg$X)
        if (sum(valid) > reg$n_reg + 2) {
          fit <- lm.fit(reg$X[valid,,drop=FALSE], reg$Y[valid])
          x_new <- build_xnew(y_t, F1_mat, p_ar, 1, T_t)
          fcst[i, hi, "AR_F1"] <- sum(fit$coefficients * x_new)
        }
      }

      # (4) FAVAR-P: p_ar lags of y + 1 lag of all r factors (parsimonious)
      for (hi in seq_along(horizons)) {
        h   <- horizons[hi]
        reg <- build_direct_reg(y_t, F_pca_t, h, p_ar, 1, T_t)
        if (is.null(reg)) next
        valid <- complete.cases(reg$Y, reg$X)
        if (sum(valid) > reg$n_reg + 2) {
          fit <- lm.fit(reg$X[valid,,drop=FALSE], reg$Y[valid])
          x_new <- build_xnew(y_t, F_pca_t, p_ar, 1, T_t)
          fcst[i, hi, "FAVAR_P"] <- sum(fit$coefficients * x_new)
        }
      }

      # (5) FAVAR: p_ar lags of y + p_ar lags of all r factors (full)
      for (hi in seq_along(horizons)) {
        h   <- horizons[hi]
        reg <- build_direct_reg(y_t, F_pca_t, h, p_ar, p_ar, T_t)
        if (is.null(reg)) next
        valid <- complete.cases(reg$Y, reg$X)
        if (sum(valid) > reg$n_reg + 2) {
          fit <- lm.fit(reg$X[valid,,drop=FALSE], reg$Y[valid])
          x_new <- build_xnew(y_t, F_pca_t, p_ar, p_ar, T_t)
          fcst[i, hi, "FAVAR"] <- sum(fit$coefficients * x_new)
        }
      }
    }
  }

  list(fcst = fcst, actual = actual, eval_idx = eval_idx,
       eval_dates = dates[eval_idx], n_eval = n_eval,
       model_names = model_names, n_models = n_models,
       horizons = horizons, n_horizons = n_horizons)
}
```

```{r run-baseline, cache=TRUE}
horizons <- c(1, 3, 6, 12)
res_base <- run_oos_evaluation(
  eval_start_date = as.Date("2008-01-01"),
  X_raw_full = X_raw_full, y_all = y_all, dates = dates
)

fcst        <- res_base$fcst
actual      <- res_base$actual
eval_idx    <- res_base$eval_idx
eval_dates  <- res_base$eval_dates
n_eval      <- res_base$n_eval
model_names <- res_base$model_names
n_models    <- res_base$n_models
n_horizons  <- res_base$n_horizons
```

The evaluation window runs from `r format(eval_dates[1])` to `r format(eval_dates[n_eval])` (`r n_eval` forecast origins).



## Forecast Evaluation Results

The out-of-sample evaluation compares the predictive accuracy of five model specifications against a Random Walk (RW) benchmark across horizons h=1,3,6,12. 

### RMSFE and Relative RMSFE

\vspace{-1em}

```{r compute-errors}
errors <- array(NA_real_, dim = c(n_eval, n_horizons, n_models),
                dimnames = list(NULL, paste0("h", horizons), model_names))
for (m in 1:n_models) {
  for (hi in 1:n_horizons) {
    errors[, hi, m] <- actual[, hi] - fcst[, hi, m]
  }
}

rmsfe <- matrix(NA_real_, n_models, n_horizons,
                dimnames = list(model_names, paste0("h=", horizons)))
for (m in 1:n_models) {
  for (hi in 1:n_horizons) {
    e <- errors[, hi, m]; v <- !is.na(e)
    if (sum(v) > 5) rmsfe[m, hi] <- sqrt(mean(e[v]^2))
  }
}

rel_rmsfe <- sweep(rmsfe, 2, rmsfe["RW", ], "/")
```

```{r tbl-rmsfe}
kbl(round(rmsfe, 3), booktabs = TRUE,
    caption = "Root Mean Squared Forecast Error by Model and Horizon") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```
\vspace{-1em}
```{r tbl-rel-rmsfe}
kbl(round(rel_rmsfe, 3), booktabs = TRUE,
    caption = "Relative RMSFE (ratio to Random Walk). Values below 1 indicate improvement.") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


### Out-of-Sample $R^2$
\vspace{-1.5em}
```{r tbl-oos-r2}
R2_oos <- matrix(NA_real_, n_models, n_horizons,
                 dimnames = list(model_names, paste0("h=", horizons)))
for (m in 1:n_models) {
  for (hi in 1:n_horizons) {
    e_m  <- errors[, hi, m]; e_rw <- errors[, hi, "RW"]
    v <- !is.na(e_m) & !is.na(e_rw)
    if (sum(v) > 5) R2_oos[m, hi] <- 1 - sum(e_m[v]^2) / sum(e_rw[v]^2)
  }
}

kbl(round(R2_oos * 100, 2), booktabs = TRUE,
    caption = "Out-of-Sample $R^2_{OOS}$ (\\%, relative to Random Walk).") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


### Diebold--Mariano Tests
\vspace{-1.3em}
```{r dm-function}
dm_test_nw <- function(e1, e2, h) {
  d <- e1^2 - e2^2; v <- !is.na(d); d <- d[v]; n <- length(d)
  if (n < 10) return(c(stat = NA, pval = NA))
  d_bar <- mean(d); bw <- max(1, floor(h^(1/3)))
  gamma0 <- var(d) * (n - 1) / n; nw_sum <- gamma0
  d_centered <- d - mean(d)
  for (j in 1:min(bw, n - 1)) {
    w <- 1 - j / (bw + 1)
    nw_sum <- nw_sum + 2*w* sum(d_centered[1:(n-j)] * d_centered[(1+j):n]) / n
  }
  se <- sqrt(nw_sum / n)
  if (se < 1e-10) return(c(stat = NA, pval = NA))
  dm_stat <- d_bar / se
  c(stat = dm_stat, pval = 2 * pt(-abs(dm_stat), df = n - 1))
}
```


```{r dm-tables, results='asis'}
comparisons <- list(
  c("AR",      "RW",      "AR vs RW"),
  c("AR_F1",   "AR",      "AR+F1 vs AR"),
  c("FAVAR_P", "AR",      "FAVAR-P vs AR"),
  c("FAVAR_P", "AR_F1",   "FAVAR-P vs AR+F1"),
  c("FAVAR",   "AR",      "FAVAR vs AR"),
  c("FAVAR",   "FAVAR_P", "FAVAR vs FAVAR-P")
)

rows <- lapply(comparisons, function(comp) {
  stats <- sapply(seq_along(horizons), function(hi) {
    h <- horizons[hi]
    res <- dm_test_nw(errors[, hi, comp[1]], errors[, hi, comp[2]], h)
    stat <- round(res[["stat"]], 2)
    sig <- ifelse(res[["pval"]] < 0.01, "***",
           ifelse(res[["pval"]] < 0.05, "**",
           ifelse(res[["pval"]] < 0.10, "*", "")))
    paste0(stat, sig)
  })
  as.data.frame(t(c(Comparison = comp[3], stats)), stringsAsFactors = FALSE)
})

dm_wide <- bind_rows(rows)
colnames(dm_wide) <- c("Comparison", paste0("h=", horizons))

kbl(dm_wide, booktabs = TRUE, row.names = FALSE,
    caption = "Diebold-Mariano test statistics across forecast horizons. Significance: * $p < 0.10$, ** $p < 0.05$, *** $p < 0.01$. HAC standard errors (Newey-West).",
    label = "dm-combined") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"), font_size = 9) %>%
  add_footnote(c("Negative DM: first model has smaller MSE."), notation = "none")
```

\vspace{-1em}

Across all horizons and comparisons, the AR benchmark dominates factor-augmented models, the full-sample DM tests suggest that factor-augmented variants increase forecast errors relative to AR benchmark. The full FAVAR with two factor lags performs worst among the factor-augmented specifications, which exhibits signs of overfitting. 

Although 126 observations appear sufficient, the effective information content is substantially lower due to the high persistence of the target variable and the strong collinearity among factor lags. As a result, the model estimates become unstable and sensitive to sampling variation. In this setting, the additional estimation noise can outweigh the incremental predictive content of the factors.

Importantly, the aggregate KLD is is calculated as a transformed mean of the balances of the business situation and the expectations. Since the factors are extracted from the same underlying industry panel, there is a structural overlap between the predictors and the target variable. Therefore, the target variable KLD already embeds much of the common cross-sectional variation, so its own lag structure implicitly captures a substantial share of the factor information.

The factor augmentation brings in additional survey dimensions (prices, MoM changes, export orders etc.) that go beyond what KLD directly measures, but these dimensions do not appear to have a stable marginal relationship with the future level of the GUS–GES geometric mean under normal conditions.



### Cumulative Squared Error Differences (CSSED)

\vspace{-1.2em}

```{r fig-cssed, fig.width=10, fig.height=7, fig.cap="CSSED: upward slope = factor model outperforms AR. Shaded regions: GFC (2008--09), EDC(2010), COVID (2020)."}
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
for (hi in seq_along(horizons)) {
  h <- horizons[hi]
  e_ar <- errors[, hi, "AR"]; e_fp <- errors[, hi, "FAVAR_P"]
  e_fv <- errors[, hi, "FAVAR"]; e_f1 <- errors[, hi, "AR_F1"]
  v_fp <- !is.na(e_ar) & !is.na(e_fp)
  v_fv <- !is.na(e_ar) & !is.na(e_fv)
  v_f1 <- !is.na(e_ar) & !is.na(e_f1)

  d_fp <- cumsum(e_ar[v_fp]^2 - e_fp[v_fp]^2)
  d_fv <- cumsum(e_ar[v_fv]^2 - e_fv[v_fv]^2)
  d_f1 <- cumsum(e_ar[v_f1]^2 - e_f1[v_f1]^2)

  ylim <- range(c(d_fp, d_fv, d_f1), na.rm = TRUE)

  plot(eval_dates[v_fp], d_fp, type = "l", col = "dodgerblue3", lwd = 2.5,
       main = sprintf("CSSED: h = %d months", h),
       xlab = "", ylab = "Cumulative SED", ylim = ylim)
  lines(eval_dates[v_fv], d_fv, col = "dodgerblue3", lwd = 1, lty = 3)
  lines(eval_dates[v_f1], d_f1, col = "darkorange", lwd = 1.5, lty = 2)
  abline(h = 0, lty = 3, col = "gray40")
  rect(as.Date("2008-01-01"), ylim[1]*2, as.Date("2009-03-01"), ylim[2]*2,
       col = adjustcolor("gray80", 0.3), border = NA)
  rect(as.Date("2009-12-01"), par("usr")[3], as.Date("2011-04-01"), par("usr")[4],
       col = adjustcolor("gray80", 0.2), border = NA)
  rect(as.Date("2020-01-01"), ylim[1]*2, as.Date("2020-12-01"), ylim[2]*2,
       col = adjustcolor("gray80", 0.3), border = NA)
  legend("bottomleft", legend = c("FAVAR-P vs AR", "FAVAR vs AR", "AR+F1 vs AR"),
         col = c("dodgerblue3", "dodgerblue3", "darkorange"),
         lty = c(1, 3, 2), lwd = c(2.5, 1, 1.5), bty = "n", cex = 0.75)
}
```


The CSSED plots reveal that relative forecast performance is strongly regime-dependent, with clear level shifts around the GFC (2008–09) and the COVID shock (2020). Factor-augmented models display marginal value mainly at cyclical turning points, periods of peak cross-sectional dispersion when the common factors carry genuine incremental information beyond the AR benchmark. 

The CSSED paths exhibit a sharp drop around 2009–10, indicating that the AR benchmark delivered substantially smaller multi-step errors in the immediate post-crisis adjustment. Over 2010–19, the CSSED paths drift upward at shorter horizons, suggesting that factor augmentation delivers modest but persistent gains over the AR benchmark through the mid-cycle expansion.

The CSSED for FAVAR-P shows a clear upward spike around 2020, indicating that the factor model outperformed the AR benchmark during this episode. This suggests that the cross-sectional information embedded in the PCA factors was particularly valuable for medium-horizon forecasting when the pandemic triggered a sharp, broad-based disruption across industries.

FAVAR and FAVAR-P tend to move together, suggesting that a second factor lag contributes little incremental predictive content. So additional factor dynamics are largely redundant and serve mainly to inflate estimation variance without improving forecast accuracy. The AR+F1 specification also fails to deliver robust gains, suggesting that the dominant factor alone provides little standalone forecasting power once the AR dynamics are already accounted for.





```{r fig-forecasts, fig.width=10, fig.height=12, fig.cap="Forecasts vs. Actual across horizons. Shaded: GFC, COVID.", include=FALSE}
### Forecast Comparison Plots

par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
for (hi in seq_along(horizons)) {
  h <- horizons[hi]
  y_range <- range(actual[, hi], fcst[, hi, "AR"], fcst[, hi, "FAVAR_P"], na.rm = TRUE)
  plot(eval_dates, actual[, hi], type = "l", lwd = 2, col = "black",
       main = sprintf("Forecasts vs. Actual (h = %d)", h),
       xlab = "", ylab = "KLD", ylim = y_range)
  rect(as.Date("2008-01-01"), par("usr")[3], as.Date("2009-06-01"), par("usr")[4],
       col = adjustcolor("gray80", 0.2), border = NA)
  rect(as.Date("2020-01-01"), par("usr")[3], as.Date("2020-12-01"), par("usr")[4],
       col = adjustcolor("gray80", 0.2), border = NA)
  lines(eval_dates, fcst[, hi, "AR"], col = "gray50", lwd = 1.2, lty = 2)
  lines(eval_dates, fcst[, hi, "FAVAR_P"], col = "dodgerblue3", lwd = 1.5)
  legend("bottomleft", legend = c("Actual", "AR(2)", "FAVAR-P"),
         col = c("black", "gray50", "dodgerblue3"),
         lty = c(1, 2, 1), lwd = c(2, 1.2, 1.5), bty = "n", cex = 0.8)
}
```





## Sub-sample Analysis

To assess whether forecast performance varies across economic regimes, we split the
evaluation window into three sub-periods.

```{r tbl-subsample, results='asis', include=F}
subperiods <- list(
  "2008-2009 (GFC Shock Phase)" = c(as.Date("2008-01-01"), as.Date("2009-03-01")), #Germany’s Industrial Trough – Q1 2009
  "2010-2012 (European Sovereign Debt Crisis)" = c(as.Date("2010-04-01"), as.Date("2012-09-01")),
  "2014-2019 (Recovery)" = c(as.Date("2014-01-01"), as.Date("2019-12-01")),
  "2020-2024 (COVID+)"   = c(as.Date("2020-01-01"), as.Date("2024-12-01"))
)
sp_counter <- 0
for (sp_name in names(subperiods)) {
  sp_counter <- sp_counter + 1
  sp_range <- subperiods[[sp_name]]
  sp_idx   <- eval_dates >= sp_range[1] & eval_dates <= sp_range[2]
  if (sum(sp_idx) < 5) next

  rmsfe_sp <- matrix(NA_real_, n_models, n_horizons,
                     dimnames = list(model_names, paste0("h=", horizons)))
  for (m in 1:n_models) {
    for (hi in 1:n_horizons) {
      e <- errors[sp_idx, hi, m]; v <- !is.na(e)
      if (sum(v) > 3) rmsfe_sp[m, hi] <- sqrt(mean(e[v]^2))
    }
  }
  rel_sp <- sweep(rmsfe_sp, 2, rmsfe_sp["RW", ], "/")

  cat(sprintf("\n\n\\textbf{%s} (n = %d):\n\\vspace{-1.7em}\n", sp_name, sum(sp_idx)))
  print(
    kbl(round(rel_sp, 3), booktabs = TRUE,
        caption = sprintf("Relative RMSFE: %s", sp_name),
        label = sprintf("sub-%d", sp_counter)) %>%
      kable_styling(full_width = FALSE, latex_options = c("HOLD_position"), font_size = 9)
  )
}
```

```{r tbl-subsample-merged, results='asis'}
subperiods <- list(
  "GFC Shock\n(2008--09)"     = c(as.Date("2008-01-01"), as.Date("2009-03-01")),
  "Euro Debt\n(2010--12)"     = c(as.Date("2010-04-01"), as.Date("2012-09-01")),
  "COVID+\n(2020--24)"        = c(as.Date("2020-01-01"), as.Date("2024-12-01"))
)

model_order  <- c("RW", "AR", "AR_F1", "FAVAR_P", "FAVAR")
model_labels <- c("RW", "AR", "AR+F1", "FAVAR-P", "FAVAR")
horizon_labs <- paste0("h=", horizons)   # c("h=1","h=3","h=6","h=12")

# Collect all relative RMSFEs 
sp_results <- lapply(subperiods, function(sp_range) {
  sp_idx <- eval_dates >= sp_range[1] & eval_dates <= sp_range[2]
  n_sp   <- sum(sp_idx)
  if (n_sp < 5) return(NULL)

  rmsfe_sp <- matrix(NA_real_, n_models, n_horizons,
                     dimnames = list(model_names, horizon_labs))
  for (m in seq_len(n_models)) {
    for (hi in seq_along(horizons)) {
      e <- errors[sp_idx, hi, m]; v <- !is.na(e)
      if (sum(v) > 3) rmsfe_sp[m, hi] <- sqrt(mean(e[v]^2))
    }
  }
  rel_sp <- sweep(rmsfe_sp, 2, rmsfe_sp["RW", ], "/")
  list(rel = rel_sp, n = n_sp)
})
sp_results <- Filter(Negate(is.null), sp_results)

# Build wide merged data frame
# Each period contributes 4 columns (h=1,3,6,12); rows = models
merged_mat <- do.call(cbind, lapply(sp_results, function(x) x$rel))
merged_mat <- merged_mat[model_order, ]   # enforce row order
rownames(merged_mat) <- model_labels

# Period sample sizes for header
n_per_period <- sapply(sp_results, function(x) x$n)
period_header_names <- mapply(function(nm, n) {
  clean_nm <- gsub("\n", " ", nm)
  sprintf("%s (n=%d)", clean_nm, n)
}, names(sp_results), n_per_period)

# Column group sizes (4 horizons per period)
n_periods      <- length(sp_results)
group_widths   <- setNames(rep(4L, n_periods), period_header_names)

# Rename columns to h=1 h=3 h=6 h=12 repeated
colnames(merged_mat) <- rep(horizon_labs, n_periods)

# Format
fmt_cell <- function(x) {
  formatC(round(x, 2), digits = 2, format = "f")
}
merged_fmt <- apply(merged_mat, c(1, 2), fmt_cell)

# Bold FAVAR-P cells where FAVAR-P < AR (column-wise comparison)
ar_row     <- merged_mat["AR",      ]
favarp_row <- merged_mat["FAVAR-P", ]
bold_cols  <- which(!is.na(favarp_row) & !is.na(ar_row) & favarp_row < ar_row)

for (j in bold_cols) {
  merged_fmt["FAVAR-P", j] <- paste0("\\textbf{", merged_fmt["FAVAR-P", j], "}")
}

# Render with kableExtra
kbl(merged_fmt,
    booktabs  = TRUE,
    escape    = FALSE,
    align     = c("l", rep("r", n_periods * 4)),
    caption   = paste0(
      "Relative RMSFE by sub-period and forecast horizon (relative to Random Walk). ",
      "Values below 1 indicate improvement over the RW benchmark. ",
      "The AR row (bold) serves as the primary benchmark for factor-augmented models."
    ),
    label     = "subsample-merged"
) |>
  kable_styling(
    full_width    = TRUE,
    latex_options = c("HOLD_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(
    c(" " = 1, group_widths),
    bold      = TRUE,
    font_size = 9,
    line      = TRUE
  ) |>
  row_spec(0, bold = TRUE) |>
  row_spec(which(model_labels == "AR"), bold = TRUE, hline_after = FALSE) |>
  column_spec(1, bold = FALSE, width = "1.8cm") |>
  footnote(
    general       = "Relative RMSFE = RMSFE(model) / RMSFE(RW). Bold FAVAR-P entries: FAVAR-P outperforms AR benchmark.",
    general_title = "",
    footnote_as_chunk = TRUE,
    escape        = FALSE
  )
```



During the GFC shock phase (2008–09), FAVAR-P performs best with the lowest relative RMSFE across all horizons. Factor-augmented specifications are also competitive with the pure AR at short
horizons during European Sovereign Debt Crisis (2010–12).

The European Sovereign Debt Crisis was characterised by fragmented transmission and policy-driven regime shifts (risk-premia spikes followed by abrupt stabilisation), so that medium- and long-horizon outcomes were dominated by non-linear turning points rather than by smooth extrapolation of the contemporaneous common component. In this environment, factor models tend to extrapolate stress too persistently, leading to large errors at h=12 relative to the more robust AR benchmark. 

During the COVID+ period (2020–2024), forecast performance deteriorates markedly and all models struggle relative to the random walk at short horizons (h=1, 3). This occurs because the pandemic induced abrupt, unprecedented level shifts in KLD that violated the mean-reverting dynamics. The factor-augmented models extracted from the same disrupted cross-section, also exhibit discontinuities that the linear projection cannot accommodate. 

At longer horizons, the AR recovers partially, beating the random walk at h=12 (0.899), indicating some remaining mean reversion at annual horizons. The factor-augmented models outperform both the RW and AR at h=6 and h=12: FAVAR-P reaches 0.934 and 0.721 respectively, and FAVAR 0.927 and 0.700. This aligns with the idea that PCA factors become most informative when a shock is broad-based and highly synchronised across industries, so that cross-sectional co-movement contains genuine signal about the medium-run trajectory of aggregate sentiment.

Overall, the results reveal that the factor information is horizon-dependent and Regime-dependent. During episodes of broad-based disruption or gradual structural divergence across industries, the factors carry genuine predictive content beyond what KLD's own lag structure provides. 



## Robustness: Alternative Training/Evaluation Splits

We re-evaluate using two alternative splits: a post-crisis split starting 2012:01, and a proportional
70/30 split.

```{r tbl-alt-splits, results='asis', cache=TRUE}
split_70pct_idx  <- floor(length(dates) * 0.70)
split_70pct_date <- dates[split_70pct_idx]
split_70pct_name <- sprintf("70/30 split (%s)", format(split_70pct_date, "%Y:%m"))

alt_splits <- list(
  "Baseline (2008:01)"    = as.Date("2008-01-01"),
  "Post-crisis (2012:01)" = as.Date("2012-01-01")
)
alt_splits[[split_70pct_name]] <- split_70pct_date

split_counter <- 0
for (split_name in names(alt_splits)) {
  split_counter <- split_counter + 1
  split_date <- alt_splits[[split_name]]

  if (split_date == as.Date("2008-01-01")) {
    alt_rel <- rel_rmsfe; alt_n <- n_eval
  } else {
    res_alt <- run_oos_evaluation(
      eval_start_date = split_date,
      X_raw_full = X_raw_full, y_all = y_all, dates = dates)
    if (is.null(res_alt)) {
      cat(sprintf("\n\n\\textbf{%s}: insufficient data.\n\n", split_name)); next
    }
    alt_n <- res_alt$n_eval
    alt_err <- array(NA_real_, dim = c(alt_n, n_horizons, n_models))
    for (m in 1:n_models) {
      for (hi in 1:n_horizons) {
        alt_err[, hi, m] <- res_alt$actual[, hi] - res_alt$fcst[, hi, m]
      }
    }
    alt_rmsfe <- matrix(NA_real_, n_models, n_horizons,
                        dimnames = list(model_names, paste0("h=", horizons)))
    for (m in 1:n_models) {
      for (hi in 1:n_horizons) {
        e <- alt_err[, hi, m]; v <- !is.na(e)
        if (sum(v) > 5) alt_rmsfe[m, hi] <- sqrt(mean(e[v]^2))
      }
    }
    alt_rel <- sweep(alt_rmsfe, 2, alt_rmsfe["RW", ], "/")
  }

  train_end <- dates[max(which(dates < split_date))]
  cat(sprintf("\n\n\\textbf{%s} (training to %s, n = %d):\n\\vspace{-1.7em}\n",
              split_name, format(train_end), alt_n))
  print(
    kbl(round(alt_rel, 3), booktabs = TRUE,
        caption = sprintf("Relative RMSFE: %s", split_name),
        label = sprintf("alt-%d", split_counter)) %>%
      kable_styling(full_width = FALSE, latex_options = c("HOLD_position"), font_size = 9)
  )
}
```

The robustness checks suggest that the baseline results are sensitive to the inclusion of crisis episodes. The relative ranking of models is broadly preserved: FAVAR-P consistently outperforms the full FAVAR, and AR outperforms AR+F1. 

Besides, the magnitude of the gains from factor augmentation varies with the composition of the evaluation period. In the post-crisis split (starting 2012) when the GFC is excluded, factor models struggle to deliver clear improvements at short horizons. As removing the GFC also removes the high-dispersion episode in which cross-sectional factor signals were most informative for forecasting aggregate KLD. Under the 70/30 split (starting 2016), where the evaluation period is dominated by the late recovery and COVID episode, short-horizon performance remains worse than RW.

Taken together, these results suggest that the competitiveness of factor-augmented models is closely tied to episodes characterized by broad, synchronized movements across industries. When such high-dispersion environments are included, factor information can add value, particularly at longer horizons.


```{r tbl-summary, include=FALSE}
## Summary
summary_df <- data.frame(
  Horizon    = paste0("h=", horizons),
  Best_Model = apply(rmsfe, 2, function(x) names(which.min(x))),
  Best_RMSFE = round(apply(rmsfe, 2, min, na.rm = TRUE), 2),
  Rel_vs_RW  = round(apply(rel_rmsfe, 2, min, na.rm = TRUE), 3),
  R2_OOS_pct = apply(R2_oos, 2, function(x) round(max(x, na.rm = TRUE) * 100, 1))
)
kbl(summary_df, booktabs = TRUE, row.names = FALSE,
    caption = "Best-performing model per forecast horizon") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```



## Discussion

Overall, the AR benchmark provides a robust baseline on average, but factor-augmented models deliver substantial gains at medium-to-long horizons during major systemic shocks. As KLD’s own lag structure already absorbs much of the common cross-sectional variation. For factor augmentation to improve upon AR, the factors would need to capture predictive content in the cross-sectional dispersion across industries, which could contain information that the aggregate trajectory cannot capture.

For applied business-cycle monitoring, this pattern suggests that in tranquil periods, simple autoregressive models are often sufficient. During systemic shocks, industries diverge sharply in timing and magnitude, creating cross-sectional dispersion that the aggregate KLD cannot reflect. Therefore, the factor structure could provide early signal about the medium-run trajectory of aggregate sentiment around turning points that AR benchmark could miss.

At the same time, the split-sample stability analysis in Section 01 revealed that Factors 3 and 5 exhibit instability in their economic interpretation across recursive estimation windows, deviating from the full-sample loading pattern. Accordingly, we treat the results here as aggregate evidence for the value of cross-sectional information, the individual factors in this section should be interpreted with caution. The subsequent analysis should focus on the more stable and economically interpretable factors, primarily Factors 1, 2 and 4, where the mapping between latent structure and observable industry dynamics is robust enough to support further inference.









