---
title: "Data Preprocessing"
author: "Weiyun Hong"
date: "2025-11-24"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    df_print: kable
subtitle: "ifo Business Survey Data - Two Version Comparison"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 150
)
```


**Two versions prepared**:

- **Version A**: Drop BES & PRS, use all years (1991-2025)
- **Version B**: Keep all questions, start from 1997

---

# 1. Setup and Data Loading

```{r libraries}
library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
library(scales)
library(viridis)

options(scipen = 999)
```


```{r load-data}
source("ifo_data_processing.R")

DATA_PATH <- file.path("Data")

results <- suppressMessages(main(DATA_PATH))

data_long <- results$data_long
```

```{r data-summary, echo=FALSE, results='asis'}
cat("Dimensions:", nrow(data_long), "rows ×", ncol(data_long), "columns\n\n")
cat("Date range:", format(min(data_long$date), "%Y-%m"), "to", 
    format(max(data_long$date), "%Y-%m"), "\n\n")
cat("Industries:", n_distinct(data_long$industry_code), "\n\n")
cat("Questions:", n_distinct(data_long$question), "\n\n")
```

## Data Structure

```{r data-structure, comment=''}
str(data_long)
```


```{r sample-data, echo=FALSE}
data_long %>%
  select(date, industry_code, industry_name, level, 
         question, question_name, value) %>%
  head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "scale_down", "hold_position"))
```

---

# 2. Add Combined Variable Names

##`industrycode_questioncode`

**Examples:**
- `C1000000_KLD` = Business climate for Food industry
- `C2800000_KLD` = Business climate for Machinery industry

```{r add-variable-name}
data_long <- data_long %>%
  mutate(
    variable = paste(industry_code, question, sep = "_")
  )
```

```{r variable-examples, echo=FALSE}
#data_long %>%
#  head(20) %>%
#  kable() %>%
#  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

Total unique variables: 2655

---

# 3. Data Quality Check

```{r overall-missing}
overall_missing <- mean(is.na(data_long$value)) * 100
```

**Overall Statistics:**

- Total observations: **`r format(nrow(data_long), big.mark=",")`**
- Missing observations: **`r format(sum(is.na(data_long$value)), big.mark=",")`**
- Overall missing rate: **`r sprintf("%.2f%%", overall_missing)`**

## Missing by Question

```{r missing-by-question}
missing_by_question <- data_long %>%
  group_by(question, question_name) %>%
  summarise(
    n_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    first_data = min(date[!is.na(value)]),
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct))
```

```{r missing-table, echo=FALSE}
missing_by_question %>%
  mutate(
    missing_pct = sprintf("%.2f%%", missing_pct),
    first_data = format(first_data, "%Y-%m")
  ) %>%
  kable(
    col.names = c("Code", "Question Name", "Observations", 
                  "Missing", "Missing %", "First Data")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(which(missing_by_question$missing_pct > 10), 
           background = "#ffcccc")
```

**Key Findings:**

```{r key-findings, echo=FALSE, results='asis'}
high_missing_questions <- missing_by_question %>%
  filter(missing_pct > 10)

if (nrow(high_missing_questions) > 0) {
  cat("\nQuestions with >10% missing:\n\n")
  for (i in 1:nrow(high_missing_questions)) {
    cat(sprintf("- %s (%s): %.2f%%, first data from %s\n",
                high_missing_questions$question[i],
                high_missing_questions$question_name[i],
                high_missing_questions$missing_pct[i],
                format(high_missing_questions$first_data[i], "%Y-%m")))
  }
} else {
  cat("\nAll questions have <10% missing rate\n")
}
```

## Visualize Missing Pattern

```{r plot-missing-time, fig.width=12, fig.height=6}
missing_over_time <- data_long %>%
  group_by(date, question) %>%
  summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop")

ggplot(missing_over_time, aes(x = date, y = missing_pct, color = question)) +
  geom_line(alpha = 0.7, size = 0.8) +
  geom_vline(xintercept = as.Date("1997-01-01"), 
             linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = as.Date("1997-01-01"), y = 50, 
           label = "1997", color = "red", hjust = -0.2, size = 5) +
  scale_color_viridis_d(option = "turbo") +
  labs(
    title = "Missing Rate Over Time by Question",
    subtitle = "Red line marks 1997 - when all questions become available",
    x = "Date",
    y = "Missing Rate (%)",
    color = "Question"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```

---

# 4. Version Comparison

Defining two data preparation versions:

VERSION A: Drop BES & PRS, Use All Years

- Questions: 13 (exclude BES, PRS)
- Time period: 1991-01 onwards
- Rationale: Maximize time span

VERSION B: Keep All Questions, Start from 1997

- Questions: 15 (all questions)
- Time period: 1997-07onwards
- Rationale: Complete panel, include expectations

```{r}
## Prepare Version A: Drop BES & PRS
data_a <- data_long %>%
  filter(!question %in% c("BES", "PRS"))

## Prepare Version B: All Questions from 1997
data_b <- data_long %>%
  filter(date >= as.Date("1997-07-01"))
```


```{r comparison-table, echo=FALSE}
comparison <- tibble(
  Metric = c("Time Period", "Number of Months", "Number of Questions", 
             "Number of Industries", "Total Variables", "Total Observations", 
             "Missing Rate"),
  `Version A` = c(
    paste(format(min(data_a$date), "%Y-%m"), "to", format(max(data_a$date), "%Y-%m")),
    format(n_distinct(data_a$date), big.mark = ","),
    as.character(n_distinct(data_a$question)),
    as.character(n_distinct(data_a$industry_code)),
    format(n_distinct(data_a$variable), big.mark = ","),
    format(nrow(data_a), big.mark = ","),
    sprintf("%.2f%%", mean(is.na(data_a$value)) * 100)
  ),
  `Version B (Preferred for DFM)` = c(
    paste(format(min(data_b$date), "%Y-%m"), "to", format(max(data_b$date), "%Y-%m")),
    format(n_distinct(data_b$date), big.mark = ","),
    as.character(n_distinct(data_b$question)),
    as.character(n_distinct(data_b$industry_code)),
    format(n_distinct(data_b$variable), big.mark = ","),
    format(nrow(data_b), big.mark = ","),
    sprintf("%.2f%%", mean(is.na(data_b$value)) * 100)
  )
)

comparison %>%
  kable(align = c("l", "c", "c")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, background = "#e6f3ff")
```



## Visual Comparison

```{r comparison-plot, fig.width=12, fig.height=6}
# Prepare data for visualization
comp_data <- tibble(
  Version = rep(c("Version A:\nDrop BES/PRS", 
                   "Version B:\nFrom 1997"), each = 3),
  Metric = rep(c("Questions", "Time Points", "Variables"), 2),
  Value = c(
    n_distinct(data_a$question),
    n_distinct(data_a$date),
    n_distinct(data_a$variable),
    n_distinct(data_b$question),
    n_distinct(data_b$date),
    n_distinct(data_b$variable)
  )
)

ggplot(comp_data, aes(x = Metric, y = Value, fill = Version)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(
    aes(label = comma(Value)),
    position = position_dodge(width = 0.7),
    vjust = -0.5,
    size = 4
  ) +
  scale_fill_manual(
    values = c("#E69F00", "#56B4E9"),
    name = NULL
  ) +
  labs(
    title = "Version Comparison: Key Metrics",
    x = NULL,
    y = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    legend.text = element_text(size = 11)
  )
```




---

# 5. Missing Value Patterns by Version

## Version A: Missing by Question

```{r version-a-missing, echo=FALSE}
data_a %>%
  group_by(question, question_name) %>%
  summarise(
    missing_pct = mean(is.na(value)) * 100,
    first_data = min(date[!is.na(value)]),
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct)) %>%
  mutate(
    missing_pct = sprintf("%.2f%%", missing_pct),
    first_data = format(first_data, "%Y-%m")
  ) %>%
  kable(col.names = c("Code", "Name", "Missing %", "First Data")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Version B: Missing by Question

```{r version-b-missing, echo=FALSE}
data_b %>%
  group_by(question, question_name) %>%
  summarise(
    missing_pct = mean(is.na(value)) * 100,
    first_data = min(date[!is.na(value)]),
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct)) %>%
  mutate(
    missing_pct = sprintf("%.2f%%", missing_pct),
    first_data = format(first_data, "%Y-%m")
  ) %>%
  kable(col.names = c("Code", "Name", "Missing %", "First Data")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

# 6. Convert to Wide Format

## Version A Wide Format

```{r version-a-wide}
data_a_wide <- data_a %>%
  select(date, industry_code, question, value) %>%
  pivot_wider(
    names_from = c(industry_code, question),
    names_sep = "_",
    values_from = value
  )
```

**Wide format dimensions:** `r nrow(data_a_wide)` rows × `r ncol(data_a_wide)` columns


## Version B Wide Format

```{r version-b-wide}
data_b_wide <- data_b %>%
  select(date, industry_code, question, value) %>%
  pivot_wider(
    names_from = c(industry_code, question),
    names_sep = "_",
    values_from = value
  )
```

**Wide format dimensions:** `r nrow(data_b_wide)` rows × `r ncol(data_b_wide)` columns

---

# 7. Data Coverage

```{r visualize-missing, fig.width=14, fig.height=8}
set.seed(123)

sample_vars <- data_b_wide %>%
  select(-date) %>%
  names() %>%
  sample(min(100, length(.)))

data_b_wide %>%
  select(date, all_of(sample_vars)) %>%
  pivot_longer(-date, names_to = "variable", values_to = "value") %>%
  mutate(is_missing = is.na(value)) %>%
  ggplot(aes(x = date, y = variable, fill = is_missing)) +
  geom_tile() +
  scale_fill_manual(
    values = c("FALSE" = "#2166ac", "TRUE" = "#d6604d"),
    labels = c("Data Available", "Missing"),
    name = NULL
  ) +
  labs(
    title = "Data Coverage Pattern - Version B (From 1997)",
    subtitle = "Sample of 100 random variables (industry × question combinations)",
    x = "Date",
    y = "Variable (industrycode_questioncode)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  )
```

## Check for Issues

```{r verify-issues, comment=''}
# Check 1: Any completely missing series?
completely_missing_a <- data_a_wide %>%
  select(-date) %>%
  summarise(across(everything(), ~all(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "all_missing") %>%
  filter(all_missing)

completely_missing_b <- data_b_wide %>%
  select(-date) %>%
  summarise(across(everything(), ~all(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "all_missing") %>%
  filter(all_missing)

cat("Version A - Completely missing series:", nrow(completely_missing_a), "\n")
cat("Version B - Completely missing series:", nrow(completely_missing_b), "\n\n")

# Check 2: Series with >50% missing
high_missing_a <- data_a_wide %>%
  select(-date) %>%
  summarise(across(everything(), ~mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 50) %>%
  arrange(desc(missing_pct))

high_missing_b <- data_b_wide %>%
  select(-date) %>%
  summarise(across(everything(), ~mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 50) %>%
  arrange(desc(missing_pct))

cat("Version A - Series with >50% missing:", nrow(high_missing_a), "\n")
if (nrow(high_missing_a) > 0) {
  print(head(high_missing_a, 10))
}

cat("\n\nVersion B - Series with >50% missing:", nrow(high_missing_b), "\n")
if (nrow(high_missing_b) > 0) {
  print(head(high_missing_b, 10))
}

```

Business Climate Index for Fish processing industry is completely missing.

---

# 8. Summary


**Version A: Drop BES & PRS**

- `r n_distinct(data_a$date)` months of data
- `r n_distinct(data_a$question)` questions
- `r format(n_distinct(data_a$variable), big.mark=",")` unique variables
- `r sprintf("%.2f%%", mean(is.na(data_a$value)) * 100)` missing rate

**Version B: All Questions from 1997**

- `r n_distinct(data_b$date)` months of data
- `r n_distinct(data_b$question)` questions (including BES & PRS)
- `r format(n_distinct(data_b$variable), big.mark=",")` unique variables
- `r sprintf("%.2f%%", mean(is.na(data_b$value)) * 100)` missing rate

## Preference for DFM Analysis

**Use Version B (All questions from 1997)** because:

1. **Complete question set** (15 questions)
   - Includes BES (employment expectations)
   - Includes PRS (production expectations)
   - Both are forward-looking indicators

2. **Balanced panel structure**
   - All questions available for all time points
   - Better for DFM estimation
   - Easier to interpret factors

3. **Sufficient time span**
   - `r n_distinct(data_b$date)` months of monthly data
   - Exceeds literature standards (20-30 years)
   - Covers multiple business cycles

4. **Lower missing rate**
   - Version B: `r sprintf("%.2f%%", mean(is.na(data_b$value)) * 100)`
   - Better than Version A: `r sprintf("%.2f%%", mean(is.na(data_a$value)) * 100)`

5. **Cross-industry analysis**
   - Variable names (industrycode_questioncode) allow
   - Easy identification of industry correlations
   - Example: C1000000_KLD vs C2800000_KLD

---

# Datasets 

```{r export-wide, results='asis'}
## Version A: Long format
# write_csv(
#  data_a,
#  "data_a_long.csv"
#)

cat("data_a_long.csv\n")
cat("  Dimensions:", nrow(data_a), "rows x", 
    ncol(data_a), "columns\n")
cat("  Key column: 'variable' (industrycode_questioncode)\n\n")

# Show sample
cat("Sample data:\n")
data_a %>%
  select(date, variable, industry_name, question_name, value) %>%
  head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


## Version B: Long format 
# write_csv(
#  data_b,
#  "data_b_long.csv"
#)

cat("data_b_long.csv\n")
cat("  Dimensions:", nrow(data_b), "rows x", 
    ncol(data_b), "columns\n")
cat("  Key column: 'variable' (industrycode_questioncode)\n\n")

# Show sample
cat("Sample data:\n")
data_b %>%
  select(date, variable, industry_name, question_name, value) %>%
  head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


## Version A: Wide format
data_a_wide <- data_a %>%
  select(date, variable, value) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  )

# write_csv(
#  data_a_wide,
#  "data_a_wide.csv"
#)

cat("data_a_wide.csv\n")
cat("  Dimensions:", nrow(data_a_wide), "rows x", 
    ncol(data_a_wide), "columns\n\n")
cat("  Format: Each column is a variable (industrycode_questioncode)\n\n")
cat("  Missing rate:", 
    sprintf("%.2f%%", mean(is.na(data_a_wide[-1])) * 100), "\n\n")

## Version B: Wide format
data_b_wide <- data_b %>%
  select(date, variable, value) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  )

#write_csv(
#  data_b_wide,
#  "data_b_wide.csv"
#)


cat("data_b_wide.csv\n")
cat("  Dimensions:", nrow(data_b_wide), "rows x", 
    ncol(data_b_wide), "columns\n\n")
cat("  Format: Each column is a variable (industrycode_questioncode)\n\n")
cat("  Missing rate:", 
    sprintf("%.2f%%", mean(is.na(data_b_wide[-1])) * 100), "\n\n")
```


**Exported Files:**

- `data_a_long.csv`: Version A long format (Drop BES/PRS)
- `data_a_wide.csv`: Version A wide format (Drop BES/PRS)
- `data_b_long.csv`: Version B long format (All questions from 1997) 
- `data_b_wide.csv`: Version B wide format (All questions from 1997) 

---