---
title: "Missing Value Analysis - ifo Business Survey Data"
author: "Weiyun Hong"
date: "2025-11-25"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 14,
  fig.height = 10,
  dpi = 150
)
```


This script provides a comprehensive analysis of missing values in the ifo Business Survey data.

```{r libraries}
library(tidyverse)
library(naniar)
library(viridis)
library(scales)
library(knitr)
library(kableExtra)
```



```{r load-data}
# Make sure you've already run the data processing
source("ifo_data_processing.R")
DATA_PATH <- file.path("Data")
results <- main(DATA_PATH)

# Assuming 'results' object is already in your environment

data_long <- results$data_long
```

# 1. Summary Statistics

## Calculate Missing Value Statistics

```{r calc-stats, include=FALSE}
# By industry-question combination
missing_by_combo <- data_long %>%
  group_by(industry_code, industry_name, level, question, question_name) %>%
  summarise(
    total_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct))

# By industry
missing_by_industry <- data_long %>%
  group_by(industry_code, industry_name, level) %>%
  summarise(
    total_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct))

# By question
missing_by_question <- data_long %>%
  group_by(question, question_name, question_type) %>%
  summarise(
    total_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(missing_pct))

# By level
missing_by_level <- data_long %>%
  group_by(level) %>%
  summarise(
    n_industries = n_distinct(industry_code),
    total_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    .groups = "drop"
  ) %>%
  arrange(level)

# By time
missing_by_time <- data_long %>%
  group_by(date) %>%
  summarise(
    total_obs = n(),
    missing_count = sum(is.na(value)),
    missing_pct = mean(is.na(value)) * 100,
    .groups = "drop"
  ) %>%
  arrange(date)
```

## Overall Missing Rate

```{r overall-stats}
overall_missing <- mean(is.na(data_long$value)) * 100

cat("📊 Overall Statistics:\n\n")
cat("Total observations:", nrow(data_long), "\n\n")
cat("Missing observations:", sum(is.na(data_long$value)), "\n\n")
cat("Overall missing rate:", sprintf("%.2f%%", overall_missing), "\n\n")
cat("\n\n")
cat("Industries:", n_distinct(data_long$industry_code), "\n\n")
cat("Questions:", n_distinct(data_long$question), "\n\n")
cat("Time points:", n_distinct(data_long$date), "\n\n")
```

## Missing by Level

```{r table-level}
missing_by_level %>%
  mutate(missing_pct = sprintf("%.2f%%", missing_pct)) %>%
  kable(
    caption = "Missing Rate by Hierarchy Level",
    col.names = c("Level", "# Industries", "Total Obs", "Missing", "Missing %")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Top 10 Missing Questions

```{r table-questions}
missing_by_question %>%
  head(10) %>%
  mutate(missing_pct = sprintf("%.2f%%", missing_pct)) %>%
  select(question, question_name, question_type, missing_pct) %>%
  kable(
    caption = "Top 10 Questions with Highest Missing Rate",
    col.names = c("Code", "Question Name", "Type", "Missing %")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# 2. Heatmap: Industry × Question Matrix

**Purpose**: Visualize missing patterns across industry-question combinations

```{r heatmap, fig.width=14, fig.height=12}
# Select top N industries with highest missing rates
top_n_industries <- 50

top_industries <- missing_by_industry %>%
  arrange(desc(missing_pct)) %>%
  head(top_n_industries) %>%
  pull(industry_code)

# Calculate missing rate matrix
missing_matrix <- data_long %>%
  filter(industry_code %in% top_industries) %>%
  group_by(industry_name, question_name) %>%
  summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop")

# Create heatmap
ggplot(missing_matrix, aes(x = question_name, y = industry_name, fill = missing_pct)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_viridis_c(
    option = "rocket",
    direction = -1,
    name = "Missing %",
    breaks = c(0, 25, 50, 75, 100)
  ) +
  labs(
    title = sprintf("Missing Value Heatmap: Top %d Industries with Highest Missing Rates", top_n_industries),
    subtitle = "Darker colors indicate higher missing rates",
    x = "Question",
    y = "Industry"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 7),
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    panel.grid = element_blank()
  )
```

**Interpretation**:

- Darker = High missing rate (close to 100%)
- Lighter = Low missing rate (close to 0%)
- If a row is entirely dark → Poor data quality for that industry
- If a column has large dark areas → That question is missing across many industries

---

# 3. Missing Rate by Hierarchy Level

**Purpose**: Compare data quality across different hierarchy levels

```{r by-level, fig.width=10, fig.height=6}
ggplot(missing_by_level, aes(x = factor(level), y = missing_pct)) +
  geom_col(aes(fill = missing_pct), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.1f%%\n(n=%d)", missing_pct, n_industries)),
    vjust = -0.5, 
    size = 3.5
  ) +
  scale_fill_viridis_c(
    option = "mako",
    direction = -1,
    name = "Missing %"
  ) +
  labs(
    title = "Missing Rate by Hierarchy Level",
    subtitle = "Label shows: Missing % (# of industries)",
    x = "Level",
    y = "Missing Rate (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "none"
  )
```

**Interpretation**:
- Level 0-1: Usually have lower missing rates (< 5%)
- Level 4-5: May have higher missing rates (15-30%)
- Level 6: Special aggregates, need separate evaluation

**Recommendation**: Prioritize using Level 0-2 for DFM analysis

---

# 4. Top 30 Industries with Highest Missing Rates

**Purpose**: Identify "problematic industries" that may need to be excluded

```{r top-industries, fig.width=12, fig.height=14}
top_n <- 30

top_data <- missing_by_industry %>%
  head(top_n) %>%
  mutate(
    industry_label = paste0(industry_code, "\n", 
                            str_trunc(industry_name, 40, "right"))
  )

ggplot(top_data, aes(x = reorder(industry_label, missing_pct), y = missing_pct)) +
  geom_col(aes(fill = level), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.1f%%", missing_pct)), 
    hjust = -0.2, 
    size = 3
  ) +
  scale_fill_viridis_c(
    option = "turbo",
    name = "Level"
  ) +
  coord_flip() +
  labs(
    title = sprintf("Top %d Industries with Highest Missing Rates", top_n),
    x = "Industry",
    y = "Missing Rate (%)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 8),
    legend.position = "right"
  ) +
  ylim(0, max(top_data$missing_pct) * 1.1)
```

**Decision Support**:
```{r decision-industries}
# Industries with > 50% missing rate
high_missing_industries <- missing_by_industry %>%
  filter(missing_pct > 50)

cat("Industries with > 50% missing rate:", nrow(high_missing_industries), "\n")
cat("Consider excluding these from analysis\n")
```

---

# 5. Missing Rate Over Time

**Purpose**: Discover temporal patterns in missing values

```{r time-overall, fig.width=12, fig.height=6}
ggplot(missing_by_time, aes(x = date, y = missing_pct)) +
  geom_line(color = "#440154", size = 1) +
  geom_smooth(method = "loess", color = "#FDE725", se = TRUE, alpha = 0.2) +
  labs(
    title = "Missing Rate Over Time",
    subtitle = "Yellow line shows smoothed trend",
    x = "Date",
    y = "Missing Rate (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

**Typical Pattern**:
- 1991-1997: High missing rate (30-40%) → Early data, incomplete
- 1997-2025: Low missing rate (< 5%) → Mature data period

**Recommendation**: Consider using data from 1997 onwards

---

# 6. Missing Rate by Question Type Over Time

**Purpose**: Identify which question types have more missing values

```{r time-by-type, fig.width=12, fig.height=6}
missing_by_time_question <- data_long %>%
  group_by(date, question_type) %>%
  summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop")

ggplot(missing_by_time_question, aes(x = date, y = missing_pct, color = question_type)) +
  geom_line(size = 1) +
  scale_color_viridis_d(option = "plasma", name = "Question Type") +
  labs(
    title = "Missing Rate Over Time by Question Type",
    x = "Date",
    y = "Missing Rate (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

**Interpretation**: Some question types (e.g., expectations) may have higher missing rates in early periods

---

# 7. Bubble Chart: Top 50 Missing Combinations

**Purpose**: Precisely locate industry×question combinations with high missing rates

```{r bubble, fig.width=14, fig.height=12}
top_n_combos <- 50

top_combos <- missing_by_combo %>%
  filter(missing_pct > 0) %>%
  arrange(desc(missing_pct)) %>%
  head(top_n_combos)

ggplot(top_combos, aes(x = question_name, y = industry_name)) +
  geom_point(aes(size = missing_pct, color = missing_pct), alpha = 0.7) +
  scale_size_continuous(
    range = c(2, 15),
    name = "Missing %"
  ) +
  scale_color_viridis_c(
    option = "inferno",
    direction = -1,
    name = "Missing %"
  ) +
  labs(
    title = sprintf("Top %d Missing Combinations (Industry × Question)", top_n_combos),
    subtitle = "Bubble size and color indicate missing rate",
    x = "Question",
    y = "Industry"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 7),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```

**Use Case**: Exclude specific industry-question combinations with extremely high missing rates

---

# 8. Level 1 Industries: Detailed Heatmap

**Purpose**: Focus on the 21 main industry sectors (most commonly used)

```{r level1-heatmap, fig.width=14, fig.height=10}
level1_data <- data_long %>%
  filter(level == 1) %>%
  group_by(industry_name, question_name) %>%
  summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop")

ggplot(level1_data, aes(x = question_name, y = industry_name, fill = missing_pct)) +
  geom_tile(color = "white", size = 0.8) +
  geom_text(
    aes(label = sprintf("%.0f%%", missing_pct)), 
    size = 2.5, 
    color = "white"
  ) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "#f7f7f7",
    high = "#b2182b",
    midpoint = 50,
    name = "Missing %",
    limits = c(0, 100)
  ) +
  labs(
    title = "Level 1 Main Industry Sectors: Missing Value Details",
    subtitle = "Numbers show exact missing rate percentages",
    x = "Question",
    y = "Industry Sector"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14),
    panel.grid = element_blank()
  )
```

**Why Important**: Level 1 represents the 21 main economic sectors, most frequently used in analysis

---

# 9. Data Quality Categories by Level

**Purpose**: Quick assessment of overall data quality across levels

```{r categories, fig.width=12, fig.height=7}
missing_categories <- data_long %>%
  group_by(industry_code, industry_name, level) %>%
  summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop") %>%
  mutate(
    category = case_when(
      missing_pct == 0 ~ "Complete (0%)",
      missing_pct < 5 ~ "Excellent (<5%)",
      missing_pct < 15 ~ "Good (5-15%)",
      missing_pct < 30 ~ "Fair (15-30%)",
      TRUE ~ "Poor (≥30%)"
    ),
    category = factor(
      category, 
      levels = c("Complete (0%)", "Excellent (<5%)", "Good (5-15%)", 
                 "Fair (15-30%)", "Poor (≥30%)")
    )
  )

category_by_level <- missing_categories %>%
  group_by(level, category) %>%
  summarise(count = n(), .groups = "drop")

ggplot(category_by_level, aes(x = factor(level), y = count, fill = category)) +
  geom_col(position = "stack", width = 0.7) +
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    color = "white",
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c("#2166ac", "#4393c3", "#92c5de", "#fddbc7", "#d6604d"),
    name = "Data Quality"
  ) +
  labs(
    title = "Data Quality Distribution by Hierarchy Level",
    subtitle = "Numbers show count of industries",
    x = "Level",
    y = "Number of Industries"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```

**Interpretation**:
- Blue shades = Good quality (Complete, Excellent, Good)
- Red shades = Poor quality (Fair, Poor)
- Level 0-1: Should be mostly blue
- If Level 2-3 has much red, consider using only Level 0-1

---

# 10. Summary


```{r summary-findings}
cat("📊 Key Findings:\n\n")

cat("1. Overall Missing Rate:", sprintf("%.2f%%\n", overall_missing))

cat("\n2. Level Analysis:\n")
missing_by_level %>%
  mutate(quality = ifelse(missing_pct < 5, "✓ Good", 
                         ifelse(missing_pct < 15, "⚠ Fair", "✗ Poor"))) %>%
  select(level, n_industries, missing_pct, quality) %>%
  print()

cat("\n3. Temporal Pattern:\n")
early_missing <- missing_by_time %>%
  filter(date < as.Date("1997-07-01")) %>%
  summarise(avg_missing = mean(missing_pct)) %>%
  pull(avg_missing)

recent_missing <- missing_by_time %>%
  filter(date >= as.Date("1997-07-01")) %>%
  summarise(avg_missing = mean(missing_pct)) %>%
  pull(avg_missing)

cat("Before 1997:", sprintf("%.2f%%\n", early_missing))
cat("1997 onwards:", sprintf("%.2f%%\n", recent_missing))

cat("\n4. Problem Industries (>30% missing):\n")
cat("Count:", nrow(filter(missing_by_industry, missing_pct > 30)), "\n")
```

## Recommendations for DFM Analysis

```{r recommendations}
cat("1. TIME RANGE:\n")
cat("   → Use data from 1997 onwards\n")
cat("   → Missing rate drops significantly after this point\n\n")

cat("2. HIERARCHY LEVEL:\n")
cat("   → Prioritize Level 0-2 for analysis\n")
cat("   → Level 3+ may have higher missing rates\n\n")

cat("3. INDUSTRY EXCLUSION:\n")
problem_industries <- missing_by_industry %>%
  filter(missing_pct > 30)
cat("   → Consider excluding", nrow(problem_industries), "industries with >30% missing\n")
cat("   → These represent", 
    sprintf("%.1f%%", nrow(problem_industries)/n_distinct(data_long$industry_code)*100),
    "of all industries\n\n")

cat("4. QUESTION SELECTION:\n")
problem_questions <- missing_by_question %>%
  filter(missing_pct > 20)
if (nrow(problem_questions) > 0) {
  cat("   → Review these questions with >20% missing:\n")
  print(problem_questions %>% select(question, question_name, missing_pct))
} else {
  cat("   → All questions have acceptable missing rates (<20%)\n")
}
```


```{r example-cleaning, include=FALSE}
# Example: Apply recommended filters
clean_data <- data_long %>%
  filter(
    level <= 2,                          # Use Level 0-2
    date >= as.Date("1997-07-01")       # Use 1997 onwards
  )

# Exclude high-missing industries
high_missing <- missing_by_industry %>%
  filter(missing_pct > 30) %>%
  pull(industry_code)

clean_data <- clean_data %>%
  filter(!industry_code %in% high_missing)

# Verify
final_missing <- mean(is.na(clean_data$value)) * 100

cat("\n Cleaned Data Summary:\n")
cat("Original observations:", nrow(data_long), "\n")
cat("Cleaned observations:", nrow(clean_data), "\n")
cat("Reduction:", sprintf("%.1f%%\n", (1 - nrow(clean_data)/nrow(data_long)) * 100))
cat("\nOriginal missing rate:", sprintf("%.2f%%\n", overall_missing))
cat("Cleaned missing rate:", sprintf("%.2f%%\n", final_missing))
cat("Improvement:", sprintf("%.2f percentage points\n", overall_missing - final_missing))
```

---




```{r}
# ===============================================================================
# BES and PRS Missing Value Analysis
# Investigate missing patterns for employment and production expectations
# ===============================================================================

library(tidyverse)
library(lubridate)

# Assuming you have 'results' loaded
# source("complete_data_processing.R")
# results <- main(DATA_PATH)

analyze_bes_prs_missing <- function(data_long) {
  
  cat("\n", rep("=", 80), "\n", sep = "")
  cat("  BES & PRS Missing Value Analysis\n")
  cat(rep("=", 80), "\n\n", sep = "")
  
  # Focus on BES and PRS
  bes_prs_data <- data_long %>%
    filter(question %in% c("BES", "PRS"))
  
  # Overall missing rate
  cat("📊 Overall Statistics:\n\n")
  
  overall_stats <- bes_prs_data %>%
    group_by(question, question_name) %>%
    summarise(
      total_obs = n(),
      missing_count = sum(is.na(value)),
      non_missing_count = sum(!is.na(value)),
      missing_pct = mean(is.na(value)) * 100,
      .groups = "drop"
    )
  
  print(overall_stats)
  
  # Missing by year
  cat("\n📅 Missing Rate by Year:\n\n")
  
  yearly_missing <- bes_prs_data %>%
    mutate(year = year(date)) %>%
    group_by(question, year) %>%
    summarise(
      n_obs = n(),
      missing_pct = mean(is.na(value)) * 100,
      .groups = "drop"
    ) %>%
    arrange(question, year)
  
  # Find the first year with data
  first_data <- yearly_missing %>%
    filter(missing_pct < 100) %>%
    group_by(question) %>%
    summarise(first_year = min(year), .groups = "drop")
  
  print(first_data)
  
  # Check missing pattern by level
  cat("\n🏗️ Missing Rate by Level:\n\n")
  
  level_missing <- bes_prs_data %>%
    group_by(question, level) %>%
    summarise(
      n_industries = n_distinct(industry_code),
      missing_pct = mean(is.na(value)) * 100,
      .groups = "drop"
    ) %>%
    arrange(question, level)
  
  print(level_missing)
  
  # Time series plot
  cat("\n📈 Creating visualizations...\n")
  
  # Plot 1: Missing rate over time
  p1 <- bes_prs_data %>%
    group_by(date, question) %>%
    summarise(missing_pct = mean(is.na(value)) * 100, .groups = "drop") %>%
    ggplot(aes(x = date, y = missing_pct, color = question)) +
    geom_line(size = 1.2) +
    geom_vline(xintercept = as.Date("1997-07-01"), 
               linetype = "dashed", color = "red", size = 1) +
    annotate("text", x = as.Date("1997-07-01"), y = 50, 
             label = "1997", color = "red", hjust = -0.2) +
    scale_color_viridis_d(option = "plasma", end = 0.8) +
    labs(
      title = "BES & PRS: Missing Rate Over Time",
      subtitle = "Red line marks 1997",
      x = "Date",
      y = "Missing Rate (%)",
      color = "Question"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
  
  print(p1)
  
  # Plot 2: Distribution of industries by first data year
  cat("\n📊 When did each industry start reporting BES/PRS?\n\n")
  
  industry_start <- bes_prs_data %>%
    filter(!is.na(value)) %>%
    mutate(year = year(date)) %>%
    group_by(question, industry_code, industry_name, level) %>%
    summarise(first_year = min(year), .groups = "drop")
  
  p2 <- industry_start %>%
    ggplot(aes(x = first_year, fill = question)) +
    geom_histogram(binwidth = 1, position = "dodge", alpha = 0.7) +
    scale_fill_viridis_d(option = "plasma", end = 0.8) +
    labs(
      title = "Distribution of Industry Start Years for BES & PRS",
      subtitle = "When did industries start reporting these questions?",
      x = "First Year with Data",
      y = "Number of Industries",
      fill = "Question"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
  
  print(p2)
  
  # Summary statistics
  cat("\n📋 Summary of Industry Start Years:\n\n")
  
  start_summary <- industry_start %>%
    group_by(question) %>%
    summarise(
      n_industries = n(),
      min_year = min(first_year),
      median_year = median(first_year),
      max_year = max(first_year),
      n_before_1997 = sum(first_year < 1997),
      n_1997_or_after = sum(first_year >= 1997),
      .groups = "drop"
    )
  
  print(start_summary)
  
  # Compare all questions
  cat("\n\n", rep("=", 80), "\n", sep = "")
  cat("  Comparison with Other Questions\n")
  cat(rep("=", 80), "\n\n", sep = "")
  
  all_questions_start <- data_long %>%
    filter(!is.na(value)) %>%
    mutate(year = year(date)) %>%
    group_by(question, question_name) %>%
    summarise(
      first_year = min(year),
      n_industries = n_distinct(industry_code),
      .groups = "drop"
    ) %>%
    arrange(first_year)
  
  cat("First year with data for each question:\n\n")
  print(all_questions_start)
  
  # Return results
  return(list(
    overall_stats = overall_stats,
    first_data_year = first_data,
    level_missing = level_missing,
    yearly_missing = yearly_missing,
    industry_start = industry_start,
    start_summary = start_summary,
    all_questions_start = all_questions_start,
    plots = list(
      time_series = p1,
      distribution = p2
    )
  ))
}

# ===============================================================================
# Strategy Comparison Function
# ===============================================================================

compare_strategies <- function(data_long) {
  
  cat("\n", rep("=", 80), "\n", sep = "")
  cat("  Strategy Comparison for DFM Analysis\n")
  cat(rep("=", 80), "\n\n", sep = "")
  
  # Strategy 1: Drop BES and PRS entirely
  strategy1 <- data_long %>%
    filter(!question %in% c("BES", "PRS"))
  
  # Strategy 2: Use all questions from 1997 onwards
  strategy2 <- data_long %>%
    filter(date >= as.Date("1997-01-01"))
  
  # Strategy 3: Use all questions but only 1991-1996 for available ones
  # This is not recommended for DFM, but let's see the stats
  
  cat("Strategy Comparison:\n\n")
  
  # Strategy 1 stats
  cat("STRATEGY 1: Drop BES & PRS, Use All Years\n")
  cat(rep("-", 50), "\n")
  cat("Time period: 1991-01 to", format(max(strategy1$date), "%Y-%m"), "\n")
  cat("Questions:", n_distinct(strategy1$question), "\n")
  cat("Industries:", n_distinct(strategy1$industry_code), "\n")
  cat("Total observations:", nrow(strategy1), "\n")
  cat("Missing rate:", sprintf("%.2f%%", mean(is.na(strategy1$value)) * 100), "\n")
  cat("Time points:", n_distinct(strategy1$date), "\n")
  cat("\n")
  
  # Strategy 2 stats
  cat("STRATEGY 2: Keep All Questions, Start from 1997\n")
  cat(rep("-", 50), "\n")
  cat("Time period: 1997-07 to", format(max(strategy2$date), "%Y-%m"), "\n")
  cat("Questions:", n_distinct(strategy2$question), "\n")
  cat("Industries:", n_distinct(strategy2$industry_code), "\n")
  cat("Total observations:", nrow(strategy2), "\n")
  cat("Missing rate:", sprintf("%.2f%%", mean(is.na(strategy2$value)) * 100), "\n")
  cat("Time points:", n_distinct(strategy2$date), "\n")
  cat("\n")
  
  # Create comparison table
  comparison <- tibble(
    Strategy = c("Drop BES & PRS", "Start from 1997"),
    Time_Period = c(
      sprintf("%s to %s", format(min(strategy1$date), "%Y-%m"), 
              format(max(strategy1$date), "%Y-%m")),
      sprintf("%s to %s", format(min(strategy2$date), "%Y-%m"), 
              format(max(strategy2$date), "%Y-%m"))
    ),
    N_Questions = c(n_distinct(strategy1$question), n_distinct(strategy2$question)),
    N_Industries = c(n_distinct(strategy1$industry_code), 
                     n_distinct(strategy2$industry_code)),
    N_Timepoints = c(n_distinct(strategy1$date), n_distinct(strategy2$date)),
    Total_Obs = c(nrow(strategy1), nrow(strategy2)),
    Missing_Pct = c(
      sprintf("%.2f%%", mean(is.na(strategy1$value)) * 100),
      sprintf("%.2f%%", mean(is.na(strategy2$value)) * 100)
    )
  )
  
  cat("\n📋 Side-by-Side Comparison:\n\n")
  print(comparison, n = Inf)
  
  # Visualize comparison
  cat("\n📈 Creating comparison visualization...\n")
  
  p <- comparison %>%
    pivot_longer(cols = c(N_Questions, N_Timepoints), 
                 names_to = "Metric", values_to = "Value") %>%
    mutate(
      Metric = case_when(
        Metric == "N_Questions" ~ "Number of Questions",
        Metric == "N_Timepoints" ~ "Number of Time Points"
      )
    ) %>%
    ggplot(aes(x = Strategy, y = Value, fill = Strategy)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = Value), vjust = -0.5, size = 4) +
    facet_wrap(~Metric, scales = "free_y", ncol = 2) +
    scale_fill_viridis_d(option = "mako", begin = 0.3, end = 0.7) +
    labs(
      title = "Strategy Comparison: Key Metrics",
      y = "Count"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "none",
      axis.text.x = element_text(angle = 15, hjust = 1)
    )
  
  print(p)
  
  return(list(
    strategy1_data = strategy1,
    strategy2_data = strategy2,
    comparison_table = comparison,
    comparison_plot = p
  ))
}


```

```{r}
# ===============================================================================
# Usage Example
# ===============================================================================

# # Run the analysis
bes_prs_analysis <- analyze_bes_prs_missing(results$data_long)
# 
# # Compare strategies
strategy_comparison <- compare_strategies(results$data_long)
# 
# # View results
print(bes_prs_analysis$start_summary)
print(strategy_comparison$comparison_table)
```








